### 2020-04-21 README.Rmd
us in

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# covid19census

<!-- badges: start -->
<!-- badges: end -->

The goal of covid19census is to ...

### 2020-04-21 R/datasets.R

@@ -1,0 +1,87 @@
+#' influenza vaccination 65 or older
+#'
+#' Percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination. Collected in 2017.
+#' @docType data
+#' @source \href{https://data.cms.gov/mapping-medicare-disparities}{Data.CMS.gov}
+#' @details \href{https://www.cms.gov/About-CMS/Agency-Information/OMH/Downloads/Mapping-Technical-Documentation.pdf}{Center for Medicare and Medicaid Services} and NORC at the University of Chicago.
+#' @return  tibble wotj `fl_65` indicating the percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination
+#' @usage data(fl65_us)
+"fl65_us"
+
+
+#' hospital beds
+#'
+#' beds of each hospital by county.
+#' @docType data
+#' @source \href{https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18}{Homeland Infrastructure Foundation-Level Data}
+#' @return  a tibble
+#' @usage data(hospbeds_us)
+"hospbeds_us"
+
+
+
+#' household composition
+#'
+#' Several metrics regarding household composition from the American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions}
+#' @return  a tibble
+#' @usage data(acm_househ_us)
+"acm_househ_us"
+
+#' age and sex
+#'
+#' Sex and age composition of the county population from the American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @return  a tibble
+#' @usage data(age_sex_us)
+"age_sex_us"
+
+#' poverty
+#'
+#' Household living below the poverty level, divided by age and race and calculate as absolute value or percentage. American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey}
+#' @return  a tibble
[…]

### 2020-04-21 R/getus_functions.R

@@ -1,0 +1,363 @@
+#' get COVID-19 updated cases
+#'
+#' extracts time series from the git repository of the \href{ https://github.com/nytimes/covid-19-data }{NYT}
+#'
+#' @return a tibble
+#' @importFrom rlang .data
+#' @importFrom magrittr %>%
+#' @import vroom
+#' @details `cases` represents the number of confirmed cases, while `cmr` the case-mortality rate (deaths / confirmed_case * 100).
+#' A good description of pitfalls and caveats associated with the use of case-mortality rate metric has been made on
+#' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
+#' @export
+get_covid <- function() {
+  dat <- vroom("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv",
+    col_types = cols(
+      date = col_date(format = ""),
+      county = col_character(),
+      state = col_character(),
+      fips = col_double(),
+      cases = col_double(),
+      deaths = col_double()
+    )
+  ) %>%
+    dplyr::mutate(cmr = .data$deaths / .data$cases * 100) %>%
+    dplyr::filter(!is.na(.data$fips))
+
+
+  message(paste0("Data of COVID-19 up to ", max(dat$date), " successfully extracted!"))
+
+  dat
+}
+
+#' get device-exposure indexes (DEX)
+#'
+#' extracts DEX from the git repository of the
+#' \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVID-19 exposure indeces}
+#'
+#' @return a tibble
+#' @details main metric is `dex_a`. In the \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{repository}, they
+#' explains: \cite{In the context of the ongoing pandemic, the DEX measure may be biased if devices sheltering-in-place
+#' are not in the sample due to lack of movement. We report adjusted DEX values to help address this selection bias.
+#' DEX-adjusted is computed assuming that the number of devices has not declined since the early-2020 peak
+#' and that unobserved devices did not visit any commercial venues.} Datataset is updated by the mantainers every weekend.
+#' @importFrom rlang .data
+#' @importFrom magrittr %>%
+#' @import vroom
+#' @export
+get_dex <- function() {
+  dat <- vroom("https://raw.githubusercontent.com/COVIDExposureIndices/COVIDExposureIndices/master/dex_data/county_dex.csv",
[…]

### 2020-04-21 R/recode_col.R

@@ -1,0 +1,13 @@
+#' recode column.
+#'
+#' recode_col recodes a vector with codes provided in another vector. The unique values of
+#' the 2 vectors need to be of same length. Codes are assigned matching unique values by position. Used internally.
+#' @param x vectore to recode.
+#' @param repl vector containing unique codes.
+#' @return a vector of same length of x.
+recode_col <- function(x, repl) {
+  intv_id <- as.character(unique(x))
+  id_recode <- repl[seq_along(intv_id)]
+  newnames <- stats::setNames(id_recode, intv_id)
+  as.vector(newnames[x])
+}

### 2020-04-21 data-raw/import_raw_us.R

@@ -1,0 +1,269 @@
+library(vroom)
+library(tidyverse)
+library(janitor)
+library(reshape2)
+
+
+# United States #-----
+
+# age and sex- age_sex_us ---------
+
+age_sex_all_us <- vroom("data-raw/us/american_comunity_survey_2018/age_sex/ACSST5Y2018.S0101_data_with_overlays_2020-04-14T063202.csv",
+                        skip = 1,
+                        col_select = c(`id`, `Geographic Area Name`, starts_with("Estimate"))
+) %>%
+  clean_names()
+
+
+to_select_age_sex <- readxl::read_xlsx("data-raw/us/american_comunity_survey_2018/age_sex/to_select_age_sex.xlsx")
+
+age_sex_us <- age_sex_all_us %>%
+  select(!!to_select_age_sex$old_names)
+
+
+names(age_sex_us) <- to_select_age_sex$new_names
+
+age_sex_us <- age_sex_us %>%
+  separate(id, into = c("us_id", "fips"), sep = 9) %>%
+  select(-us_id) %>%
+  mutate(fips = as.numeric(fips))
+
+
+# flu vax - fl65_us  ----------
+fl65_all_us <- vroom("data-raw/us/flu_us.csv",
+  col_types = cols(
+    year = col_double(),
+    geography = col_character(),
+    measure = col_character(),
+    adjustment = col_character(),
+    analysis = col_character(),
+    domain = col_character(),
+    condition = col_character(),
+    primary_sex = col_character(),
+    primary_age = col_character(),
+    primary_dual = col_character(),
+    fips = col_double(),
+    county = col_character(),
+    state = col_character(),
+    urban = col_character(),
+    primary_race = col_character(),
+    primary_denominator = col_character(),
+    analysis_value = col_double()
+  )
+)
+
+fl65_us <- fl65_all_us %>%
+  rename("imm65" = "analysis_value") %>%
+  dplyr::select(state, county, fips, imm65)
+
+# hospital beds - hospbeds_us  ----------
+hospbeds_all_us <- vroom("data-raw/us/hospital_beds.csv",
+  col_types = cols(
+    X = col_double(),
+    Y = col_double(),
+    OBJECTID = col_double(),
+    ID = col_character(),
+    NAME = col_character(),
+    ADDRESS = col_character(),
+    CITY = col_character(),
+    STATE = col_character(),
+    ZIP = col_character(),
+    ZIP4 = col_character(),
[…]

### 2020-04-21 tests/testthat/test-get.R

@@ -1,0 +1,59 @@
+library(dplyr)
+
+test_that("get_all executes", {
+  expect_error(get_all(), NA)
+})
+
+test_that("get_covid executes", {
+  expect_error(get_covid(), NA)
+})
+
+y <- get_all()
+
+x <- get_covid()
+
+test_that("get_all and get_covid have same fips", {
+  expect_equal(length(unique(y$fips)), length(unique(x$fips)))
+})
+
+
+###################################
+
+test_that("get_dex executes", {
+  expect_error(get_dex(), NA)
+})
+
+##################################
+
+
+expected_states <- structure(list(state = c(
+  "Alaska", "Alabama", "Arizona", "Arkansas",
+  "California", "Colorado", "Connecticut", "District of Columbia",
+  "Delaware", "Florida", "Georgia", "Hawaii", "Iowa", "Idaho",
+  "Illinois", "Indiana", "Kansas", "Kentucky", "Louisiana", "Massachusetts",
+  "Maryland", "Maine", "Michigan", "Minnesota", "Missouri", "Mississippi",
+  "Montana", "North Carolina", "North Dakota", "Nebraska", "New Hampshire",
+  "New Jersey", "New Mexico", "Nevada", "New York", "Ohio", "Oklahoma",
+  "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
+  "Tennessee", "Texas", "Utah", "Virginia", "Vermont", "Washington",
+  "Wisconsin", "West Virginia", "Wyomin"
+), abbr = c(
+  "AK", "AL",
+  "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI", "IA",
+  "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN",
+  "MO", "MS", "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY",
+  "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VA",
+  "VT", "WA", "WI", "WV", "WY"
+)), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA, -51L))
+
+test_that("get_test executes", {
+  expect_error(get_tests(), NA)
+})
+
+tests <- get_tests() %>%
+  select(state, abbr) %>%
+  distinct()
+
+test_that("get_tests states abbr ", {
+  expect_equal(tests, expected_states)
+})

### 2020-04-21 R/getus_functions.R
getfunc

@@ -10,7 +10,7 @@
 #' A good description of pitfalls and caveats associated with the use of case-mortality rate metric has been made on
 #' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
 #' @export
-get_covid <- function() {
+getus_covid <- function() {
   dat <- vroom("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv",
     col_types = cols(
       date = col_date(format = ""),
@@ -45,7 +45,7 @@ get_covid <- function() {
 #' @importFrom magrittr %>%
 #' @import vroom
 #' @export
-get_dex <- function() {
+getus_dex <- function() {
   dat <- vroom("https://raw.githubusercontent.com/COVIDExposureIndices/COVIDExposureIndices/master/dex_data/county_dex.csv",
     skip = 1,
     col_names = c(
@@ -106,7 +106,7 @@ get_dex <- function() {
 #' Other details regarding the score system used are reported in the \href{https://covidtracking.com/about-data}{maintainers webpage}
 #'  @import vroom
 #'  @export
-get_tests <- function() {
+getus_tests <- function() {
   dat <- vroom::vroom("https://covidtracking.com/api/states/daily.csv",
     col_types = cols(
       date = col_date(format = "%Y%m%d"),
@@ -142,11 +142,11 @@ get_tests <- function() {
 
   message(paste0("Data of tests up to ", max(dat$date, na.rm = TRUE), " successfully imported!"))
 
-  # we don't have unincorporated territories in get_all()
+  # we don't have unincorporated territories in getus_all()
   dat2 <- dat[dat$abbr %in% state_abbr$abbr, ]
 
   # we use the function recode_col from the current package to recode the state abbr to state names
-  # that are exactly the same of those generate by get_all()
+  # that are exactly the same of those generate by getus_all()
   dat2$state <- recode_col(dat2$abbr, state_abbr$state)
 
   dat2 <- dat2[, c(
@@ -331,14 +331,14 @@ get_tests <- function() {
 #' @importFrom rlang .data
 #' @importFrom magrittr %>%
 #' @import vroom
-#' @seealso \code{\link{get_covid}},\code{\link{get_tests}}, \code{\link{get_dex}},
+#' @seealso \code{\link{getus_covid}},\code{\link{getus_tests}}, \code{\link{getus_dex}},
 #' @export
[…]

### 2020-04-21 tests/testthat/test-get.R

@@ -1,16 +1,16 @@
 library(dplyr)
 
 test_that("get_all executes", {
-  expect_error(get_all(), NA)
+  expect_error(getus_all(), NA)
 })
 
 test_that("get_covid executes", {
-  expect_error(get_covid(), NA)
+  expect_error(getus_covid(), NA)
 })
 
-y <- get_all()
+y <- getus_all()
 
-x <- get_covid()
+x <- getus_covid()
 
 test_that("get_all and get_covid have same fips", {
   expect_equal(length(unique(y$fips)), length(unique(x$fips)))
@@ -20,7 +20,7 @@ test_that("get_all and get_covid have same fips", {
 ###################################
 
 test_that("get_dex executes", {
-  expect_error(get_dex(), NA)
+  expect_error(getus_dex(), NA)
 })
 
 ##################################
@@ -47,13 +47,13 @@ expected_states <- structure(list(state = c(
 )), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA, -51L))
 
 test_that("get_test executes", {
-  expect_error(get_tests(), NA)
+  expect_error(getus_tests(), NA)
 })
 
-tests <- get_tests() %>%
+tests <- getus_tests() %>%
   select(state, abbr) %>%
   distinct()
 
-test_that("get_tests states abbr ", {
+test_that("getus_tests states abbr ", {
   expect_equal(tests, expected_states)
 })

### 2020-04-21 data-raw/it/how_get_data.Rmd
success merging of the 2 packages

---
title: "Intersection COVID-19: Italy"
date: "`r Sys.time()`"
categories:
  - covid-19
tags: 
  - covid-19
  - Italy
lastmod: "`r Sys.time()`"
keywords: []
description: ''
output:
  html_document:
    df_print: kable
    toc: true
    theme: journal
---

# Italy

## Influenza Vaccination Coverage

Data of the influenza vaccination coverage in Italy can be found in the web site of ["Ministero della Salute"](http://www.salute.gov.it/imgs/C_17_tavole_19_allegati_iitemAllegati_0_fileAllegati_itemFile_3_file.pdf). Data are from 1999‐2000 to 2018‐2019. So... pretty cool.

Data of the general population are in a `pdf` table in page 1. We can use `{tabulizer}` to get them. **Last data are 07-15-2019**

```{r dropbox, include=FALSE}
# this make it possible to get and write data directly on the cloud independently of the 
# computer/operating  system and so on.
# https://github.com/karthik/rdrop2

library(rdrop2)
# log in dropbox
drop_auth()

```

```{r getting_it_im, warning=FALSE, message=FALSE}
library(DT)
library(Hmisc)
library(rvest)
library(htmltools)
library(htmlwidgets)
library(janitor)
library(lubridate)
library(lemon)
library(miniUI)
library(plotly)
library(scales)
library(tabulizer) # this library require specific version of Java and can be a pain to install
library(tidyverse)
library(xml2)
library(vroom)

web_page <- "http://www.salute.gov.it/imgs/C_17_tavole_19_allegati_iitemAllegati_0_fileAllegati_itemFile_3_file.pdf"

area_sel <- c(
  top = 92.6606873324687, left = 49.819066386841, bottom = 436.401946791649,
  right = 772.17388698945
) # got this using locate_areas()

suppressWarnings(
  fl_tab <- extract_tables(
    web_page,
    output = "data.frame",
    pages = 1,
    area =  list(area_sel),
    guess = FALSE
  )[[1]]
)

# inFLuenza coverage ITaly

fl_it <- fl_tab %>%
  rename(region = X) %>%
  {
    x <- .
    x <- x[3:nrow(x), ]
    names(x) <- str_remove_all(names(x), "[[:punct:]]|X")
    x
  } %>%
  mutate_all(list(~ gsub(., pattern = ",", replacement = "."))) %>%
  rename("2019" = ncol(.)) %>%
  {
    x <- .
    x[x == ""] <- NA
    x$region[x$region == "P. A. Trento"] <- "P.A. Trento"
    x <- x[!is.na(x$region), ]
    x
  }

# fl_it
```

Let's plot

```{r plot_fl_it2019}
fl_it_2019 <- fl_it %>%
  rename(perc_imm = "2019") %>%
  mutate(perc_imm = as.numeric(perc_imm)) %>%
  select(region, perc_imm)

fl_it_2019 %>%
  ggplot(aes(reorder(region, perc_imm), perc_imm)) +
  geom_col(fill = "grey") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = NULL, y = NULL, title = "% of population that received influenza vaccination")
```

We actually have influenza vaccination coverage for people that are 65 > 
It is in the same file but at page 3

```{r fl65_it}
area_sel2 <- c(
  top = 75.63785, left = 51.12191, bottom = 434.73870 ,
  right =   795.57018 
) # got this using locate_areas()

suppressWarnings(
  fl65_tab <- extract_tables(
    web_page,
    output = "data.frame",
    pages = 3,
    area =  list(area_sel2),
    guess = FALSE
  )[[1]]
)

# inFLuenza coverage ITaly

fl65_it <- fl65_tab %>%
  rename(region = X) %>%
  {
    x <- .
    x <- x[3:nrow(x), ]
    names(x) <- str_remove_all(names(x), "[[:punct:]]|X")
    x
  } %>%
  mutate_all(list(~ gsub(., pattern = ",", replacement = "."))) %>%
  rename("2019" = ncol(.)) %>%
  {
    x <- .
    x[x == ""] <- NA
    x$region[x$region == "P. A. Trento"] <- "P.A. Trento"
    x <- x[!is.na(x$region), ]
    x
  }

fl65_it_2019 <- fl65_it %>%
  rename(perc_imm65 = "2019") %>%
  mutate(perc_imm65 = as.numeric(perc_imm65)) %>%
  select(region, perc_imm65)
```

Let's put it together with the other flu data

```{r fl_it_all}
fl_it_2019 <- inner_join(fl65_it_2019, fl_it_2019, by = "region")
```

## Demographics

Istat maintains a database of a variety of statistical indicators accessible from its [website](http://dati.istat.it/). 
Info on the age of the population can be found going to `Popolazione famiglie` > `Popolazione resident al 1 gennaio` > `Per fascia d'eta`.

Details about the dataset can be found [here](http://siqual.istat.it/SIQual/visualizza.do?id=0019900&refresh=true&language=IT)

Let's check the demographic.

```{r dem_it_p}
dem_it_p <- vroom(
  "data/italy/pop_it.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO15 = col_character(),
    `Tipo di indicatore demografico` = col_character(),
    SEXISTAT1 = col_double(),
    Sesso = col_character(),
    ETA1 = col_character(),
    Età  = col_character(),
    STATCIV2 = col_double(),
    `Stato civile` = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_character(),
    Flags = col_character()
  )
) %>%
  rename(
    region = Territorio,
    index = `Tipo di indicatore demografico`,
    sex = Sesso,
    age = `Età`,
    marital_status = `Stato civile`,
    year = `Seleziona periodo`,
    value = Value
  ) %>%
  select(region, index, sex, age, marital_status, year, value) %>%
  filter(age != "totale", sex != "totale", marital_status == "totale") %>%
  mutate(
    age = as.numeric(str_remove_all(age, "[[:alpha:]]")),
    sex = recode(sex, femmine = "female", maschi = "male")
  )

dem_it_p$region <- recode(
  dem_it_p$region,
  "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
  "Provincia Autonoma Bolzano / Bozen" = "P.A. Bolzano",
  "Provincia Autonoma Trento"  = "P.A. Trento",
  "Friuli-Venezia Giulia"  = "Friuli Venezia Giulia",
  "Emilia-Romagna" = "Emilia Romagna"
)
```

These data have `Trentino Alto Adige / Südtirol` but also `PA Trento` and `PA Bolazono`. Let's see if `Trentino Alto Adige` is the sum of the 2.

```{r trentino}
dem_it_p %>%
  filter(region %in% list("P.A. Bolzano", "P.A. Trento", "Trentino Alto Adige / Südtirol")) %>%
  group_by(region) %>%
  summarize(tot = sum(value))
```

Yes, it is, we need to remove it.
Let's do that and compute some percentage.

```{r percent}
dem_it_p <- dem_it_p %>%
  filter(region != "Trentino Alto Adige / Südtirol") %>%
  group_by(region) %>%
  mutate(tot = sum(value)) %>% # this makes total for regions
  ungroup() %>%
  mutate(perc_pop = value / tot * 100)
```

```{r plot_dem, fig.height= 10, fig.width=10}
dem_it_p %>%
  ggplot(aes(age, perc_pop, group = sex, color = sex)) +
  geom_line() +
  facet_wrap(~region)
```

The Istituto Superiore Sanità (HHS) provides info regarding the mortality in different age groups fro patients positive for COVID-19 ([link](https://www.epicentro.iss.it/coronavirus/sars-cov-2-decessi-italia)). In an [initial report](https://www.epicentro.iss.it/coronavirus/sars-cov-2-decessi-italia) the age variable was in bin of 10 years (0-9, 10-19..), left closed and right closed; ages more than 90 were binned together.

If we want to consider some variables with modeling packages, we might need a wider format. Let's get that too.

```{r dem_i_binned}
dem_it_p <- dem_it_p %>%
  mutate(age_bins = cut(age, seq(0, 100, 10), right = FALSE, include.lowest = TRUE)) %>%
  mutate(age_bins = recode(age_bins, "[90,100]" = "[90,100+]")) %>%
  group_by_at(vars(-age, -value, -marital_status, -index, -year, -perc_pop)) %>%
  summarize(value = sum(value), perc_pop = sum(perc_pop)) %>%
  ungroup()

# if we want to use caret or any package for regression we need to have a variable age-sex bins as columns
dem_it_wider <- dem_it_p %>%
  mutate(sex_age_bin = paste0("perc_", sex, "_", age_bins)) %>% # now that we have the perc we create the new factor
  select(region, tot, perc_pop, sex_age_bin) %>%
  # now we do the magic
  pivot_wider(names_from = sex_age_bin, values_from = perc_pop) %>%
  rename(pop_tot = tot)

glimpse(dem_it_p)
```

Let's check if the percentage adds up to 100

```{r check}
apply(dem_it_wider[, 3:ncol(dem_it_wider)], 1, sum)
```

We have the demographic data.

**We might want to add the population for square kilometer.**

We can get that from wikipedia.

```{r regions_area}
wiki <- xml2::read_html("https://it.wikipedia.org/wiki/Regioni_d%27Italia")

regions_area_p <- wiki %>%
  rvest::html_nodes("table") %>%
  rvest::html_table(fill = TRUE) %>%
  {
    x <- .
    x[[3]]
  }

# let's clean the names
regions_area <- regions_area_p %>%
  select(Regione, `Superficie (km²)`) %>%
  rename(region = Regione, area_km2 = `Superficie (km²)`) %>%
  mutate(region = str_replace(region, "-", " "), area_km2 = as.numeric(str_remove(area_km2, "[[:blank:]]"))) %>%
  filter(!region %in% c("Trentino Alto Adige", "Italia"))

#  we add those missing regions
regions_area <- bind_rows(regions_area, tribble(
  ~region, ~area_km2,
  "P.A. Bolzano", 7398,
  "P.A. Trento", 2397
))
```

## Health

### Chronic Conditions

On the Istat [website](http://dati.istat.it/) in `Salute e Sanità` > `Stato di salute e consumo di farmaci` > `Stato di Salute in regioni e comuni`. Details on the datasets can be found [here](http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=IT)

Data are of **2018**

```{r healt_it}

chronic_it_p <- vroom("data/italy/chronic_conditions.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO_AVQ = col_character(),
    `Tipo dato` = col_character(),
    MISURA_AVQ = col_character(),
    Misura = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_logical(),
    Flags = col_logical()
  )
) %>%
  clean_names() %>%
  filter(misura == "valori in migliaia") %>% # we keep the absolute values
  mutate(value = value * 1000) # values are in thousands

# we do some cleaning and translations
chronic_it <- chronic_it_p %>%
  rename(region = territorio, ch_condition = tipo_dato_avq) %>%
  mutate(region = str_replace(region, "-", " ")) %>%
  filter(!region %in% c("Trentino Alto Adige / Südtirol", "Italia")) %>%
  # make name consistent
  mutate(region = recode(region,
    "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
    "Provincia Autonoma Bolzano / Bozen" =  "P.A. Bolzano",
    "Provincia Autonoma Trento"   =  "P.A. Trento",
    "Fiuli Venezia Giulia"  = "Friuli Venezia Giulia"
  )) %>%
  select(region, ch_condition, value) %>%
  mutate(ch_condition = tolower(str_remove(ch_condition, "0_"))) %>%
  # we make it wilder
  pivot_wider(names_from = ch_condition, values_from = value) %>%
  rename(chronic_aleast_2cron = aleast_2cron, chronic_aleast_1cron = aleast_1cron, chronic_good_h = good_h)
```

### Cigarette Smoking

On the Istat [website](http://dati.istat.it/) in `Salute e Sanità` > `Stili di vita e fattori di rischio` > `Abitudini al fumo`, `Regioni`. Details on the datasets can be found [here](http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=IT).

 **14 years old or more**

```{r smoking_it, warning=FALSE, message=FALSE}
smoking_it <- vroom("data/italy/smoking.csv",
  col_types = cols(
    ITTER107 = col_character(),
    Territorio = col_character(),
    TIPO_DATO_AVQ = col_character(),
    `Tipo dato` = col_character(),
    MISURA_AVQ = col_character(),
    Misura = col_character(),
    TIME = col_double(),
    `Seleziona periodo` = col_double(),
    Value = col_double(),
    `Flag Codes` = col_logical(),
    Flags = col_logical()
  )
) %>%
  rename(
    region = Territorio,
    smoking_status = TIPO_DATO_AVQ,
    value = Value
  ) %>%
  filter(Misura == "valori in migliaia") %>%
  select(region, smoking_status, value) %>%
  mutate(region = str_replace(region, "-", " ")) %>%
  filter(!region %in% c("Trentino Alto Adige / Südtirol", "Italia")) %>%
  # make name consistent
  mutate(region = recode(region,
    "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
    "Provincia Autonoma Bolzano / Bozen" =  "P.A. Bolzano",
    "Provincia Autonoma Trento"   =  "P.A. Trento",
    "Fiuli Venezia Giulia"  = "Friuli Venezia Giulia"
  )) %>%
  mutate(smoking_status = recode(smoking_status,
    "14_FUMO_SI" = "smoking_current",
    "14_FUMO_EX" = "smoking_ex",
    "14_FUMO_NO" = "smoking_no"
  )) %>%
  filter(smoking_status %in% c("smoking_current", "smoking_ex", "smoking_no")) %>%
  # values are in thousands
  mutate(value = value * 1000) %>%
  pivot_wider(names_from = smoking_status, values_from = value)
```

### Obesity

On the Istat [website](http://dati.istat.it/) in `Salute e Sanità` > `Stili di vita e fattori di rischio` > `Indici di Massa Corporea`, `Regioni`. Details on the datasets can be found [here](http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=IT).

**18 years old or more**

```{r obesity_it, warning=FALSE, message=FALSE}
bweight_it <- vroom("data/italy/bweight_it.csv", col_types = cols(
  ITTER107 = col_character(),
  Territorio = col_character(),
  TIPO_DATO_AVQ = col_character(),
  `Tipo dato` = col_character(),
  MISURA_AVQ = col_character(),
  Misura = col_character(),
  SEXISTAT1 = col_double(),
  Sesso = col_character(),
  TIME = col_double(),
  `Seleziona periodo` = col_double(),
  Value = col_double(),
  `Flag Codes` = col_logical(),
  Flags = col_logical()
)) %>%
  filter(TIME == 2018) %>%
  rename(
    region = Territorio,
    bweight_status = TIPO_DATO_AVQ,
    value = Value,
    ses = Sesso
  ) %>%
  filter(Misura == "valori in migliaia") %>%
  select(region, bweight_status, value) %>%
  mutate(region = str_replace(region, "-", " ")) %>%
  filter(!region %in% c("Trentino Alto Adige / Südtirol", "Italia")) %>%
  # make name consistent
  mutate(region = recode(region,
    "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
    "Provincia Autonoma Bolzano / Bozen" =  "P.A. Bolzano",
    "Provincia Autonoma Trento"   =  "P.A. Trento",
    "Fiuli Venezia Giulia"  = "Friuli Venezia Giulia"
  )) %>%
  mutate(bweight_status = recode(bweight_status,
    "18_BMI_SOTTO" = "bweight_under",
    "18_BMI_NORMO" = "bweight_normal",
    "18_BMI_SOVRA" = "bweight_over",
    "18_BMI_OBE" = "bweight_obese"
  )) %>%
  filter(bweight_status %in% c("bweight_under", "bweight_normal", "bweight_over", "bweight_obese")) %>%
  # values are in thousands
  mutate(value = value * 1000) %>%
  pivot_wider(names_from = bweight_status, values_from = value)
```

### Cancer

Data on tumor prevalence (year: ##2016##) can be found in the website of Istituto Superiore Sanità [tumors](http://www.registri-tumori.it/PDF/AIOM2016/I_numeri_del_cancro_2016.pdf). Tab is at pag 35.

** year was 2016 **

```{r cancer_it}
web_page <- "http://www.registri-tumori.it/PDF/AIOM2016/I_numeri_del_cancro_2016.pdf"

area_sel <- c(
  top = 240.95886, left = 27.75322, bottom = 534.19809,
  right = 436.61248
) # got this using locate_areas()

suppressWarnings(
  cancer_tab <- extract_tables(
    web_page,
    output = "data.frame",
    pages = 35,
    area =  list(area_sel),
    guess = FALSE
  )[[1]]
  %>%
    na.omit()
)

names(cancer_tab) <-
  c(
    "region",
    paste("cancer",
      c(
        "all",
        "breast",
        "colon",
        "prostate",
        "bladder",
        "lymphoma",
        "head_neck",
        "uterus",
        "lung"
      ),
      sep = "_"
    )
  )

# clean names
cancer_it <- cancer_tab %>%
  mutate(region = recode(region,
    "Trentino Alto" = "Trentino Alto Adige",
    "Emilia" = "Emilia Romagna",
    "Friuli Venezia" = "Friuli Venezia Giulia",
    "Valle D’Aosta" = "Valle d'Aosta"
  )) %>%
  filter(region != "Trentino Alto Adige") %>%
  # in italy . used as , in numbers
  mutate_if(is.double, ~ . * 1000)

# no data on those  but we want to include them anyway for the join

cancer_it[20, "region"] <- "P.A. Bolzano"
cancer_it[21, "region"] <- "P.A. Trento"
```

## COVID-19 Cases

Data of COVID-19 cases and many more info can be found on the website of "protezione civile" ([here]("https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv")).

```{r get_data_cases_it}
get_cases_italy <- function() {
  dat <- vroom(
    file = "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",
    col_types = cols(
      data = col_datetime(format = ""),
      stato = col_character(),
      codice_regione = col_double(),
      denominazione_regione = col_character(),
      lat = col_double(),
      long = col_double(),
      ricoverati_con_sintomi = col_double(),
      terapia_intensiva = col_double(),
      totale_ospedalizzati = col_double(),
      isolamento_domiciliare = col_double(),
      totale_positivi = col_double(),
      variazione_totale_positivi = col_double(),
      nuovi_positivi = col_double(),
      dimessi_guariti = col_double(),
      deceduti = col_double(),
      totale_casi = col_double(),
      tamponi = col_double(),
      note_it = col_character(),
      note_en = col_character()
    )
  ) %>%
    rename(region = denominazione_regione) %>%
    mutate(region = str_replace(region, "-", " ")) %>%
    select(-totale_positivi, -variazione_totale_positivi, -nuovi_positivi, -dimessi_guariti) %>% 
      rename(
    date = data,
    state = stato,
    region_cod = codice_regione,
    hospitalized_with_symptoms = ricoverati_con_sintomi,
    intensive_care_unit = terapia_intensiva,
    total_hospitalized = totale_ospedalizzati,
    home_quarantine = isolamento_domiciliare,
    deaths = deceduti,
    tests = tamponi,
    total_cases = totale_casi
  ) %>%
  # select(!c("totale_attualmente_positivi", "nuovi_attualmente_positivi" )) %>%
  mutate(
    cmr = deaths / total_cases * 100,
    chr = total_hospitalized / total_cases * 100,
    cir = intensive_care_unit / total_cases * 100
  ) %>%
  pivot_longer(cols = c("cmr", "chr", "cir", "hospitalized_with_symptoms", "intensive_care_unit", "total_hospitalized", "home_quarantine", "deaths", "total_cases", "tests"), names_to = "COVID19_var", values_to = "value")

  dat
}
```

## Join togheter and correlate

Let's put together the datasets of influenza coverage and the all the others.

```{r get_cmr}
cmr_it <- get_cases_italy()
```

```{r join_all}
# nuovi attualmente positivi not reliable...look at critiques in recent Italian news
# Join fl with cmr
fl_cmr_it <- inner_join(cmr_it, fl_it_2019, by = "region") %>%
  mutate_at(vars("perc_imm65","perc_imm"), as.numeric)

# join  with demographics
fl_cmr_dem_it <- inner_join(fl_cmr_it, dem_it_wider, by = "region")

# adding regions km2
fl_cmr_dem_it <- inner_join(fl_cmr_dem_it, regions_area, by = "region") %>%
  mutate(pop_km2 = pop_tot / area_km2) %>%
  # reorder col
  select(c(
    "date", "state", "region_cod", "region", "lat", "long", "note_it", "note_en",
    "perc_imm", "perc_imm65", "COVID19_var", "value", "pop_tot", "area_km2", "pop_km2", "perc_female_[0,10)",
    "perc_female_[10,20)", "perc_female_[20,30)", "perc_female_[30,40)",
    "perc_female_[40,50)", "perc_female_[50,60)", "perc_female_[60,70)",
    "perc_female_[70,80)", "perc_female_[80,90)", "perc_female_[90,100+]",
    "perc_male_[0,10)", "perc_male_[10,20)", "perc_male_[20,30)",
    "perc_male_[30,40)", "perc_male_[40,50)", "perc_male_[50,60)",
    "perc_male_[60,70)", "perc_male_[70,80)", "perc_male_[80,90)",
    "perc_male_[90,100+]"
  ))

# join health data
fl_cmr_dem_chronic_it <- inner_join(fl_cmr_dem_it, chronic_it, by = "region")

# join cancer data
fl_cmr_dem_health_cancer_it <- inner_join(fl_cmr_dem_chronic_it, cancer_it, by = "region")

# join smoking data
fl_cmr_dem_health_cancer_smoking_it <- inner_join(fl_cmr_dem_health_cancer_it, smoking_it, by = "region")

# join bweight
fl_cmr_dem_health_cancer_smoking_bweight_it <- inner_join(fl_cmr_dem_health_cancer_smoking_it, bweight_it, by = "region")

# something wrong with bweight

all_dat <- fl_cmr_dem_health_cancer_smoking_bweight_it

# make cancer and health as percentage
all_dat <- all_dat %>%
  mutate_at(vars(chronic_osteo:bweight_obese), list(~ (. / pop_tot * 100))) %>% 
  mutate(date = as.Date(date, "%Y-%m-%d"))

# all data are as percentage so we remove the perc_ from names
names(all_dat) <- str_remove(names(all_dat), "perc_")

glimpse(all_dat)
```

Now we have our datasets ready for being analyzed...what we need to do is just filter for the `COVID19_var` and date of interest and then run the stats.

### 2020-04-21 R/datasets.R

@@ -1,87 +1,0 @@
-#' influenza vaccination 65 or older
-#'
-#' Percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination. Collected in 2017.
-#' @docType data
-#' @source \href{https://data.cms.gov/mapping-medicare-disparities}{Data.CMS.gov}
-#' @details \href{https://www.cms.gov/About-CMS/Agency-Information/OMH/Downloads/Mapping-Technical-Documentation.pdf}{Center for Medicare and Medicaid Services} and NORC at the University of Chicago.
-#' @return  tibble wotj `fl_65` indicating the percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination
-#' @usage data(fl65_us)
-"fl65_us"
-
-
-#' hospital beds
-#'
-#' beds of each hospital by county.
-#' @docType data
-#' @source \href{https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18}{Homeland Infrastructure Foundation-Level Data}
-#' @return  a tibble
-#' @usage data(hospbeds_us)
-"hospbeds_us"
-
-
-
-#' household composition
-#'
-#' Several metrics regarding household composition from the American Community Survey of 2018
-#' @docType data
-#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
-#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions}
-#' @return  a tibble
-#' @usage data(acm_househ_us)
-"acm_househ_us"
-
-#' age and sex
-#'
-#' Sex and age composition of the county population from the American Community Survey of 2018
-#' @docType data
-#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
-#' @return  a tibble
-#' @usage data(age_sex_us)
-"age_sex_us"
-
-#' poverty
-#'
-#' Household living below the poverty level, divided by age and race and calculate as absolute value or percentage. American Community Survey of 2018
-#' @docType data
-#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
-#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey}
-#' @return  a tibble
[…]

### 2020-04-21 R/datasets_it.R

@@ -1,0 +1,139 @@
+#' body-mass index
+#'
+#' Body mass index in regions of Italy, in the general population.
+#' Data were collected in 2018 and indicate absolute number of people underweight, normalweight, overweight or obese.
+#' @docType data
+#' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
+#' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
+#' @usage data(bweight_it)
+"bweight_it"
+
+#' cancer patients
+#'
+#' Number of cancer patients in each region by type.
+#' Data were collected in 2016 and indicate absolute number of people diagnosed with cancer. Data for P.A. Trento and
+#' P.A. Bolzano are missing (but we have Trentino Alto Adige)
+#' @docType data
+#' @source \href{http://www.registri-tumori.it/PDF/AIOM2016/I_numeri_del_cancro_2016.pdf}{Istituto Superiore Sanita'}
+#' @usage data(cancer_it)
+#' @return a tibble
+"cancer_it"
+
+#' smoking status
+#'
+#' Number of people age 14 years and over that self-refer as smoker, non smoker, or past smoker by region and type.
+#' Data were collected in 2018 and are absolute number of people.
+#' @docType data
+#' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
+#' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
+#' @usage data(smoking_it)
+#' @return a tibble
+"smoking_it"
+
+#' chronic conditions
+#'
+#' Number of people suffering of chronic conditions by region and type.
+#' Data were collected in 2018 and indicate absolute number of people.
+#' @docType data
+#' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
+#' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
+#' @usage data(chronic_it)
+#' @return a tibble
+"chronic_it"
+
+
+#' influenza vaccination coverage, general population, time series
+#'
+#' Influenza vaccination coverage in Italy in the \strong{general population} from 1999 to 2019. Data are percent of region population
+#' @docType data
[…]

### 2020-04-21 R/datasets_us.R

@@ -1,0 +1,85 @@
+#' influenza vaccination 65 or older
+#'
+#' Percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination. Collected in 2017.
+#' @docType data
+#' @source \href{https://data.cms.gov/mapping-medicare-disparities}{Data.CMS.gov}
+#' @details \href{https://www.cms.gov/About-CMS/Agency-Information/OMH/Downloads/Mapping-Technical-Documentation.pdf}{Center for Medicare and Medicaid Services} and NORC at the University of Chicago.
+#' @return  tibble wotj `fl_65` indicating the percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination
+#' @usage data(fl65_us)
+"fl65_us"
+#' hospital beds
+#'
+#' beds of each hospital by county.
+#' @docType data
+#' @source \href{https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18}{Homeland Infrastructure Foundation-Level Data}
+#' @return  a tibble
+#' @usage data(hospbeds_us)
+"hospbeds_us"
+
+
+
+#' household composition
+#'
+#' Several metrics regarding household composition from the American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions}
+#' @return  a tibble
+#' @usage data(acm_househ_us)
+"acm_househ_us"
+
+#' age and sex
+#'
+#' Sex and age composition of the county population from the American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @return  a tibble
+#' @usage data(age_sex_us)
+"age_sex_us"
+
+#' poverty
+#'
+#' Household living below the poverty level, divided by age and race and calculate as absolute value or percentage. American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey}
+#' @return  a tibble
[…]

### 2020-04-21 R/get_covid.R

@@ -1,0 +1,81 @@
+#' get COVID-19 updated cases
+#'
+#' extracts and translates time series form the git repository of the \href{https://github.com/pcm-dpc/COVID-19}{protezione civile}
+#'
+#' @return a tibble with following columns:
+#' \describe{
+#'      \item{\strong{date}}{date of data}
+#'      \item{\strong{state}}{state}
+#'      \item{\strong{region_code}}{region abbreviation}
+#'      \item{\strong{region}}{full name of region}
+#'      \item{\strong{lat}}{lat}
+#'      \item{\strong{long}}{long}
+#'      \item{\strong{COVID19_var}}{
+#'          \itemize{
+#'             \item cmr: case-mortality rate for that region and that date (deaths/total_cases * 100)
+#'             \item hospitalized_with_symptoms: number of people hospitalized with symptoms
+#'             \item intensive_care_unit: number of people in intensive care units
+#'             \item home_quarantine: number of people COVID-19 positive in home quarantine
+#'             \item deaths: number of deaths
+#'             \item total_cases: number of COVID-19 positive cases detected
+#'             \item tests: number of tests performed
+#'          }
+#'      }
+#'      \item{\strong{value}}{value relative to COVID19_var}
+#'
+#' }
+#' @importFrom rlang .data
+#' @importFrom magrittr %>%
+#' @import vroom
+#' @export
+get_covid <- function() {
+  dat <- vroom(
+    file = "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",
+    col_types = cols(
+      data = col_datetime(format = ""),
+      stato = col_character(),
+      codice_regione = col_double(),
+      denominazione_regione = col_character(),
+      lat = col_double(),
+      long = col_double(),
+      ricoverati_con_sintomi = col_double(),
+      terapia_intensiva = col_double(),
+      totale_ospedalizzati = col_double(),
+      isolamento_domiciliare = col_double(),
+      totale_positivi = col_double(),
+      variazione_totale_positivi = col_double(),
+      nuovi_positivi = col_double(),
+      dimessi_guariti = col_double(),
+      deceduti = col_double(),
+      totale_casi = col_double(),
+      tamponi = col_double(),
+      note_it = col_character(),
+      note_en = col_character()
+    )
+  ) %>%
[…]

### 2020-04-21 R/getit_functions.R

@@ -1,0 +1,179 @@
+#' get COVID-19 updated cases
+#'
+#' extracts and translates time series form the git repository of the \href{https://github.com/pcm-dpc/COVID-19}{protezione civile}
+#'
+#' @return a tibble with following columns:
+#' \describe{
+#'      \item{\strong{date}}{date of data}
+#'      \item{\strong{state}}{state}
+#'      \item{\strong{region_code}}{region abbreviation}
+#'      \item{\strong{region}}{full name of region}
+#'      \item{\strong{lat}}{lat}
+#'      \item{\strong{long}}{long}
+#'      \item{\strong{COVID19_var}}{
+#'          \itemize{
+#'             \item cmr: case-mortality rate for that region and that date (deaths/total_cases * 100)
+#'             \item hospitalized_with_symptoms: number of people hospitalized with symptoms
+#'             \item intensive_care_unit: number of people in intensive care units
+#'             \item home_quarantine: number of people COVID-19 positive in home quarantine
+#'             \item deaths: number of deaths
+#'             \item total_cases: number of COVID-19 positive cases detected
+#'             \item tests: number of tests performed
+#'          }
+#'      }
+#'      \item{\strong{value}}{value relative to COVID19_var}
+#'
+#' }
+#' @importFrom rlang .data
+#' @importFrom magrittr %>%
+#' @import vroom
+#' @export
+getit_covid <- function() {
+  dat <- vroom(
+    file = "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",
+    col_types = cols(
+      data = col_datetime(format = ""),
+      stato = col_character(),
+      codice_regione = col_double(),
+      denominazione_regione = col_character(),
+      lat = col_double(),
+      long = col_double(),
+      ricoverati_con_sintomi = col_double(),
+      terapia_intensiva = col_double(),
+      totale_ospedalizzati = col_double(),
+      isolamento_domiciliare = col_double(),
+      totale_positivi = col_double(),
+      variazione_totale_positivi = col_double(),
+      nuovi_positivi = col_double(),
+      dimessi_guariti = col_double(),
+      deceduti = col_double(),
+      totale_casi = col_double(),
+      tamponi = col_double(),
+      note_it = col_character(),
+      note_en = col_character()
+    )
+  ) %>%
[…]

### 2020-04-21 data-raw/import_raw.R

@@ -1,0 +1,819 @@
+library(vroom)
+library(tidyverse)
+library(janitor)
+library(reshape2)
+
+
+# United States #-----
+
+# xxxxxxxxxxxxxxxx-------------------
+# age and sex- age_sex_us ---------
+
+age_sex_all_us <- vroom("data-raw/us/american_comunity_survey_2018/age_sex/ACSST5Y2018.S0101_data_with_overlays_2020-04-14T063202.csv",
+  skip = 1,
+  col_select = c(`id`, `Geographic Area Name`, starts_with("Estimate"))
+) %>%
+  clean_names()
+
+
+to_select_age_sex <- readxl::read_xlsx("data-raw/us/american_comunity_survey_2018/age_sex/to_select_age_sex.xlsx")
+
+age_sex_us <- age_sex_all_us %>%
+  select(!!to_select_age_sex$old_names)
+
+
+names(age_sex_us) <- to_select_age_sex$new_names
+
+age_sex_us <- age_sex_us %>%
+  separate(id, into = c("us_id", "fips"), sep = 9) %>%
+  select(-us_id) %>%
+  mutate(fips = as.numeric(fips))
+
+
+# flu vax - fl65_us  ----------
+fl65_all_us <- vroom("data-raw/us/flu_us.csv",
+  col_types = cols(
+    year = col_double(),
+    geography = col_character(),
+    measure = col_character(),
+    adjustment = col_character(),
+    analysis = col_character(),
+    domain = col_character(),
+    condition = col_character(),
+    primary_sex = col_character(),
+    primary_age = col_character(),
+    primary_dual = col_character(),
+    fips = col_double(),
+    county = col_character(),
+    state = col_character(),
+    urban = col_character(),
+    primary_race = col_character(),
+    primary_denominator = col_character(),
+    analysis_value = col_double()
+  )
+)
+
+fl65_us <- fl65_all_us %>%
+  rename("imm65" = "analysis_value") %>%
+  dplyr::select(state, county, fips, imm65)
+
+# hospital beds - hospbeds_us  ----------
+hospbeds_all_us <- vroom("data-raw/us/hospital_beds.csv",
+  col_types = cols(
+    X = col_double(),
+    Y = col_double(),
+    OBJECTID = col_double(),
+    ID = col_character(),
+    NAME = col_character(),
+    ADDRESS = col_character(),
+    CITY = col_character(),
[…]

### 2020-04-21 data-raw/import_raw_us.R

@@ -1,269 +1,0 @@
-library(vroom)
-library(tidyverse)
-library(janitor)
-library(reshape2)
-
-
-# United States #-----
-
-# age and sex- age_sex_us ---------
-
-age_sex_all_us <- vroom("data-raw/us/american_comunity_survey_2018/age_sex/ACSST5Y2018.S0101_data_with_overlays_2020-04-14T063202.csv",
-                        skip = 1,
-                        col_select = c(`id`, `Geographic Area Name`, starts_with("Estimate"))
-) %>%
-  clean_names()
-
-
-to_select_age_sex <- readxl::read_xlsx("data-raw/us/american_comunity_survey_2018/age_sex/to_select_age_sex.xlsx")
-
-age_sex_us <- age_sex_all_us %>%
-  select(!!to_select_age_sex$old_names)
-
-
-names(age_sex_us) <- to_select_age_sex$new_names
-
-age_sex_us <- age_sex_us %>%
-  separate(id, into = c("us_id", "fips"), sep = 9) %>%
-  select(-us_id) %>%
-  mutate(fips = as.numeric(fips))
-
-
-# flu vax - fl65_us  ----------
-fl65_all_us <- vroom("data-raw/us/flu_us.csv",
-  col_types = cols(
-    year = col_double(),
-    geography = col_character(),
-    measure = col_character(),
-    adjustment = col_character(),
-    analysis = col_character(),
-    domain = col_character(),
-    condition = col_character(),
-    primary_sex = col_character(),
-    primary_age = col_character(),
-    primary_dual = col_character(),
-    fips = col_double(),
-    county = col_character(),
-    state = col_character(),
-    urban = col_character(),
-    primary_race = col_character(),
-    primary_denominator = col_character(),
-    analysis_value = col_double()
-  )
-)
-
-fl65_us <- fl65_all_us %>%
-  rename("imm65" = "analysis_value") %>%
-  dplyr::select(state, county, fips, imm65)
-
-# hospital beds - hospbeds_us  ----------
-hospbeds_all_us <- vroom("data-raw/us/hospital_beds.csv",
-  col_types = cols(
-    X = col_double(),
-    Y = col_double(),
-    OBJECTID = col_double(),
-    ID = col_character(),
-    NAME = col_character(),
-    ADDRESS = col_character(),
-    CITY = col_character(),
-    STATE = col_character(),
-    ZIP = col_character(),
-    ZIP4 = col_character(),
[…]

### 2020-04-21 tests/testthat/test-get.R

@@ -1,59 +1,0 @@
-library(dplyr)
-
-test_that("get_all executes", {
-  expect_error(getus_all(), NA)
-})
-
-test_that("get_covid executes", {
-  expect_error(getus_covid(), NA)
-})
-
-y <- getus_all()
-
-x <- getus_covid()
-
-test_that("get_all and get_covid have same fips", {
-  expect_equal(length(unique(y$fips)), length(unique(x$fips)))
-})
-
-
-###################################
-
-test_that("get_dex executes", {
-  expect_error(getus_dex(), NA)
-})
-
-##################################
-
-
-expected_states <- structure(list(state = c(
-  "Alaska", "Alabama", "Arizona", "Arkansas",
-  "California", "Colorado", "Connecticut", "District of Columbia",
-  "Delaware", "Florida", "Georgia", "Hawaii", "Iowa", "Idaho",
-  "Illinois", "Indiana", "Kansas", "Kentucky", "Louisiana", "Massachusetts",
-  "Maryland", "Maine", "Michigan", "Minnesota", "Missouri", "Mississippi",
-  "Montana", "North Carolina", "North Dakota", "Nebraska", "New Hampshire",
-  "New Jersey", "New Mexico", "Nevada", "New York", "Ohio", "Oklahoma",
-  "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
-  "Tennessee", "Texas", "Utah", "Virginia", "Vermont", "Washington",
-  "Wisconsin", "West Virginia", "Wyomin"
-), abbr = c(
-  "AK", "AL",
-  "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI", "IA",
-  "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN",
-  "MO", "MS", "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY",
-  "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VA",
-  "VT", "WA", "WI", "WV", "WY"
-)), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA, -51L))
-
-test_that("get_test executes", {
-  expect_error(getus_tests(), NA)
-})
-
-tests <- getus_tests() %>%
-  select(state, abbr) %>%
-  distinct()
-
-test_that("getus_tests states abbr ", {
-  expect_equal(tests, expected_states)
-})

### 2020-04-21 tests/testthat/test-getit.R

@@ -1,0 +1,61 @@
+library(stringr)
+
+test_that("get_all executes", {
+  expect_error(getit_all(), NA)
+})
+
+y <- getit_all()
+
+test_that("get_all returns regions_21", {
+  expect_equal(length(unique(y$region)), 21)
+})
+
+
+test_that("sum smoking is < 100 ", {
+  expect_equal(
+    sum(!unique(
+      apply(
+        y[, names(y)[str_detect(names(y), "smoking")]],
+        1, sum
+      )
+    ) <= 100),
+    0
+  )
+})
+
+
+test_that("getit_covid executes", {
+  expect_error(get_covid(), NA)
+})
+
+x <- getit_covid()
+
+# we take only the last year
+house_it <- house_it %>%
+  dplyr::filter(.data$year == 2018) %>%
+  dplyr::select(.data$region, .data$phouse) %>%
+  dplyr::rename("p_house" = "phouse")
+
+# check the datasets that we are merging-----
+t_data <- lapply(list(
+  x,
+  cancer_it,
+  chronic_it,
+  dem_65bin_fm,
+  bweight_it,
+  firstaid_it,
+  hospbed_it,
+  house_it,
+  netinc_it,
+  pm2.5_it,
+  regions_area,
+  smoking_it
+), function(x) {
+  unique(x$region) %in% unique(dem_65bin_fm$region)
+})
+
+
+test_that("all datasets and function have same regions", {
+  expect_equal(sum(unlist(lapply(t_data, sum))), 252)
+})
+

### 2020-04-21 tests/testthat/test-getus.R

@@ -1,0 +1,59 @@
+library(dplyr)
+
+test_that("get_all executes", {
+  expect_error(getus_all(), NA)
+})
+
+test_that("get_covid executes", {
+  expect_error(getus_covid(), NA)
+})
+
+y <- getus_all()
+
+x <- getus_covid()
+
+test_that("get_all and get_covid have same fips", {
+  expect_equal(length(unique(y$fips)), length(unique(x$fips)))
+})
+
+
+###################################
+
+test_that("get_dex executes", {
+  expect_error(getus_dex(), NA)
+})
+
+##################################
+
+
+expected_states <- structure(list(state = c(
+  "Alaska", "Alabama", "Arizona", "Arkansas",
+  "California", "Colorado", "Connecticut", "District of Columbia",
+  "Delaware", "Florida", "Georgia", "Hawaii", "Iowa", "Idaho",
+  "Illinois", "Indiana", "Kansas", "Kentucky", "Louisiana", "Massachusetts",
+  "Maryland", "Maine", "Michigan", "Minnesota", "Missouri", "Mississippi",
+  "Montana", "North Carolina", "North Dakota", "Nebraska", "New Hampshire",
+  "New Jersey", "New Mexico", "Nevada", "New York", "Ohio", "Oklahoma",
+  "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
+  "Tennessee", "Texas", "Utah", "Virginia", "Vermont", "Washington",
+  "Wisconsin", "West Virginia", "Wyomin"
+), abbr = c(
+  "AK", "AL",
+  "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI", "IA",
+  "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN",
+  "MO", "MS", "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY",
+  "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VA",
+  "VT", "WA", "WI", "WV", "WY"
+)), class = c("tbl_df", "tbl", "data.frame"), row.names = c(NA, -51L))
+
+test_that("getus_test executes", {
+  expect_error(getus_tests(), NA)
+})
+
+tests <- getus_tests() %>%
+  select(state, abbr) %>%
+  distinct()
+
+test_that("getus_tests states abbr ", {
+  expect_equal(tests, expected_states)
+})

### 2020-04-22 README.Rmd
changed shape of gitit_covid, added check.url, started readme

@@
 <!-- badges: start -->
+[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
 <!-- badges: end -->
 
-The goal of covid19census is to ...
+The package `covid19census` provides functions to extract COVID-19 cases, deaths, hospitalizations and test of U.S.
+and Italy at the county and regional level, respectively, and then combine them with other population metric (age, sex, prevalence of chronic conditions, income indexes, access to health services and many others.)
+
+## Installation
+
+Download the package in a local folder and the run the following code.
+
+```{r intallation, eval=FALSE, include=TRUE}
+devtools::install_local("path_to_package")
+```
+
+Alternatively, you can install it directly from github.
+
+```{r intallation2, eval=FALSE, include=TRUE}
+# the repo at the moment private so you need to be authorized 
+# don't share the token
+library(devtools)
+
+devtools::install_github("c1au6i0/rstract", auth_token = "355580fb57d57b58228b4617c14c3f8234741715")
+```
+
+## Usage
+
+A dataset that includes COVID-19 data and **all the other demographic variables** can be obtained executing `getus_all` (get-U.S-all) and `getit_all` (get-Italy-all)
+
+```{r all, message=FALSE}
+all_us <- getus_all()
+
+all_it <- getus_all()
+```
+
+## U.S. Datasets and sources
+
+The function `get  all` executes
+
+`get  covid` and joins the resulting `tibble` with datasets from the [Homeland Infrastructure Foundation](https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18), the [Census](https://data.census.gov/cedsci/table?q=United%20States), [Mapping Medicare Disparities](https://data.cms.gov/mapping-medicare-disparities), and [activity indexes](https://github.com/COVIDExposureIndices/COVIDExposureIndices) calculated by  Victor Couture, Jonathan Dingel, Allison Green, Jessie Handbury, and Kevin Williams based on smartphone movement data provided by PlaceIQ.
+
+In particular, this is a list of datasets included:
+
+* `act_ind`: county-level device exposure index (DEX), an index of activity.
+* `acm_househ`: several metric regarding household composition (2018).
+* `age_sex`: age and sex distribution (2018).
+* `fl65`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
+* `hospbeds`: total hospital beds in each county (2019).
+* `mmd`: data of 2017 regarding the  prevalence found in medicare beneficiaries of
+
+    * many medical and chronic conditions (Alzheimer, chronic kidney, obesity, depression, obstructive pulmonary, disease, arthritis, diabetes, osteoporosis, asthma, atrial  fibrillation, ischemic hearth, myocardial infarction, hypertension).
+    * several type of cancer.
+    * emergency,  medical admissions and annual visits.
+    * pneumoccocal vaccine.
+    * tabacco use.
+
+*Note that* info for some counties are missing in some datasets. For example, `hospbeds` contains info on 2545 counties, `fl65` has 3224 counties, whereas datasets from the Census have 3220 counties.
+
+## Datasets Italy
+
+The function `get_all` executes `get_covid` and aggregates the resulting `tibble` with some of the datasets below and, normalizes some of the variables:
+
+* `act_it`:  Change in  retail and recreation activity as reported by google [not joined with `get_all()`  but accessible].
+* `bweight_it`: Body mass index in regions of Italy, in the general population.
+* `cancer_it`: Number of cancer patients in each region by type.
+* `chronic_it`: Number of people suffering of chronic conditions by region and type.
+* `dem_it`: Number of people by region, sex, age.
+* `firstaid_it`: Number of people using first aid or medical guard (3 months). 
+* `fl_it`: Influenza vaccination coverage in Italy for 2009-2019 season for general population and aged 65 and more.
+* `fl_it_2019`: Influenza vaccination coverage in Italy for 2018-2019 season for general population and 65 and more.
+* `hospbed_it`: Inpatient hospital beds per 1000 people.
+* `house_it`:  household crowding index from 2014 to 2018 in each region.
+* `netinc_it`:  median net annual households income (including imputed rents, in euros).
+* `pm2.5_it`:  emission of particulate 2.5 in tons per region.
+* `regions_area`: Area in square meters of each region. Used to calculate density per region.
+* `smoking_it`: Number of people aged 14 years and over that self-refered as smoker, non smoker, or past smoker by region.
+
+## Raw Data and other scripts
+
+Raw data and the code used to import it can be found in the [data-raw](https://github.com/c1au6i0/convid19census/tree/master/data-raw) folder.

### 2020-04-22 R/datasets_us.R

@@ -7,16 +7,15 @@
 #' @return  tibble wotj `fl_65` indicating the percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination
 #' @usage data(fl65_us)
 "fl65_us"
+
 #' hospital beds
 #'
-#' beds of each hospital by county.
+#' beds of each hospital by county (2019).
 #' @docType data
 #' @source \href{https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18}{Homeland Infrastructure Foundation-Level Data}
 #' @return  a tibble
 #' @usage data(hospbeds_us)
 "hospbeds_us"
-
-
 
 #' household composition
 #'

### 2020-04-22 R/get_covid.R

@@ -1,81 +1,0 @@
-#' get COVID-19 updated cases
-#'
-#' extracts and translates time series form the git repository of the \href{https://github.com/pcm-dpc/COVID-19}{protezione civile}
-#'
-#' @return a tibble with following columns:
-#' \describe{
-#'      \item{\strong{date}}{date of data}
-#'      \item{\strong{state}}{state}
-#'      \item{\strong{region_code}}{region abbreviation}
-#'      \item{\strong{region}}{full name of region}
-#'      \item{\strong{lat}}{lat}
-#'      \item{\strong{long}}{long}
-#'      \item{\strong{COVID19_var}}{
-#'          \itemize{
-#'             \item cmr: case-mortality rate for that region and that date (deaths/total_cases * 100)
-#'             \item hospitalized_with_symptoms: number of people hospitalized with symptoms
-#'             \item intensive_care_unit: number of people in intensive care units
-#'             \item home_quarantine: number of people COVID-19 positive in home quarantine
-#'             \item deaths: number of deaths
-#'             \item total_cases: number of COVID-19 positive cases detected
-#'             \item tests: number of tests performed
-#'          }
-#'      }
-#'      \item{\strong{value}}{value relative to COVID19_var}
-#'
-#' }
-#' @importFrom rlang .data
-#' @importFrom magrittr %>%
-#' @import vroom
-#' @export
-get_covid <- function() {
-  dat <- vroom(
-    file = "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv",
-    col_types = cols(
-      data = col_datetime(format = ""),
-      stato = col_character(),
-      codice_regione = col_double(),
-      denominazione_regione = col_character(),
-      lat = col_double(),
-      long = col_double(),
-      ricoverati_con_sintomi = col_double(),
-      terapia_intensiva = col_double(),
-      totale_ospedalizzati = col_double(),
-      isolamento_domiciliare = col_double(),
-      totale_positivi = col_double(),
-      variazione_totale_positivi = col_double(),
-      nuovi_positivi = col_double(),
-      dimessi_guariti = col_double(),
-      deceduti = col_double(),
-      totale_casi = col_double(),
-      tamponi = col_double(),
-      note_it = col_character(),
-      note_en = col_character()
-    )
-  ) %>%
[…]

### 2020-04-22 R/getit_functions.R

@@ -4,76 +4,109 @@
 #'
 #' @return a tibble with following columns:
 #' \describe{
-#'      \item{\strong{date}}{date of data}
-#'      \item{\strong{state}}{state}
-#'      \item{\strong{region_code}}{region abbreviation}
-#'      \item{\strong{region}}{full name of region}
-#'      \item{\strong{lat}}{lat}
-#'      \item{\strong{long}}{long}
-#'      \item{\strong{COVID19_var}}{
-#'          \itemize{
-#'             \item cmr: case-mortality rate for that region and that date (deaths/total_cases * 100)
-#'             \item hospitalized_with_symptoms: number of people hospitalized with symptoms
-#'             \item intensive_care_unit: number of people in intensive care units
-#'             \item home_quarantine: number of people COVID-19 positive in home quarantine
-#'             \item deaths: number of deaths
-#'             \item total_cases: number of COVID-19 positive cases detected
-#'             \item tests: number of tests performed
-#'          }
-#'      }
-#'      \item{\strong{value}}{value relative to COVID19_var}
-#'
+#'      \item{date}{date of data}
+#'      \item{state}{state}
+#'      \item{region_code}{region abbreviation}
+#'      \item{region}{full name of region}
+#'      \item{lat}{lat}
+#'      \item{long}{long}
+#'      \item {cmr}{case-mortality rate for that region and that date (deaths/total_cases * 100)}
+#'      \item {hospitalized_with_symptoms}{number of people hospitalized with symptoms}
+#'      \item {intensive_care_unit}{number of people in intensive care units}
+#'      \item {home_quarantine}{number of people COVID-19 positive in home quarantine}
+#'      \item {deaths}{number of deaths}
+#'      \item {total_cases}{number of COVID-19 positive cases detected}
+#'      \item {tests}{number of tests performed}
+#'      \item{value}{value relative to COVID19_var}
 #' }
 #' @importFrom rlang .data
 #' @importFrom magrittr %>%
 #' @import vroom
 #' @export
 getit_covid <- function() {
+
+  url_data <-  "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv"
+
+  if(RCurl::url.exists(url_data) == FALSE){
+    stop("Something wrong with the repository or your internet connection!")
+  }
+
   dat <- vroom(
[…]

### 2020-04-22 R/getus_functions.R

@@ -11,7 +11,15 @@
 #' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
 #' @export
 getus_covid <- function() {
-  dat <- vroom("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv",
+
+  url_data <-  "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"
+
+  if(RCurl::url.exists(url_data) == FALSE){
+    stop("Something wrong with the repository or your internet connection!")
+  }
+
+
+  dat <- vroom(url_data,
     col_types = cols(
       date = col_date(format = ""),
       county = col_character(),
@@ -46,7 +54,15 @@ getus_covid <- function() {
 #' @import vroom
 #' @export
 getus_dex <- function() {
-  dat <- vroom("https://raw.githubusercontent.com/COVIDExposureIndices/COVIDExposureIndices/master/dex_data/county_dex.csv",
+
+  url_data <-  "https://raw.githubusercontent.com/COVIDExposureIndices/COVIDExposureIndices/master/dex_data/county_dex.csv"
+
+  if(RCurl::url.exists(url_data) == FALSE){
+    stop("Something wrong with the repository or your internet connection!")
+  }
+
+
+  dat <- vroom(url_data,
     skip = 1,
     col_names = c(
       "fips",
@@ -107,7 +123,14 @@ getus_dex <- function() {
 #'  @import vroom
 #'  @export
 getus_tests <- function() {
-  dat <- vroom::vroom("https://covidtracking.com/api/states/daily.csv",
+  url_data <-  "https://covidtracking.com/api/states/daily.csv"
+
+  if(RCurl::url.exists(url_data) == FALSE){
+    stop("Something wrong with the repository or your internet connection!")
+  }
+
+
+  dat <- vroom::vroom(url_data,
     col_types = cols(
       date = col_date(format = "%Y%m%d"),
       state = col_character(),

### 2020-04-22 README.Rmd
fixed errors docu and testhat

@@
 
-devtools::install_github("c1au6i0/rstract", auth_token = "355580fb57d57b58228b4617c14c3f8234741715")
+devtools::install_github("c1au6i0/covid19census", auth_token = "355580fb57d57b58228b4617c14c3f8234741715")
 ```
@@
 ```{r all, message=FALSE}
+library(covid19census)
+
+# to retrive u.s data
 all_us <- getus_all()
 
-all_it <- getus_all()
+# to retrive italy data
+all_it <- getit_all()
 ```
 
-## U.S. Datasets and sources
+## U.S. Datasets and Sources
 
@@
 
-## Datasets Italy
+## Italy: Datasets and Sources
 

### 2020-04-22 R/getit_functions.R

@@ -2,23 +2,31 @@
 #'
 #' extracts and translates time series form the git repository of the \href{https://github.com/pcm-dpc/COVID-19}{protezione civile}
 #'
-#' @return a tibble with following columns:
+#' @return a tibble with following 19 variables:
 #' \describe{
-#'      \item{date}{date of data}
+#'      \item{date}{in `ISO 8601` format}
 #'      \item{state}{state}
 #'      \item{region_code}{region abbreviation}
 #'      \item{region}{full name of region}
 #'      \item{lat}{lat}
 #'      \item{long}{long}
-#'      \item {cmr}{case-mortality rate for that region and that date (deaths/total_cases * 100)}
-#'      \item {hospitalized_with_symptoms}{number of people hospitalized with symptoms}
-#'      \item {intensive_care_unit}{number of people in intensive care units}
-#'      \item {home_quarantine}{number of people COVID-19 positive in home quarantine}
-#'      \item {deaths}{number of deaths}
-#'      \item {total_cases}{number of COVID-19 positive cases detected}
-#'      \item {tests}{number of tests performed}
-#'      \item{value}{value relative to COVID19_var}
+#'      \item{cmr}{case-mortality rate for that region and that date (deaths/total_cases * 100)}
+#'      \item{total_cases}{number of COVID-19 positive cases detected}
+#'      \item{deaths}{number of deaths}
+#'      \item{tests}{number of tests performed}
+#'      \item{hospitalized_with_symptoms}{number of people hospitalized with symptoms, that day}
+#'      \item{intensive_care_unit}{number of people in intensive care units, that day}
+#'      \item{total_hospitalized}{hospitalized_with_symptoms + intensive_care_unit}
+#'      \item{home_quarantine}{number of people COVID-19 positive in home quarantine, that day}
+#'      \item{total_positives}{total currently positives: hospitalized_with_symptoms + intensive_care_unit + home_quarantine}
+#'      \item{change_positives}{change in the number of positive cases: total_positives that day - total_positives preceding day}
+#'      \item{new_positives}{number of new positive cases: total_cases that day - total_cases preceding day}
+#'      \item{recovered_released}{recovered - released from hospital}
+#'      \item{people_tested}{number of people tested}
 #' }
+#' @details caveats and problems related the calculation by the Protezione Civile of some variables  were rised by
[…]

### 2020-04-22 tests/testthat/test-getit.R

@@ -1,31 +1,8 @@
 library(stringr)
 
-test_that("get_all executes", {
-  expect_error(getit_all(), NA)
-})
-
-y <- getit_all()
-
-test_that("get_all returns regions_21", {
-  expect_equal(length(unique(y$region)), 21)
-})
-
-
-test_that("sum smoking is < 100 ", {
-  expect_equal(
-    sum(!unique(
-      apply(
-        y[, names(y)[str_detect(names(y), "smoking")]],
-        1, sum
-      )
-    ) <= 100),
-    0
-  )
-})
-
-
+# getit_covid --------------------
 test_that("getit_covid executes", {
-  expect_error(get_covid(), NA)
+  expect_error(getit_covid(), NA)
 })
 
 x <- getit_covid()
@@ -57,5 +34,29 @@ t_data <- lapply(list(
 
 test_that("all datasets and function have same regions", {
   expect_equal(sum(unlist(lapply(t_data, sum))), 252)
+})
+
+
+test_that("get_all executes", {
+  expect_error(getit_all(), NA)
+})
+
+y <- getit_all()
+
+test_that("get_all returns regions_21", {
+  expect_equal(length(unique(y$region)), 21)
+})
+
+
+test_that("sum smoking is < 100 ", {
+  expect_equal(
+    sum(!unique(
+      apply(
+        y[, names(y)[str_detect(names(y), "smoking")]],
+        1, sum
+      )
+    ) <= 100),
+    0
+  )
 })
 

### 2020-04-22 README.Rmd
vignette

@@
 
-A dataset that includes COVID-19 data and **all the other demographic variables** can be obtained executing `getus_all` (get-U.S-all) and `getit_all` (get-Italy-all)
+A dataset that includes COVID-19 data and **all the other demographic variables** can be obtained executing `getus_all` (get-U.S-all) to get U.S data, or `getit_all` (get-Italy-all) to get Italy data.
 
@@
 
-# to retrive u.s data
+# to retrieve u.s data
 all_us <- getus_all()
 
-# to retrive italy data
+# to retrieve italy data
 all_it <- getit_all()
@@
 
+Details on the other functions and the datasets can be found in the Vignette of the package.
+
 ## U.S. Datasets and Sources
 
-The function `get  all` executes
+Data regarding COVID-19 number of cases and deaths (at the county level) were obtained from the [New York Time git repository](https://github.com/nytimes/covid-19-data), whereas number of tests and hospitalizations from the [the Covid Tracking Project].
 
-`get  covid` and joins the resulting `tibble` with datasets from the [Homeland Infrastructure Foundation](https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18), the [Census](https://data.census.gov/cedsci/table?q=United%20States), [Mapping Medicare Disparities](https://data.cms.gov/mapping-medicare-disparities), and [activity indexes](https://github.com/COVIDExposureIndices/COVIDExposureIndices) calculated by  Victor Couture, Jonathan Dingel, Allison Green, Jessie Handbury, and Kevin Williams based on smartphone movement data provided by PlaceIQ.
+The other datsets were obtained from the [Homeland Infrastructure Foundation](https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18), the [Census](https://data.census.gov/cedsci/table?q=United%20States), [Mapping Medicare Disparities](https://data.cms.gov/mapping-medicare-disparities), and [activity indexes](https://github.com/COVIDExposureIndices/COVIDExposureIndices) calculated by  Victor Couture, Jonathan Dingel, Allison Green, Jessie Handbury, and Kevin Williams based on smartphone movement data provided by `PlaceIQ`.
 
-In particular, this is a list of datasets included:
+This is a list of U.S. datasets that can be accessed directly or in a `tibble` containing also COVID-19 metrics:
 
@@
 * `mmd`: data of 2017 regarding the  prevalence found in medicare beneficiaries of 
-
     * many medical and chronic conditions (Alzheimer, chronic kidney, obesity, depression, obstructive pulmonary, disease, arthritis, diabetes, osteoporosis, asthma, atrial  fibrillation, ischemic hearth, myocardial infarction, hypertension).
@@
     * pneumoccocal vaccine.
-    * tabacco use.
+    * tobacco use.
 
-*Note that* info for some counties are missing in some datasets. For example, `hospbeds` contains info on 2545 counties, `fl65` has 3224 counties, whereas datasets from the Census have 3220 counties.
+*Note that* info for some counties are missing in some datasets. For example, `hospbeds` contains info on 2545 counties, `fl65` has 3224 counties, whereas datasets from the Census have 3220.
 
@@
 
-The function `get_all` executes `get_covid` and aggregates the resulting `tibble` with some of the datasets below and, normalizes some of the variables:
+Data regarding COVID-19 deaths, confirmed cases, test and hospitalizations are extracted from the [Protezione Civile repository](https://github.com/pcm-dpc/COVID-19)
 
-* `act_it`:  Change in  retail and recreation activity as reported by google [not joined with `get_all()`  but accessible].
+Demographic metrics and other stats regarding medical conditions and vaccinations were obtained from the [ISTAT](http://dati.istat.it/?lang=en), [Ministero della Salute](http://www.salute.gov.it/portale/home.html).
+Data regarding `pm2.5` were collected by [Istituto Superiore Per La protexione Ambientale](https://annuario.isprambiente.it/pon/basic/14)
+
+This is a list of Italy datasets that can be accessed directly or in a `tibble` containing also COVID-19 metrics:
+
 * `bweight_it`: Body mass index in regions of Italy, in the general population.
@@
 * `regions_area`: Area in square meters of each region. Used to calculate density per region.
-* `smoking_it`: Number of people aged 14 years and over that self-refered as smoker, non smoker, or past smoker by region.
+* `smoking_it`: Number of people aged 14 years and over that self-referred as smoker, non smoker, or past smoker by region.
 
-## Raw Data and other scripts
+## Raw Data and other Scripts
 

### 2020-04-22 R/datasets_us.R

@@ -33,8 +33,8 @@
 #' @docType data
 #' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
 #' @return  a tibble
-#' @usage data(age_sex_us)
-"age_sex_us"
+#' @usage data(dem_us)
+"dem_us"
 
 #' poverty
 #'

### 2020-04-22 data-raw/import_raw.R

@@ -756,35 +756,35 @@ smoking_it <- vroom("data-raw/it/smoking.csv",
 # usethis -----
 
 # external data
-# usethis::use_data(
-#   # US data
-#   acm_househ_us,
-#   age_sex_us,
-#   fl65_us,
-#   hospbeds_us,
-#   mmd_us,
-#   poverty_us,
-#   # state_abbr,
-#   # italy
-#   bweight_it,
-#   cancer_it,
-#   chronic_it,
-#   # dem_65bin_fm,
-#   dem_it,
-#   firstaid_it,
-#   fl_it,
-#   fl65_it,
-#   hospbed_it,
-#   house_it,
-#   netinc_it,
-#   pm2.5_it,
-#   regions_area,
-#   smoking_it,
-#
-#
-#   # internal = TRUE,
-#   overwrite = TRUE
-# )
+
+dem_us <-   age_sex_us
+usethis::use_data(
+  # US data
+  acm_househ_us,
+  dem_us,
+  fl65_us,
+  hospbeds_us,
+  mmd_us,
+  poverty_us,
+  # state_abbr,
+  # italy
+  bweight_it,
+  cancer_it,
+  chronic_it,
+  # dem_65bin_fm,
+  dem_it,
+  firstaid_it,
+  fl_it,
+  fl65_it,
+  hospbed_it,
+  house_it,
+  netinc_it,
+  pm2.5_it,
+  regions_area,
+  smoking_it,
+  # internal = TRUE,
+  overwrite = TRUE
+)
 #
 # # internal data
 # usethis::use_data(

### 2020-04-23 inst/img/make_hex.R
hexstiker readme

@@ -1,0 +1,56 @@
+library(devtools)
+
+# devtools::install_github("c1au6i0/covid19census")
+library(kableExtra)
+library(covid19census)
+library(dplyr)
+library(showtext)
+library(hexSticker)
+library(ggplot2)
+library(png)
+library(grid)
+
+tab2 <- getus_all()
+
+tab2 <- ff[1:50, 1:50]
+
+
+names(tab2) <- paste0("x", 1:50)
+
+
+tab2 %>%
+  kable(escape = F, align = "c") %>%
+  column_spec(1:50, bold = T, border_right = T) %>%
+  kable_styling(
+    latex_options = c("striped", "scale_down"),
+    full_width = FALSE,
+    font_size = 5
+  )
+
+
+
+tab1 <- getit_all()
+glimpse(tab1)
+
+imgurl <- readPNG("inst/img/image_logo.png")
+imgurl <- rasterGrob(imgurl, interpolate = TRUE)
+
+font_add_google(name = "Quantico", "gira")
+showtext_auto()
+
+sticker(
+  imgurl,
+  package = "covid19census",
+  p_size = 5,
+  p_y = 1.5,
+  s_x = 1.1,
+  s_y = 0.7,
+  s_width = 2.5,
+  s_height = 2.5,
+  p_color = "#b83335",
+  p_family = "gira",
+  filename = "inst/img/hexsticker2.png",
+  h_color = "#b83335",
+  h_fill = "white"
+)
+

### 2020-04-23 README.Rmd
updated token
removed token
hexstiker readme
fixed readme typo
vignette link test

@@
 
-# covid19census
+# covid19census  <img src='inst/img/hexsticker.png' align="right" height="139" />
 
@@
 
+### Overview
+
 The package `covid19census` provides functions to extract COVID-19 cases, deaths, hospitalizations and test of U.S.
@@
 
-## Installation
+### Installation
 
@@
 
-devtools::install_github("c1au6i0/covid19census", auth_token = "355580fb57d57b58228b4617c14c3f8234741715")
+devtools::install_github("c1au6i0/covid19census")
 ```
 
-## Usage
+### Usage
 
@@
 
-Details on the other functions and the datasets can be found in the Vignette of the package.
+Details on the other functions and the datasets can be found in the [vignette](vignette_how_to.html) of the package.
 
-## U.S. Datasets and Sources
+### U.S. Datasets and Sources
 
@@
 
-## Italy: Datasets and Sources
+### Italy: Datasets and Sources
 
@@
 
-## Raw Data and other Scripts
+### Raw Data and other Scripts
 

### 2020-04-23 README.Rmd
vignette test2
readme links

@@
 
-Details on the other functions and the datasets can be found in the [vignette](vignette_how_to.html) of the package.
+Details on the other functions and the datasets can be found in the vignette of the package `vignette("v_covid19census")`.
 
@@
 Raw data and the code used to import it can be found in the [data-raw](https://github.com/c1au6i0/convid19census/tree/master/data-raw) folder.
+
+### Terms of Use
+
+Refer to the [License](https://github.com/c1au6i0/covid19census/blob/master/LICENSE.md) for details on the terms of use and copyrights.

### 2020-04-23 README.Rmd
readme fsize
typos

@@
 
-### Overview
+## Overview
 
 The package `covid19census` provides functions to extract COVID-19 cases, deaths, hospitalizations and test of U.S.
-and Italy at the county and regional level, respectively, and then combine them with other population metric (age, sex, prevalence of chronic conditions, income indexes, access to health services and many others.)
+and Italy at the county and regional level, respectively, and then combine them with other population metrics (age, sex, prevalence of chronic conditions, income indexes, access to health services and many others.)
 
-### Installation
+## Installation
 
@@
 
-### Usage
+## Usage
 
@@
 
-### U.S. Datasets and Sources
+## U.S. Datasets and Sources
 
@@
 
-### Italy: Datasets and Sources
+## Italy: Datasets and Sources
 
@@
 
-### Raw Data and other Scripts
+## Raw Data and other Scripts
 
@@
 
-### Terms of Use
+## Terms of Use
 

### 2020-04-23 inst/img/make_hex.R
change stiker

@@ -45,8 +45,8 @@ sticker(
   p_y = 1.5,
   s_x = 1.1,
   s_y = 0.7,
-  s_width = 2.5,
-  s_height = 2.5,
+  s_width = 2.3,
+  s_height = 2.3,
   p_color = "#b83335",
   p_family = "gira",
   filename = "inst/img/hexsticker2.png",

### 2020-04-23 README.Rmd
fixed typos

@@
 The package `covid19census` provides functions to extract COVID-19 cases, deaths, hospitalizations and test of U.S.
-and Italy at the county and regional level, respectively, and then combine them with other population metrics (age, sex, prevalence of chronic conditions, income indexes, access to health services and many others.)
+and Italy at the county and regional level, respectively, and then combine them with other population metrics (age, sex, prevalence of chronic conditions, income indexes, access to health services and many others).
 
@@
 
-Download the package in a local folder and the run the following code.
+Download the package in a local folder and then run the following code.
 

### 2020-04-23 R/getus_functions.R
fixed namespace getus_test

@@ -98,7 +98,7 @@ getus_dex <- function() {
 #' was used verbatim for the description below
 #' \describe{
 #'     \item{date}{in `ISO 8601` format}
-#'     \item{state}{state name}]
+#'     \item{state}{state name}
 #'     \item{abbr}{abbreviation}
 #'     \item{positive}{total cumulative positive test results}
 #'     \item{negative}{total cumulative negative test results}
@@ -108,7 +108,7 @@ getus_dex <- function() {
 #'     \item{icu_curr}{current people in ICU}
 #'     \item{icu_cumul}{cumulative people in ICU}
 #'     \item{ventilator_curr}{current people using ventilator}
-#'     \item{ventilator_cumul}{cumulativepeople using ventilator}
+#'     \item{ventilator_cumul}{cumulative people using ventilator}
 #'     \item{recovered}{total people recoverd}
 #'     \item{hash}{unique ID changed every time the data updates}
 #'     \item{date_checked}{date of the time we last visited their website}
@@ -120,8 +120,8 @@ getus_dex <- function() {
 #'     \item{total_test_increase}{increase from the day before}
 #' }
 #' Other details regarding the score system used are reported in the \href{https://covidtracking.com/about-data}{maintainers webpage}
-#'  @import vroom
-#'  @export
+#' @import vroom
+#' @export
 getus_tests <- function() {
   url_data <-  "https://covidtracking.com/api/states/daily.csv"
 

### 2020-04-23 inst/some_checks.Rmd
fixed some docu + article

---
title: "Checks on testing"
output: html_notebook
---

```{r}
library(covid19census)
library(dplyr)
library(tidyr)
library(ggplot2)
library(broom)
library(psych)
library(stringr)
library(DataExplorer)
library(plotly)
theme_set(theme_bw())
```

```{r}
t_us <-  covid19census::getus_tests()
```

```{r warning=FALSE}
names(t_us)
```

```{r, warning=FALSE, fig.height=5, fig.width=5}
t_us %>% 
  select(date, state, positive, negative, pending) %>% 
  pivot_longer(cols = c("positive", "negative", "pending"), names_to = "det_tests") %>%

  ggplot(aes(date, log2(value), colour = det_tests)) +
    geom_line() +
    facet_wrap(vars(state)) +
    scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
    theme(legend.position = "top",
        legend.title = element_blank())

```

```{r warning=FALSE, fig.height=5, fig.width=5}
t_us %>%
  select(date,
         state,
         hospitalized_curr,
         hospitalized_cumul
         # icu_curr,
         # icu_cumul

         ) %>%
  pivot_longer(
    cols = c(
      "hospitalized_curr",
      "hospitalized_cumul"
      # "icu_curr",
      # "icu_cumul"
    ),
    names_to = "det_tests"
  ) %>%

  ggplot(aes(date, log2(value), colour = det_tests)) +
  geom_line() +
  facet_wrap(vars(state)) +
  scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
  theme(legend.position = "top",
        legend.title = element_blank())

```
Same states have hospitalized current but not hospitalized cumulative and vicevera ????

** Possible different meaning of variables depending on the state  **

```{r warning=FALSE, fig.height=5, fig.width=5}
t_us %>%
  select(date,
         state,
         # hospitalized_curr,
         # hospitalized_cumul
         # icu_curr,
         # icu_cumul
         "ventilator_curr",
         "ventilator_cumul"

         ) %>%
  pivot_longer(
    cols = c(
      # "hospitalized_curr",
      # "hospitalized_cumul"
        "ventilator_curr",
         "ventilator_cumul"
    ),
    names_to = "det_tests"
  ) %>%

  ggplot(aes(date, log2(value), colour = det_tests)) +
  geom_line() +
  facet_wrap(vars(state)) +
  scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
  theme(legend.position = "top",
        legend.title = element_blank())

```
```{r warning=FALSE, fig.height=5, fig.width=5}
t_us %>%
  select(date,
         state,
         # hospitalized_curr,
         # hospitalized_cumul
         icu_curr,
         icu_cumul

         ) %>%
  pivot_longer(
    cols = c(
      # "hospitalized_curr",
      # "hospitalized_cumul"
      "icu_curr",
      "icu_cumul"
    ),
    names_to = "det_tests"
  ) %>%

  ggplot(aes(date, log2(value), colour = det_tests)) +
  geom_line() +
  facet_wrap(vars(state)) +
  scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
  theme(legend.position = "top",
        legend.title = element_blank())

```

Lets see some totals
```{r warning=FALSE, fig.height=10, fig.width=6}
ggplotly(
t_us %>% 
  filter(date == max(date)) %>% 
  select(-hash, -date_checked, -fips) %>%
  pivot_longer(cols = positive:total_test_increase, names_to = "metric", values_to = "value") %>% 
  ggplot(aes(state, value)) +
    geom_col() +
    facet_wrap(vars(metric), scales = "free_y", ncol = 1) + 
    scale_x_discrete(guide = guide_axis( angle = 90)),
    height = 2000,
    width = 600
)

```

### 2020-04-23 R/getus_functions.R

@@ -119,7 +119,13 @@ getus_dex <- function() {
 #'     \item{positive_increase}{increase in positive results from day before}
 #'     \item{total_test_increase}{increase from the day before}
 #' }
-#' Other details regarding the score system used are reported in the \href{https://covidtracking.com/about-data}{maintainers webpage}
+#' Other details regarding the score system used are reported in the \href{https://covidtracking.com/about-data}{maintainers webpage}.\cr
+#' \strong{Note for the use of some of some this variables by covidtracking authors:} \cr
+#' \emph{States are currently reporting two fundamentally unlike statistics: current hospital/ICU admissions and cumulative hospitalizations/ICU admissions.
+#' Across the country, this reporting is also sparse.
+#' In short: it is impossible to assemble anything resembling the real statistics for hospitalizations,
+#' ICU admissions, or ventilator usage across the United States. As a result, we will no longer provide
+#' national-level summary hospitalizations, ICU admissions, or ventilator usage statistics on our site.}
 #' @import vroom
 #' @export
 getus_tests <- function() {

### 2020-04-24 README.Rmd
readme
logo
draft implementation

@@
 
-# covid19census  <img src='inst/img/hexsticker.png' align="right" height="139" />
+# covid19census  <img src='inst/img/hexsticker2.png' align="right" height="139" >
 
@@
 
-```{r all, message=FALSE}
+```{r all, message=FALSE, eval=FALSE}
 library(covid19census)
@@
 
-Data regarding COVID-19 number of cases and deaths (at the county level) were obtained from the [New York Time git repository](https://github.com/nytimes/covid-19-data), whereas number of tests and hospitalizations from the [the Covid Tracking Project].
+Data regarding COVID-19 number of cases and deaths (at the county level) were obtained from the [New York Time git repository](https://github.com/nytimes/covid-19-data), whereas number of tests and hospitalizations from the [the Covid Tracking Project](https://covidtracking.com/api}).
 

### 2020-04-24 inst/some_checks.Rmd
draft implementation

@@
 
+```
+WHO ratio positive/total to talk about reopening
+
+```{r, message=FALSE}
+t_us %>% 
+  mutate(ratio_pos = positive/(positive + negative)) %>% 
+  filter(date == max(date)) %>%
+
+  ggplot(aes(reorder(state, ratio_pos), ratio_pos)) +
+    geom_col() +
+    scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90)) +
+    labs(x = NULL,
+         y = "positive/(positive + negative)",
+         title = "Probability of a positive test") +
+    theme(legend.position = "top",
+        legend.title = element_blank())
 ```

### 2020-04-24 R/getit_functions.R
draft alghorithm

@@ -126,7 +126,7 @@ getit_covid <- function() {
 #' get COVID-19 cases and other statistics
 #'
 #' extracts and translates time series form the git repository of the \href{https://github.com/pcm-dpc/COVID-19}{protezione civile} and
-#' combines them with other statistics related to italian population
+#' combines them with other statistics related to italian population.
 #'
 #' @return a tibble with following 64 variables:
 #' \describe{
@@ -175,7 +175,7 @@ getit_covid <- function() {
 #' Prevalence of types of cancer patients (2016), influenza-vaccination coverage (2019) and the number of hospital beds per 1000 people (2017) were obtained from
 #' \href{http://www.dati.salute.gov.it/}{Ministero della Salute}. Note that cancer patients prevalence was calculated using
 #'  region population esitmates of 2019. Data of particulate 2.5 (2017) comes from the
-#' \href{https://annuario.isprambiente.it/pon/basic/14}{Istituto Superiore Per La protexione Ambientale}.
+#' \href{https://annuario.isprambiente.it/pon/basic/14}{Istituto Superiore Per La protezione Ambientale}.
 #' @source  \href{https://github.com/pcm-dpc/COVID-19}{protezione civile}, \href{http://dati.istat.it/?lang=en}{ISTAT}
 #' @importFrom rlang .data
 #' @importFrom magrittr %>%

### 2020-04-24 R/getus_functions.R

@@ -206,11 +206,13 @@ getus_tests <- function() {
 #'  \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables},
 #'  \href{https://data.cms.gov/mapping-medicare-disparities}{Mapping Medicare Disparities},
 #'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices}
-#' @return A tibble with 309 variables. Data regarding the household composition, population age and poverty levels,  sex and extension
+#' @return A tibble with 304 variables. Data regarding the household composition, population sex and age and poverty levels,
 #'  were scraped from the 2018 American Community Survey (ACS). Poverty was defined at the family level and not the household level in
-#'  the ACS. Medical conditions, tabacco use, cancer and, data relative to the number of medical and emergency visits
-#'  was obtained from the 2017 Mapping Medicare Disparities. From relative documentation listed in the source: "Prevalence rates are calculated by searching
-#'  for certain diagnosis codes in \strong{Medicare beneficiaries’ claims}. The admission rate by admission type is the frequency of a specific type of inpatient admission per 1,000 inpatient admissions in a year." The number of hospital beds per county was calculated from data of the
+#'  the ACS. Medical conditions, tabaco use, cancer and, data relative to the number of medical and emergency visits
+#'  was obtained from the 2017 Mapping Medicare Disparities. From relative documentation listed in the source: "Prevalence rates are calculated
+#'  by searching for certain diagnosis codes in \strong{Medicare beneficiaries’ claims}. T
+#'  he admission rate by admission type is the frequency of a specific type of inpatient admission per 1,000 inpatient admissions in a year."
+#'  The number of hospital beds per county was calculated from data of the
 #'  2020 Homeland Infrastructure Foundation. \cr
 #'  \strong{Note that data on test and hospitalizations are at the state level!}
 #'  \describe{

### 2020-04-24 R/getit_functions.R
figure and table draft

@@ -117,7 +117,7 @@ getit_covid <- function() {
       )
    ]
 
-  message(paste0("Data till ", max(dat$date), " successfully extracted!"))
+  message(paste0("Italy COVID-19 data up to ", max(dat$date), " successfully retrived!"))
 
   dat
 }

### 2020-04-24 R/getus_functions.R

@@ -33,7 +33,7 @@ getus_covid <- function() {
     dplyr::filter(!is.na(.data$fips))
 
 
-  message(paste0("Data of COVID-19 up to ", max(dat$date), " successfully extracted!"))
+  message(paste0("US COVID-19 data up to ", max(dat$date), " successfully retrived!"))
 
   dat
 }
@@ -84,7 +84,7 @@ getus_dex <- function() {
     )
   )
 
-  message(paste0("Data of activity up to ", max(dat$date), " successfully imported!"))
+  message(paste0("US mobility data up to ", max(dat$date), " successfully retrived!"))
 
   dat
 }
@@ -169,7 +169,7 @@ getus_tests <- function() {
 
   colnames(dat)[colnames(dat) == "state"] <- "abbr"
 
-  message(paste0("Data of tests up to ", max(dat$date, na.rm = TRUE), " successfully imported!"))
+  message(paste0("US test data up to ", max(dat$date, na.rm = TRUE), " successfully retrived!"))
 
   # we don't have unincorporated territories in getus_all()
   dat2 <- dat[dat$abbr %in% state_abbr$abbr, ]

### 2020-04-25 README.Rmd
draft: corrplot and docum

@@
 
-This is a list of U.S. datasets that can be accessed directly or in a `tibble` containing also COVID-19 metrics:
+This is a list of U.S. datasets that can be accessed directly or in a `dataframe` containing also COVID-19 metrics:
 
@@
 
-This is a list of Italy datasets that can be accessed directly or in a `tibble` containing also COVID-19 metrics:
+This is a list of Italy datasets that can be accessed directly or in a `dataframe` containing also COVID-19 metrics:
 

### 2020-04-25 R/getit_functions.R

@@ -2,7 +2,7 @@
 #'
 #' extracts and translates time series form the git repository of the \href{https://github.com/pcm-dpc/COVID-19}{protezione civile}
 #'
-#' @return a tibble with following 19 variables:
+#' @return a dataframe with following 19 variables:
 #' \describe{
 #'      \item{date}{in `ISO 8601` format}
 #'      \item{state}{state}
@@ -128,7 +128,7 @@ getit_covid <- function() {
 #' extracts and translates time series form the git repository of the \href{https://github.com/pcm-dpc/COVID-19}{protezione civile} and
 #' combines them with other statistics related to italian population.
 #'
-#' @return a tibble with following 64 variables:
+#' @return a dataframe with following 64 variables:
 #' \describe{
 #'      \item{date}{date of data}
 #'      \item{state}{state}

### 2020-04-25 R/getus_functions.R

@@ -2,7 +2,7 @@
 #'
 #' extracts time series from the git repository of the \href{ https://github.com/nytimes/covid-19-data }{NYT}
 #'
-#' @return a tibble
+#' @return a dataframe
 #' @importFrom rlang .data
 #' @importFrom magrittr %>%
 #' @import vroom
@@ -43,7 +43,7 @@ getus_covid <- function() {
 #' extracts DEX from the git repository of the
 #' \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVID-19 exposure indeces}
 #'
-#' @return a tibble
+#' @return a dataframe
 #' @details main metric is `dex_a`. In the \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{repository}, they
 #' explains: \cite{In the context of the ongoing pandemic, the DEX measure may be biased if devices sheltering-in-place
 #' are not in the sample due to lack of movement. We report adjusted DEX values to help address this selection bias.
@@ -93,7 +93,7 @@ getus_dex <- function() {
 #'
 #' extracts information on tests, hospitalizations and other metrics at the \strong{State level} maintained by the
 #' \href{https://covidtracking.com/api}{the COVID Tracking Project}
-#' @return a tibble with 15 variables
+#' @return a dataframe with 15 variables
 #' @details a description of the variable can be found in the \href{https://covidtracking.com/api}{the COVID Tracking Project} and when possible
 #' was used verbatim for the description below
 #' \describe{
@@ -206,7 +206,7 @@ getus_tests <- function() {
 #'  \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables},
 #'  \href{https://data.cms.gov/mapping-medicare-disparities}{Mapping Medicare Disparities},
 #'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices}
-#' @return A tibble with 304 variables. Data regarding the household composition, population sex and age and poverty levels,
+#' @return A dataframe with 304 variables. Data regarding the household composition, population sex and age and poverty levels,
 #'  were scraped from the 2018 American Community Survey (ACS). Poverty was defined at the family level and not the household level in
 #'  the ACS. Medical conditions, tabaco use, cancer and, data relative to the number of medical and emergency visits
[…]

### 2020-04-27 README.Rmd
travis

@@
 [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
+[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
 <!-- badges: end -->

### 2020-04-27 README.Rmd
docu badge

--- 2020-04-25 README.Rmd
@@
 [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
+[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
+[![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 <!-- badges: end -->

### 2020-04-28 R/datasets_us.R
added race

@@ -82,3 +82,44 @@
 #' @usage data(mmd_us)
 #' @seealso \code{\link{getus_all}} for more details regarding the variables
 "mmd_us"
+
+#' race
+#'
+#' Estimate population of each county by race. American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey}
+#' @return  a tibble
+#' @usage data(poverty_us)
+"race_us"
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2020-04-28 R/getus_functions.R

@@ -206,9 +206,9 @@ getus_tests <- function() {
 #'  \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables},
 #'  \href{https://data.cms.gov/mapping-medicare-disparities}{Mapping Medicare Disparities},
 #'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices}
-#' @return A dataframe with 304 variables. Data regarding the household composition, population sex and age and poverty levels,
+#' @return A dataframe with 314 variables. Data regarding the household composition, population sex, age, race, ancestry and poverty levels,
 #'  were scraped from the 2018 American Community Survey (ACS). Poverty was defined at the family level and not the household level in
-#'  the ACS. Medical conditions, tabaco use, cancer and, data relative to the number of medical and emergency visits
+#'  the ACS. Medical conditions, tobacco use, cancer and, data relative to the number of medical and emergency visits
 #'  was obtained from the 2017 Mapping Medicare Disparities. From relative documentation listed in the source: "Prevalence rates are calculated
 #'  by searching for certain diagnosis codes in \strong{Medicare beneficiaries’ claims}. T
 #'  he admission rate by admission type is the frequency of a specific type of inpatient admission per 1,000 inpatient admissions in a year."
@@ -290,14 +290,20 @@ getus_tests <- function() {
 #'   \item{perc_withcomputer}{percent that owns or use computer}
 #'   \item{perc_withinternet}{percet that has acces to internet}
 #'   \item{\strong{POPULATION AND SEX}}{---------------}
-#'   \item{tot_pop}{total population}
-#'   \item{tot_male}{total male}
-#'   \item{tot_female}{total female}
-#'   \item{tot_ \emph{age_sex}}{total population by age bin and sex}
+#'   \item{total_pop}{total population}
+#'   \item{total_male}{total male}
+#'   \item{total_female}{total female}
+#'   \item{total_ \emph{age_sex}}{total population by age bin and sex}
 #'   \item{perc_ \emph{age_sex}}{percent population by age bin and sex}
+#'   \item{\strong{POPULATION AND RACE}}{---------------}
[…]

### 2020-04-28 data-raw/import_raw.R

@@ -2,10 +2,12 @@ library(vroom)
 library(tidyverse)
 library(janitor)
 library(reshape2)
+library(tabulizer)
 
+# Unpackit_us.zip and run the script
 
+# xxxxxxxxxxxxxxxx-------------------
 # United States #-----
-
 # xxxxxxxxxxxxxxxx-------------------
 # age and sex- age_sex_us ---------
 
@@ -209,7 +211,25 @@ poverty_us <- poverty_all_us %>%
 
 state_abbr <- vroom("data-raw/us/state_abbr.csv", col_types = cols(.default = "c"))
 
+# race: race_us ------------
+to_select_race <- readxl::read_xlsx("data-raw/us/american_comunity_survey_2018/race/to_select_race.xlsx")
 
+race_all_us <- vroom("data-raw/us/american_comunity_survey_2018/race/ACSDT5Y2018.B02001_data_with_overlays_2020-04-28T102942.csv",
+                           col_types = cols(.default = "c"), skip = 1
+) %>%
+  clean_names() %>%
+  select(!!to_select_race$old_names)
+
+names(race_all_us) <- to_select_race$new_names
+
+race_us <- race_all_us %>%
+  separate(id, into = c("us_id", "fips"), sep = 9) %>%
+  select(-us_id) %>%
+  separate(geographic_area_name, c("county", "state"), sep = ",") %>%
+  mutate_at(vars(-county, -state), as.numeric)
+
+
+# xxxxxxxxxxxxxxxx-------------------
 # Italy ------------------------
 # xxxxxxxxxxxxxxxx-------------------
 # bweight_it -----
@@ -780,6 +800,7 @@ usethis::use_data(
   house_it,
   netinc_it,
   pm2.5_it,
+  race_us,
   regions_area,
   smoking_it,
   # internal = TRUE,
@@ -787,33 +808,34 @@ usethis::use_data(
 )
 #
 # # internal data
-# usethis::use_data(
-#   # US data
-#   acm_househ_us,
-#   age_sex_us,
-#   fl65_us,
-#   hospbeds_us,
-#   mmd_us,
-#   poverty_us,
-#   state_abbr,
-#   # italy
-#   bweight_it,
-#   cancer_it,
-#   chronic_it,
-#   dem_65bin_fm,
-#   dem_it,
-#   firstaid_it,
[…]

### 2020-04-28 R/getus_functions.R
documentation getus_all extended

@@ -214,6 +214,8 @@ getus_tests <- function() {
 #'  he admission rate by admission type is the frequency of a specific type of inpatient admission per 1,000 inpatient admissions in a year."
 #'  The number of hospital beds per county was calculated from data of the
 #'  2020 Homeland Infrastructure Foundation. \cr
+#'  The following list of variables is divided in sections \emph{COVID-19 VARS, HOUSEHOLDS MARITAL STATUS AND COMPOSITION, HOUSEHOLDS EDUCATION DEGREES,
+#'  ANCESTRY, COMPUTER OR INTERNET, POPULATION AND SEX, POPULATION AND RACE, MEDICAL AND VACCINES, POVERTY, STATE LEVEL TESTS AND HOSPITALIZATIONS, ACTIVITY}. \cr
 #'  \strong{Note that data on test and hospitalizations are at the state level!}
 #'  \describe{
 #'   \item{date}{formatted `ISO 8601`}
@@ -221,7 +223,7 @@ getus_tests <- function() {
 #'   \item{state}{state}
 #'   \item{fips}{federal information processing standard, a unique numeric identifier of a county}
 #'   \item{urban}{urban or rural (not sure about the  definition)}
-#'   \item{\strong{COVID-19 vars}}{---------------}
+#'   \item{\strong{COVID-19 VARS}}{---------------}
 #'   \item{cases}{confirmed COVID-19 cases (cumulates with date)}
 #'   \item{deaths}{number of deaths attributed to COVID-19}
 #'   \item{cmr}{case-mortality rate (deaths / confirmed cases * 100)}
@@ -296,7 +298,7 @@ getus_tests <- function() {
 #'   \item{total_ \emph{age_sex}}{total population by age bin and sex}
 #'   \item{perc_ \emph{age_sex}}{percent population by age bin and sex}
 #'   \item{\strong{POPULATION AND RACE}}{---------------}
-#'   \item{total \emph{race}}{total number of people of that race. It can be normalized using total_population}
+#'   \item{total \emph{race}}{total number of people of that race (white, black, native, asian, island, other). It can be normalized using total_population}
 #'   \item{totat_other1race}{estimate total some other race alone}
 #'   \item{totat_2races}{estimate total two or more races}
 #'   \item{total_2races_other}{estimate total two or more races including other race}

### 2020-04-28 README.Rmd
added race
pm2.5 for us added

@@
 The other datsets were obtained from the [Homeland Infrastructure Foundation](https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18), the [Census](https://data.census.gov/cedsci/table?q=United%20States), [Mapping Medicare Disparities](https://data.cms.gov/mapping-medicare-disparities), and [activity indexes](https://github.com/COVIDExposureIndices/COVIDExposureIndices) calculated by  Victor Couture, Jonathan Dingel, Allison Green, Jessie Handbury, and Kevin Williams based on smartphone movement data provided by `PlaceIQ`.
+Emissions of pm2.5 in micro g/m3 (2016) were reported by \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group} and
+#'  aggregate by \href{https://github.com/wxwx1993/PM_COVID/blob/master/additional_preprocessing_code/download_pm25_values.md}{Ista Zahn and Ben Sabath}
 
@@
 
-* `act_ind`: county-level device exposure index (DEX), an index of activity.
-* `acm_househ`: several metric regarding household composition (2018).
-* `age_sex`: age and sex distribution (2018).
-* `fl65`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
-* `hospbeds`: total hospital beds in each county (2019).
-* `mmd`: data of 2017 regarding the  prevalence found in medicare beneficiaries of 
+* `act_ind_us`: county-level device exposure index (DEX), an index of activity.
+* `acm_househ_us`: several metric regarding household composition, including poverty, education, ancestry (2018).
+* `age_sex_us`: age and sex distribution (2018).
+* `race_us`: estimated races in population (2018).
+* `fl65_us`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
+* `hospbeds_us`: total hospital beds in each county (2019).
+* `mmd_us`: data of 2017 regarding the  prevalence found in medicare beneficiaries of 
     * many medical and chronic conditions (Alzheimer, chronic kidney, obesity, depression, obstructive pulmonary, disease, arthritis, diabetes, osteoporosis, asthma, atrial  fibrillation, ischemic hearth, myocardial infarction, hypertension).
@@
     * tobacco use.
+* `pm2.5_us`: Emissions of pm2.5 in micro g/m3
 

### 2020-04-28 R/datasets_it.R

@@ -128,7 +128,7 @@
 "netinc_it"
 
 
-#' particulate 2.5
+#' particulate 2.5 italy
 #'
 #' Emission of pm2.5 in tons per region from 1990 to 2017
 #' @docType data

### 2020-04-28 R/datasets_us.R

@@ -90,36 +90,16 @@
 #' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
 #' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey}
 #' @return  a tibble
-#' @usage data(poverty_us)
+#' @usage data(race_us)
 "race_us"
 
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
+#' particulate 2.5
+#'
+#' Emission of pm2.5 in  micro g/m3 per county from 2000 to 2016
+#' @docType data
+#' @source \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group},
+#' \href{https://github.com/wxwx1993/PM_COVID/blob/master/Data/county_pm25.csv}{wxwk1993 processed data}
+#' @details  \href{https://github.com/wxwx1993/PM_COVID/blob/master/additional_preprocessing_code/download_pm25_values.md}{Ista Zahn and Ben Sabath repo}
+#' @return  a tibble
+#' @usage data(pm2.5_us)
+"pm2.5_us"

### 2020-04-28 R/getit_functions.R

@@ -32,10 +32,9 @@
 #' @import vroom
 #' @export
 getit_covid <- function() {
+  url_data <- "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv"
 
-  url_data <-  "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv"
-
-  if(RCurl::url.exists(url_data) == FALSE){
+  if (RCurl::url.exists(url_data) == FALSE) {
     stop("Something wrong with the repository or your internet connection!")
   }
 
@@ -86,36 +85,37 @@ getit_covid <- function() {
       note_it = "_",
       note_eng = "_"
     )
-  )  %>%
+  ) %>%
     dplyr::mutate(
       cmr = .data$deaths / .data$total_cases * 100
     ) %>%
     dplyr::mutate(region = stringr::str_replace(.data$region, "-", " "))
 
   # just reorder
-  dat <- dat[,
-      c(
-        "date",
-        "state",
-        "region_cod",
-        "region",
-        "lat",
-        "long",
-        "cmr",
-        "total_cases",
-        "deaths",
-        "tests",
-        "hospitalized_with_symptoms",
-        "intensive_care_unit",
-        "total_hospitalized",
-        "home_quarantine",
-        "total_positives",
-        "change_positives",
-        "new_positives",
-        "recovered_released",
-        "people_tested"
-      )
-   ]
+  dat <- dat[
+    ,
+    c(
+      "date",
+      "state",
+      "region_cod",
+      "region",
+      "lat",
+      "long",
+      "cmr",
+      "total_cases",
+      "deaths",
+      "tests",
+      "hospitalized_with_symptoms",
+      "intensive_care_unit",
+      "total_hospitalized",
+      "home_quarantine",
+      "total_positives",
+      "change_positives",
+      "new_positives",
+      "recovered_released",
+      "people_tested"
+    )
+  ]
 
   message(paste0("Italy COVID-19 data up to ", max(dat$date), " successfully retrived!"))
 

### 2020-04-28 R/getus_functions.R

@@ -11,10 +11,9 @@
 #' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
 #' @export
 getus_covid <- function() {
+  url_data <- "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"
 
-  url_data <-  "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"
-
-  if(RCurl::url.exists(url_data) == FALSE){
+  if (RCurl::url.exists(url_data) == FALSE) {
     stop("Something wrong with the repository or your internet connection!")
   }
 
@@ -54,10 +53,9 @@ getus_covid <- function() {
 #' @import vroom
 #' @export
 getus_dex <- function() {
+  url_data <- "https://raw.githubusercontent.com/COVIDExposureIndices/COVIDExposureIndices/master/dex_data/county_dex.csv"
 
-  url_data <-  "https://raw.githubusercontent.com/COVIDExposureIndices/COVIDExposureIndices/master/dex_data/county_dex.csv"
-
-  if(RCurl::url.exists(url_data) == FALSE){
+  if (RCurl::url.exists(url_data) == FALSE) {
     stop("Something wrong with the repository or your internet connection!")
   }
 
@@ -129,9 +127,9 @@ getus_dex <- function() {
 #' @import vroom
 #' @export
 getus_tests <- function() {
-  url_data <-  "https://covidtracking.com/api/states/daily.csv"
+  url_data <- "https://covidtracking.com/api/states/daily.csv"
 
-  if(RCurl::url.exists(url_data) == FALSE){
+  if (RCurl::url.exists(url_data) == FALSE) {
     stop("Something wrong with the repository or your internet connection!")
   }
 
@@ -205,168 +203,163 @@ getus_tests <- function() {
 #'  \href{https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18}{Homeland Infrastructure Foundation-Level Data},
 #'  \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables},
 #'  \href{https://data.cms.gov/mapping-medicare-disparities}{Mapping Medicare Disparities},
-#'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices}
+#'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices},
[…]

### 2020-04-28 data-raw/import_raw.R

@@ -11,14 +11,14 @@ library(tabulizer)
 # xxxxxxxxxxxxxxxx-------------------
 # age and sex- age_sex_us ---------
 
-age_sex_all_us <- vroom("data-raw/us/american_comunity_survey_2018/age_sex/ACSST5Y2018.S0101_data_with_overlays_2020-04-14T063202.csv",
+age_sex_all_us <- vroom("data-raw/it_us/us/american_comunity_survey_2018/age_sex/ACSST5Y2018.S0101_data_with_overlays_2020-04-14T063202.csv",
   skip = 1,
   col_select = c(`id`, `Geographic Area Name`, starts_with("Estimate"))
 ) %>%
   clean_names()
 
 
-to_select_age_sex <- readxl::read_xlsx("data-raw/us/american_comunity_survey_2018/age_sex/to_select_age_sex.xlsx")
+to_select_age_sex <- readxl::read_xlsx("data-raw/it_us/us/american_comunity_survey_2018/age_sex/to_select_age_sex.xlsx")
 
 age_sex_us <- age_sex_all_us %>%
   select(!!to_select_age_sex$old_names)
@@ -33,7 +33,7 @@ age_sex_us <- age_sex_us %>%
 
 
 # flu vax - fl65_us  ----------
-fl65_all_us <- vroom("data-raw/us/flu_us.csv",
+fl65_all_us <- vroom("data-raw/it_us/us/flu_us.csv",
   col_types = cols(
     year = col_double(),
     geography = col_character(),
@@ -60,7 +60,7 @@ fl65_us <- fl65_all_us %>%
   dplyr::select(state, county, fips, imm65)
 
 # hospital beds - hospbeds_us  ----------
-hospbeds_all_us <- vroom("data-raw/us/hospital_beds.csv",
+hospbeds_all_us <- vroom("data-raw/it_us/us/hospital_beds.csv",
   col_types = cols(
     X = col_double(),
     Y = col_double(),
@@ -114,7 +114,7 @@ hospbeds_us <- hospbeds_all_us %>%
   filter(!is.na(fips))
 
 # Mapping Medicare Disparities - mmd_us  ----------
-to_imp <- list.files("data-raw/us/data_cms/", full.names = TRUE)
+to_imp <- list.files("data-raw/it_us/us/data_cms/", full.names = TRUE)
 mmd_all_us <- vroom(to_imp,
   col_types = cols(
[…]

### 2020-04-28 tests/testthat/test-getit.R

@@ -59,4 +59,3 @@ test_that("sum smoking is < 100 ", {
     0
   )
 })
-

### 2020-04-29 R/datasets_us.R
added netinc, hispanic  fixed (life is sad)

@@ -103,3 +103,15 @@
 #' @return  a tibble
 #' @usage data(pm2.5_us)
 "pm2.5_us"
+
+#' netinc_us
+#'
+#' Median Household income, 2018
+#' @docType data
+#' @source \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group},
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey}
+#' @return  a tibble
+#' @usage data(netinc_us)
+"netinc_us"
+

### 2020-04-29 R/getus_functions.R

@@ -205,7 +205,7 @@ getus_tests <- function() {
 #'  \href{https://data.cms.gov/mapping-medicare-disparities}{Mapping Medicare Disparities},
 #'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices},
 #'  \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group}
-#' @return A dataframe with 314 variables. Data regarding the household composition, population sex, age, race, ancestry and poverty levels,
+#' @return A dataframe with 326 variables. Data regarding the household composition, population sex, age, race, ancestry and poverty levels,
 #'  were scraped from the 2018 American Community Survey (ACS). Poverty was defined at the family level and not the household level in
 #'  the ACS. Medical conditions, tobacco use, cancer and, data relative to the number of medical and emergency visits
 #'  was obtained from the 2017 Mapping Medicare Disparities. From relative documentation listed in the source: "Prevalence rates are calculated
@@ -289,11 +289,26 @@ getus_tests <- function() {
 #'     \item{total_ \emph{age_sex}}{total population by age bin and sex}
 #'     \item{perc_ \emph{age_sex}}{percent population by age bin and sex}
 #'     \item{\strong{POPULATION AND RACE}}{---------------}
-#'     \item{total \emph{race}}{total number of people of that race (white, black, native, asian, island, other). It can be normalized using total_population}
-#'     \item{totat_other1race}{estimate total some other race alone}
-#'     \item{totat_2races}{estimate total two or more races}
-#'     \item{total_2races_other}{estimate total two or more races including other race}
-#'     \item{total_2races_exlusion}{estimate total two or more races exluding some other arce and three or more races}
+#'     \item{total_nlat}{total not hispanic or latinos. It can be normalized using total_population}
+#'     \item{total_nlat_white_alone}{total white in that category}
+#'     \item{total_nlat_blackaa_alone}{total black or african american}
+#'     \item{total_nlat_native_alone}{total american indian and alaska native}
+#'     \item{total_nlat_asian_alone}{total asian}
+#'     \item{total_nlat_island_alone}{total native hawaiaian and other pacific islander}
+#'     \item{total_nlat_other_race}{total other race}
[…]

### 2020-04-29 data-raw/import_raw.R

@@ -214,7 +214,7 @@ state_abbr <- vroom("data-raw/it_us/us/state_abbr.csv", col_types = cols(.defaul
 # race: race_us ------------
 to_select_race <- readxl::read_xlsx("data-raw/it_us/us/american_comunity_survey_2018/race/to_select_race.xlsx")
 
-race_all_us <- vroom("data-raw/it_us/us/american_comunity_survey_2018/race/ACSDT5Y2018.B02001_data_with_overlays_2020-04-28T102942.csv",
+race_all_us <- vroom("data-raw/it_us/us/american_comunity_survey_2018/race/ACSDT5Y2018.B03002_data_with_overlays_2020-04-28T232845.csv",
   col_types = cols(.default = "c"), skip = 1
 ) %>%
   clean_names() %>%
@@ -225,7 +225,7 @@ names(race_all_us) <- to_select_race$new_names
 race_us <- race_all_us %>%
   separate(id, into = c("us_id", "fips"), sep = 9) %>%
   select(-us_id) %>%
-  separate(geographic_area_name, c("county", "state"), sep = ",") %>%
+  separate(state_county, c("county", "state"), sep = ",") %>%
   mutate_at(vars(-county, -state), as.numeric)
 
 # pm2.5_us ---------------------
@@ -240,6 +240,23 @@ pm2.5_us <- vroom("https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Da
   )
 )
 
+# netinc_us -----
+
+
+netincome_all_us <- vroom("data-raw/it_us/us/american_comunity_survey_2018/netincome/ACSST5Y2018.S1901_data_with_overlays_2020-04-29T001634.csv",
+                     col_types = cols(.default = "c"), skip = 1
+) %>%
+  clean_names() %>%
+  select(id, geographic_area_name, estimate_households_median_income_dollars)
+
+names(netincome_all_us)
+
+netinc_us <- netincome_all_us %>%
+  separate(id, into = c("us_id", "fips"), sep = 9) %>%
+  select(-us_id) %>%
+  separate(geographic_area_name, c("county", "state"), sep = ",") %>%
+  rename(median_income = estimate_households_median_income_dollars) %>%
+  mutate_at(vars(-county, -state), as.numeric)
 
 # xxxxxxxxxxxxxxxx-------------------
 # Italy ------------------------
[…]

### 2020-04-30 R/datasets_us.R
added season, pm2.5 mean

@@ -115,3 +115,15 @@
 #' @usage data(netinc_us)
 "netinc_us"
 
+#' seasonal temperature and humidity
+#'
+#' Seasonal temperature and humidity
+#' @docType data
+#' @source \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group},
+#' \href{https://github.com/wxwx1993/PM_COVID/blob/master/Data/county_pm25.csv}{wxwk1993 processed data}
+#' @details  \href{https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Data/temp_seasonal_county.csv}{Ista Zahn and Ben Sabath repo}
+#' @return  a tibble
+#' @usage data(season_us)
+"season_us"
+
+

### 2020-04-30 R/getus_functions.R

@@ -127,7 +127,7 @@ getus_dex <- function() {
 #' @import vroom
 #' @export
 getus_tests <- function() {
-  url_data <- "https://covidtracking.com/api/states/daily.csv"
+  url_data <- "https://covidtracking.com/api/v1/states/daily.csv"
 
   if (RCurl::url.exists(url_data) == FALSE) {
     stop("Something wrong with the repository or your internet connection!")
@@ -205,17 +205,17 @@ getus_tests <- function() {
 #'  \href{https://data.cms.gov/mapping-medicare-disparities}{Mapping Medicare Disparities},
 #'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices},
 #'  \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group}
-#' @return A dataframe with 326 variables. Data regarding the household composition, population sex, age, race, ancestry and poverty levels,
+#' @return A dataframe with 330 variables. Data regarding the household composition, population sex, age, race, ancestry and poverty levels,
 #'  were scraped from the 2018 American Community Survey (ACS). Poverty was defined at the family level and not the household level in
 #'  the ACS. Medical conditions, tobacco use, cancer and, data relative to the number of medical and emergency visits
 #'  was obtained from the 2017 Mapping Medicare Disparities. From relative documentation listed in the source: "Prevalence rates are calculated
 #'  by searching for certain diagnosis codes in \strong{Medicare beneficiaries’ claims}. The admission rate by admission type is the frequency of
 #'  a specific type of inpatient admission per 1,000 inpatient admissions in a year."
 #'  The number of hospital beds per county was calculated from data of the2020 Homeland Infrastructure Foundation.
-#'  Emissions of particulate 2.5 in micro g/m3 (2016) were reported by \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group} and
[…]

### 2020-04-30 data-raw/import_raw.R

@@ -240,9 +240,19 @@ pm2.5_us <- vroom("https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Da
   )
 )
 
+# season_us ---------------------
+# Thank you  Ista Zahn and Ben Sabath for hints on the sources
+# https://github.com/wxwx1993/PM_COVID/blob/master/additional_preprocessing_code/download_pm25_values.md
+# The Atmospheric Composition Analysis Group at Dalhouse University
+
+season_us <- vroom("https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Data/temp_seasonal_county.csv",
+                    col_types = cols(.default = "d")
+                  )
+
+names(season_us) <- c("fips", "year", "summer_temp", "summer_hum", "winter_temp", "winter_hum")
+
 # netinc_us -----
-
-
+# from census
 netincome_all_us <- vroom("data-raw/it_us/us/american_comunity_survey_2018/netincome/ACSST5Y2018.S1901_data_with_overlays_2020-04-29T001634.csv",
                      col_types = cols(.default = "c"), skip = 1
 ) %>%
@@ -832,15 +842,29 @@ usethis::use_data(
   race_us,
   regions_area,
   smoking_it,
+  season_us,
   # internal = TRUE,
   overwrite = TRUE
 )
 #
 
-pm2.5_us <- pm2.5_us %>%
-  filter(year == 2016) %>%
-  rename(pm2.5 = pm25) %>%
-  select(fips, pm2.5)
+
+# they take the mean of all the different years, we do the same
+pm2.5_us <-
+  pm2.5_us %>%
+  group_by(fips) %>%
+  summarize(pm2.5 = mean(pm25)) %>%
+  ungroup()
+
+season_us <-
+  season_us %>%
+  group_by(fips) %>%
+  # easy summarization of each variable...
+  summarize_at(
+    vars(summer_temp:winter_hum), mean
+    ) %>%
+  ungroup()
+
 
 # # internal data
 usethis::use_data(
@@ -870,6 +894,7 @@ usethis::use_data(
   race_us,
   regions_area,
   smoking_it,
+  season_us,
   internal = TRUE,
   overwrite = TRUE
 )

### 2020-04-30 README.Rmd
added netinc, hispanic  fixed (life is sad)
cleaned up names

@@
 
-* `act_ind_us`: county-level device exposure index (DEX), an index of activity.
-* `acm_househ_us`: several metric regarding household composition, including poverty, education, ancestry (2018).
-* `age_sex_us`: age and sex distribution (2018).
-* `race_us`: estimated races in population (2018).
-* `fl65_us`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
-* `hospbeds_us`: total hospital beds in each county (2019).
-* `mmd_us`: data of 2017 regarding the  prevalence found in medicare beneficiaries of 
+* `us_acm_househ`: several metric regarding household composition, including poverty, education, ancestry (2018).
+* `us_age_sex`: age and sex distribution (2018).
+* `us_fl65`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
+* `us_hospbeds`: total hospital beds in each county (2019).
+* `us_netinc`: median househould income (2018).
+* `us_mmd`: data of 2017 regarding the  prevalence found in medicare beneficiaries of 
     * many medical and chronic conditions (Alzheimer, chronic kidney, obesity, depression, obstructive pulmonary, disease, arthritis, diabetes, osteoporosis, asthma, atrial  fibrillation, ischemic hearth, myocardial infarction, hypertension).
@@
     * tobacco use.
-* `pm2.5_us`: Emissions of pm2.5 in micro g/m3
+* `us_pm2.5`: Emissions of pm2.5 in micro g/m3
+* `us_race`: estimated races in population (2018).
 
@@
 
-* `bweight_it`: Body mass index in regions of Italy, in the general population.
-* `cancer_it`: Number of cancer patients in each region by type.
-* `chronic_it`: Number of people suffering of chronic conditions by region and type.
-* `dem_it`: Number of people by region, sex, age.
-* `firstaid_it`: Number of people using first aid or medical guard (3 months). 
-* `fl_it`: Influenza vaccination coverage in Italy for 2009-2019 season for general population and aged 65 and more.
-* `fl_it_2019`: Influenza vaccination coverage in Italy for 2018-2019 season for general population and 65 and more.
-* `hospbed_it`: Inpatient hospital beds per 1000 people.
-* `house_it`:  household crowding index from 2014 to 2018 in each region.
-* `netinc_it`:  median net annual households income (including imputed rents, in euros).
-* `pm2.5_it`:  emission of particulate 2.5 in tons per region.
-* `regions_area`: Area in square meters of each region. Used to calculate density per region.
-* `smoking_it`: Number of people aged 14 years and over that self-referred as smoker, non smoker, or past smoker by region.
+* `it_bweight`: Body mass index in regions of Italy, in the general population.
+* `it_cancer`: Number of cancer patients in each region by type.
+* `it_chronic`: Number of people suffering of chronic conditions by region and type.
+* `it_dem`: Number of people by region, sex, age.
+* `it_firstaid`: Number of people using first aid or medical guard (3 months). 
+* `it_fl`: Influenza vaccination coverage in Italy for 2009-2019 season for general population and aged 65 and more.
+* `it_fl_2019`: Influenza vaccination coverage in Italy for 2018-2019 season for general population and 65 and more.
+* `it_hospbed`: Inpatient hospital beds per 1000 people.
+* `it_house`:  household crowding index from 2014 to 2018 in each region.
+* `it_netinc`:  median net annual households income (including imputed rents, in euros).
+* `it_pm2.5`:  emission of particulate 2.5 in tons per region.
+* `it_regions`: Area in square meters of each region. Used to calculate density per region.
+* `it_smoking`: Number of people aged 14 years and over that self-referred as smoker, non smoker, or past smoker by region.
 

### 2020-04-30 data-raw/it_us/it/how_get_data.Rmd
cleaned up names

--- 2020-04-21 data-raw/it/how_get_data.Rmd
@@
 
-fl_it <- fl_tab %>%
+it_fl <- fl_tab %>%
   rename(region = X) %>%
@@
 
-# fl_it
+# it_fl
 ```
@@
 
-```{r plot_fl_it2019}
-fl_it_2019 <- fl_it %>%
+```{r plot_it_fl2019}
+it_fl_2019 <- it_fl %>%
   rename(perc_imm = "2019") %>%
@@
 
-fl_it_2019 %>%
+it_fl_2019 %>%
   ggplot(aes(reorder(region, perc_imm), perc_imm)) +
@@
 
-```{r fl65_it}
+```{r it_fl65}
 area_sel2 <- c(
@@
 
-fl65_it <- fl65_tab %>%
+it_fl65 <- fl65_tab %>%
   rename(region = X) %>%
@@
 
-fl65_it_2019 <- fl65_it %>%
+it_fl65_2019 <- it_fl65 %>%
   rename(perc_imm65 = "2019") %>%
@@
 
-```{r fl_it_all}
-fl_it_2019 <- inner_join(fl65_it_2019, fl_it_2019, by = "region")
+```{r it_fl_all}
+it_fl_2019 <- inner_join(it_fl65_2019, it_fl_2019, by = "region")
 ```
@@
 
-```{r dem_it_p}
-dem_it_p <- vroom(
+```{r it_dem_p}
+it_dem_p <- vroom(
   "data/italy/pop_it.csv",
@@
 
-dem_it_p$region <- recode(
-  dem_it_p$region,
+it_dem_p$region <- recode(
+  it_dem_p$region,
   "Valle d'Aosta / Vallée d'Aoste" = "Valle d'Aosta",
@@
 ```{r trentino}
-dem_it_p %>%
+it_dem_p %>%
   filter(region %in% list("P.A. Bolzano", "P.A. Trento", "Trentino Alto Adige / Südtirol")) %>%
@@
 ```{r percent}
-dem_it_p <- dem_it_p %>%
+it_dem_p <- it_dem_p %>%
   filter(region != "Trentino Alto Adige / Südtirol") %>%
@@
 ```{r plot_dem, fig.height= 10, fig.width=10}
-dem_it_p %>%
+it_dem_p %>%
   ggplot(aes(age, perc_pop, group = sex, color = sex)) +
@@
 ```{r dem_i_binned}
-dem_it_p <- dem_it_p %>%
+it_dem_p <- it_dem_p %>%
   mutate(age_bins = cut(age, seq(0, 100, 10), right = FALSE, include.lowest = TRUE)) %>%
@@
 # if we want to use caret or any package for regression we need to have a variable age-sex bins as columns
-dem_it_wider <- dem_it_p %>%
+it_dem_wider <- it_dem_p %>%
   mutate(sex_age_bin = paste0("perc_", sex, "_", age_bins)) %>% # now that we have the perc we create the new factor
@@
 
-glimpse(dem_it_p)
+glimpse(it_dem_p)
 ```
@@
 ```{r check}
-apply(dem_it_wider[, 3:ncol(dem_it_wider)], 1, sum)
+apply(it_dem_wider[, 3:ncol(it_dem_wider)], 1, sum)
 ```
@@
 
-```{r regions_area}
+```{r it_regions}
 wiki <- xml2::read_html("https://it.wikipedia.org/wiki/Regioni_d%27Italia")
 
-regions_area_p <- wiki %>%
+it_regions_p <- wiki %>%
   rvest::html_nodes("table") %>%
@@
 # let's clean the names
-regions_area <- regions_area_p %>%
+it_regions <- it_regions_p %>%
   select(Regione, `Superficie (km²)`) %>%
@@
 #  we add those missing regions
-regions_area <- bind_rows(regions_area, tribble(
+it_regions <- bind_rows(it_regions, tribble(
   ~region, ~area_km2,
@@
 
-chronic_it_p <- vroom("data/italy/chronic_conditions.csv",
+it_chronic_p <- vroom("data/italy/chronic_conditions.csv",
   col_types = cols(
@@
 # we do some cleaning and translations
-chronic_it <- chronic_it_p %>%
+it_chronic <- it_chronic_p %>%
   rename(region = territorio, ch_condition = tipo_dato_avq) %>%
@@
 
-```{r smoking_it, warning=FALSE, message=FALSE}
-smoking_it <- vroom("data/italy/smoking.csv",
+```{r it_smoking, warning=FALSE, message=FALSE}
+it_smoking <- vroom("data/italy/smoking.csv",
   col_types = cols(
@@
 ```{r obesity_it, warning=FALSE, message=FALSE}
-bweight_it <- vroom("data/italy/bweight_it.csv", col_types = cols(
+it_bweight <- vroom("data/italy/it_bweight.csv", col_types = cols(
   ITTER107 = col_character(),
@@
 
-```{r cancer_it}
+```{r it_cancer}
 web_page <- "http://www.registri-tumori.it/PDF/AIOM2016/I_numeri_del_cancro_2016.pdf"
@@
 # clean names
-cancer_it <- cancer_tab %>%
+it_cancer <- cancer_tab %>%
   mutate(region = recode(region,
@@
 
-cancer_it[20, "region"] <- "P.A. Bolzano"
-cancer_it[21, "region"] <- "P.A. Trento"
+it_cancer[20, "region"] <- "P.A. Bolzano"
+it_cancer[21, "region"] <- "P.A. Trento"
 ```
@@
 # Join fl with cmr
-fl_cmr_it <- inner_join(cmr_it, fl_it_2019, by = "region") %>%
+fl_cmr_it <- inner_join(cmr_it, it_fl_2019, by = "region") %>%
   mutate_at(vars("perc_imm65","perc_imm"), as.numeric)
@@
 # join  with demographics
-fl_cmr_dem_it <- inner_join(fl_cmr_it, dem_it_wider, by = "region")
+fl_cmr_it_dem <- inner_join(fl_cmr_it, it_dem_wider, by = "region")
 
 # adding regions km2
-fl_cmr_dem_it <- inner_join(fl_cmr_dem_it, regions_area, by = "region") %>%
+fl_cmr_it_dem <- inner_join(fl_cmr_it_dem, it_regions, by = "region") %>%
   mutate(pop_km2 = pop_tot / area_km2) %>%
@@
 # join health data
-fl_cmr_dem_chronic_it <- inner_join(fl_cmr_dem_it, chronic_it, by = "region")
+fl_cmr_it_dem_chronic <- inner_join(fl_cmr_it_dem, it_chronic, by = "region")
 
 # join cancer data
-fl_cmr_dem_health_cancer_it <- inner_join(fl_cmr_dem_chronic_it, cancer_it, by = "region")
+fl_cmr_dem_health_it_cancer <- inner_join(fl_cmr_it_dem_chronic, it_cancer, by = "region")
 
 # join smoking data
-fl_cmr_dem_health_cancer_smoking_it <- inner_join(fl_cmr_dem_health_cancer_it, smoking_it, by = "region")
+fl_cmr_dem_health_cancer_it_smoking <- inner_join(fl_cmr_dem_health_it_cancer, it_smoking, by = "region")
 
 # join bweight
-fl_cmr_dem_health_cancer_smoking_bweight_it <- inner_join(fl_cmr_dem_health_cancer_smoking_it, bweight_it, by = "region")
+fl_cmr_dem_health_cancer_it_smoking_bweight <- inner_join(fl_cmr_dem_health_cancer_it_smoking, it_bweight, by = "region")
 
@@
 
-all_dat <- fl_cmr_dem_health_cancer_smoking_bweight_it
+all_dat <- fl_cmr_dem_health_cancer_it_smoking_bweight
 

### 2020-04-30 R/datasets_it.R

@@ -5,8 +5,8 @@
 #' @docType data
 #' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
 #' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
-#' @usage data(bweight_it)
-"bweight_it"
+#' @usage data(it_bweight)
+"it_bweight"
 
 #' cancer patients
 #'
@@ -15,9 +15,9 @@
 #' P.A. Bolzano are missing (but we have Trentino Alto Adige)
 #' @docType data
 #' @source \href{http://www.registri-tumori.it/PDF/AIOM2016/I_numeri_del_cancro_2016.pdf}{Istituto Superiore Sanita'}
-#' @usage data(cancer_it)
+#' @usage data(it_cancer)
 #' @return a tibble
-"cancer_it"
+"it_cancer"
 
 #' smoking status
 #'
@@ -26,9 +26,9 @@
 #' @docType data
 #' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
 #' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
-#' @usage data(smoking_it)
+#' @usage data(it_smoking)
 #' @return a tibble
-"smoking_it"
+"it_smoking"
 
 #' chronic conditions
 #'
@@ -37,9 +37,9 @@
 #' @docType data
 #' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
 #' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
-#' @usage data(chronic_it)
+#' @usage data(it_chronic)
 #' @return a tibble
-"chronic_it"
+"it_chronic"
 
 
 #' influenza vaccination coverage, general population, time series
@@ -47,22 +47,22 @@
 #' Influenza vaccination coverage in Italy in the \strong{general population} from 1999 to 2019. Data are percent of region population
 #' @docType data
 #' @source \href{http://www.salute.gov.it/imgs/C_17_tavole_19_allegati_iitemAllegati_0_fileAllegati_itemFile_3_file.pdf}{Ministero della Salute}
-#' @usage data(fl_it)
-"fl_it"
+#' @usage data(it_fl)
+"it_fl"
 
 #' influenza vaccination coverage 2019
 #'
[…]

### 2020-04-30 R/datasets_us.R

@@ -5,8 +5,8 @@
 #' @source \href{https://data.cms.gov/mapping-medicare-disparities}{Data.CMS.gov}
 #' @details \href{https://www.cms.gov/About-CMS/Agency-Information/OMH/Downloads/Mapping-Technical-Documentation.pdf}{Center for Medicare and Medicaid Services} and NORC at the University of Chicago.
 #' @return  tibble wotj `fl_65` indicating the percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination
-#' @usage data(fl65_us)
-"fl65_us"
+#' @usage data(us_fl65)
+"us_fl65"
 
 #' hospital beds
 #'
@@ -14,8 +14,8 @@
 #' @docType data
 #' @source \href{https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18}{Homeland Infrastructure Foundation-Level Data}
 #' @return  a tibble
-#' @usage data(hospbeds_us)
-"hospbeds_us"
+#' @usage data(us_hospbeds)
+"us_hospbeds"
 
 #' household composition
 #'
@@ -24,8 +24,8 @@
 #' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
 #' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions}
 #' @return  a tibble
-#' @usage data(acm_househ_us)
-"acm_househ_us"
+#' @usage data(us_acm_househ)
+"us_acm_househ"
 
 #' age and sex
 #'
@@ -33,8 +33,8 @@
 #' @docType data
 #' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
 #' @return  a tibble
-#' @usage data(dem_us)
-"dem_us"
+#' @usage data(us_dem)
+"us_dem"
 
 #' poverty
 #'
@@ -43,8 +43,8 @@
 #' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
 #' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey}
 #' @return  a tibble
-#' @usage data(poverty_us)
[…]

### 2020-04-30 R/getit_functions.R

@@ -182,19 +182,19 @@ getit_covid <- function() {
 #' @importFrom dplyr vars
 #' @importFrom stats na.omit
 #' @export
-#' @seealso for details regarding  the methodology of specific datasets check  \code{\link{bweight_it}}, \code{\link{cancer_it}},
-#' \code{\link{chronic_it}}, \code{\link{dem_it}}, \code{\link{firstaid_it}}, \code{\link{fl_it}}, \code{\link{fl65_it}},\code{\link{fl_it}},
-#' \code{\link{hospbed_it}}, \code{\link{house_it}}, \code{\link{pm2.5_it}}
+#' @seealso for details regarding  the methodology of specific datasets check  \code{\link{it_bweight}}, \code{\link{it_cancer}},
+#' \code{\link{it_chronic}}, \code{\link{it_dem}}, \code{\link{it_firstaid}}, \code{\link{it_fl}}, \code{\link{it_fl65}},\code{\link{it_fl}},
+#' \code{\link{it_hospbed}}, \code{\link{it_house}}, \code{\link{it_pm2.5}}
 getit_all <- function() {
   cmr_it <- getit_covid()
 
-  house_it <- house_it %>%
+  it_house <- it_house %>%
     dplyr::filter(.data$year == 2018) %>%
     dplyr::select(.data$region, .data$phouse) %>%
     dplyr::rename("p_house" = "phouse")
 
   # get data and combine it together
-  dat1 <- plyr::join_all(list(cmr_it, house_it, fl_it_2019, dem_65bin_fm, regions_area), by = "region", type = "inner") %>%
+  dat1 <- plyr::join_all(list(cmr_it, it_house, it_fl_2019, dem_65bin_fm, it_regions), by = "region", type = "inner") %>%
     dplyr::mutate_at(vars("perc_imm65", "perc_imm"), as.numeric) %>%
     dplyr::mutate(pop_km2 = .data$pop_tot / .data$area_km2)
 
@@ -227,7 +227,7 @@ getit_all <- function() {
   )
 
   # inner join the rest
-  dat2 <- plyr::join_all(list(dat1[, col_order], chronic_it, cancer_it, smoking_it, bweight_it, firstaid_it, hospbed_it, netinc_it, pm2.5_it), by = "region", type = "inner")
[…]

### 2020-04-30 R/getus_functions.R

@@ -394,7 +394,7 @@ getus_all <- function() {
     dplyr::select(-.data$death, -.data$death_increase, -.data$abbr, -.data$hash, -.data$fips)
 
   # we keep only fips and vars
-  to_join <- lapply(list(acm_househ_us, age_sex_us, race_us, fl65_us, hospbeds_us, mmd_us, poverty_us, netinc_us, pm2.5_us, season_us), function(x) {
+  to_join <- lapply(list(us_acm_househ, us_age_sex, us_race, us_fl65, us_hospbeds, us_mmd, us_poverty, us_netinc, us_pm2.5, us_season), function(x) {
     x[, !names(x) %in% c("state_county", "county", "state", "year", "abbr")]
   })
 

### 2020-04-30 data-raw/import_raw.R

@@ -9,9 +9,9 @@ library(tabulizer)
 # xxxxxxxxxxxxxxxx-------------------
 # United States #-----
 # xxxxxxxxxxxxxxxx-------------------
-# age and sex- age_sex_us ---------
+# us_age_sex AKA us_dem  ---------
 
-age_sex_all_us <- vroom("data-raw/it_us/us/american_comunity_survey_2018/age_sex/ACSST5Y2018.S0101_data_with_overlays_2020-04-14T063202.csv",
+us_age_sex_all <- vroom("data-raw/it_us/us/american_comunity_survey_2018/age_sex/ACSST5Y2018.S0101_data_with_overlays_2020-04-14T063202.csv",
   skip = 1,
   col_select = c(`id`, `Geographic Area Name`, starts_with("Estimate"))
 ) %>%
@@ -20,20 +20,20 @@ age_sex_all_us <- vroom("data-raw/it_us/us/american_comunity_survey_2018/age_sex
 
 to_select_age_sex <- readxl::read_xlsx("data-raw/it_us/us/american_comunity_survey_2018/age_sex/to_select_age_sex.xlsx")
 
-age_sex_us <- age_sex_all_us %>%
+us_age_sex <- us_age_sex_all %>%
   select(!!to_select_age_sex$old_names)
 
 
-names(age_sex_us) <- to_select_age_sex$new_names
+names(us_age_sex) <- to_select_age_sex$new_names
 
-age_sex_us <- age_sex_us %>%
+us_age_sex <- us_age_sex %>%
   separate(id, into = c("us_id", "fips"), sep = 9) %>%
   select(-us_id) %>%
   mutate(fips = as.numeric(fips))
 
 
-# flu vax - fl65_us  ----------
-fl65_all_us <- vroom("data-raw/it_us/us/flu_us.csv",
+# us_fl65 ----------
+us_fl65_all <- vroom("data-raw/it_us/us/flu_us.csv",
   col_types = cols(
     year = col_double(),
     geography = col_character(),
@@ -55,12 +55,12 @@ fl65_all_us <- vroom("data-raw/it_us/us/flu_us.csv",
   )
 )
 
-fl65_us <- fl65_all_us %>%
+us_fl65<- us_fl65_all %>%
   rename("imm65" = "analysis_value") %>%
   dplyr::select(state, county, fips, imm65)
 
-# hospital beds - hospbeds_us  ----------
[…]

### 2020-04-30 tests/testthat/test-getit.R

@@ -8,7 +8,7 @@ test_that("getit_covid executes", {
 x <- getit_covid()
 
 # we take only the last year
-house_it <- house_it %>%
+it_house <- it_house %>%
   dplyr::filter(.data$year == 2018) %>%
   dplyr::select(.data$region, .data$phouse) %>%
   dplyr::rename("p_house" = "phouse")
@@ -16,17 +16,17 @@ house_it <- house_it %>%
 # check the datasets that we are merging-----
 t_data <- lapply(list(
   x,
-  cancer_it,
-  chronic_it,
+  it_cancer,
+  it_chronic,
   dem_65bin_fm,
-  bweight_it,
-  firstaid_it,
-  hospbed_it,
-  house_it,
-  netinc_it,
-  pm2.5_it,
-  regions_area,
-  smoking_it
+  it_bweight,
+  it_firstaid,
+  it_hospbed,
+  it_house,
+  it_netinc,
+  it_pm2.5,
+  it_regions,
+  it_smoking
 ), function(x) {
   unique(x$region) %in% unique(dem_65bin_fm$region)
 })

### 2020-04-30 data-raw/import_raw.R
na.rm mean

@@ -853,7 +853,7 @@ usethis::use_data(
 us_pm2.5 <-
   us_pm2.5 %>%
   group_by(fips) %>%
-  summarize(pm2.5 = mean(pm25)) %>%
+  summarize(pm2.5 = mean(pm25, na.rm = TRUE)) %>%
   ungroup()
 
 us_season <-
@@ -861,7 +861,7 @@ us_season <-
   group_by(fips) %>%
   # easy summarization of each variable...
   summarize_at(
-    vars(summer_temp:winter_hum), mean
+    vars(summer_temp:winter_hum), mean, na.rm = TRUE
     ) %>%
   ungroup()
 

### 2020-05-01 R/getus_functions.R
jhu covid

@@ -1,4 +1,79 @@
-#' get COVID-19 updated cases
+#' get COVID-19 from JHU
+#'
+#' extracts time series from the git repository of the \href{ https://github.com/CSSEGISandData }{JHU}
+#'
+#' @return a dataframe
+#' @importFrom rlang .data
+#' @importFrom magrittr %>%
+#' @import vroom
+#' @details `cases` represents the number of confirmed cases, while `cmr` the case-mortality rate (deaths / confirmed_case * 100).
+#' A good description of pitfalls and caveats associated with the use of case-mortality rate metric has been made on
+#' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
+#' @export
+getus_covid_jhu <- function() {
+
+  url_base <-  "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"
+
+  metric_files <- list("confirmed", "deaths")
+
+  # JHU has 2 files with date in long format (confirmed and deaths)...whatever
+  dat_l <- lapply(metric_files, function(x){
+    url_full <- paste0(url_base, x, "_US.csv")
+
+    if (RCurl::url.exists(url_data) == FALSE) {
+      stop("Something wrong with the repository or your internet connection!")
+    }
+
+   dat <- vroom(url_full, col_types = c(.default = "?" )) %>%
+      pivot_longer(cols = dplyr::matches("[0-9]{1,}/")  , names_to = "date", values_to = "value") %>%
+      janitor::clean_names() %>%
+      dplyr::filter(.data$country_region == "US") %>%
+      dplyr::select(
+        .data$date, .data$combined_key, .data$fips, .data$value
+      )  %>%
+     tidyr::separate( .data$combined_key,
+                      sep = ", ",
+                      into = c("county", "state", "country"),
+                      fill = "left"
+                      ) %>%
+     # JHU has unincorporated U.S territories and the cruises data
+     # that ends up to be NA because they have not counties in the dataframe
+     dplyr::filter(!is.na(.data$county), !is.na(.data$fips)) %>%
+     dplyr::select(-.data$country)
+
+   names(dat)[names(dat) == "value"] <- x
+
+   dat
+  }
+  )
+
+  names(dat_l) <- metric_files
+
[…]

### 2020-05-01 R/utilis.R

@@ -2,9 +2,10 @@
 #'
 #' recode_col recodes a vector with codes provided in another vector. The unique values of
 #' the 2 vectors need to be of same length. Codes are assigned matching unique values by position. Used internally.
-#' @param x vectore to recode.
+#' @param x vector to recode.
 #' @param repl vector containing unique codes.
 #' @return a vector of same length of x.
+#' @keywords internal
 recode_col <- function(x, repl) {
   intv_id <- as.character(unique(x))
   id_recode <- repl[seq_along(intv_id)]

### 2020-05-01 R/getus_functions.R
fixed warnings and testhat

@@ -9,7 +9,7 @@
 #' @details `cases` represents the number of confirmed cases, while `cmr` the case-mortality rate (deaths / confirmed_case * 100).
 #' A good description of pitfalls and caveats associated with the use of case-mortality rate metric has been made on
 #' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
-#' @export
+#' @keywords internal
 getus_covid_jhu <- function() {
 
   url_base <-  "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"
@@ -20,12 +20,12 @@ getus_covid_jhu <- function() {
   dat_l <- lapply(metric_files, function(x){
     url_full <- paste0(url_base, x, "_US.csv")
 
-    if (RCurl::url.exists(url_data) == FALSE) {
+    if (RCurl::url.exists(url_full) == FALSE) {
       stop("Something wrong with the repository or your internet connection!")
     }
 
-   dat <- vroom(url_full, col_types = c(.default = "?" )) %>%
-      pivot_longer(cols = dplyr::matches("[0-9]{1,}/")  , names_to = "date", values_to = "value") %>%
+   dat <- vroom(url_full, col_types = c(.default = "?" ), progress = FALSE) %>%
+      tidyr::pivot_longer(cols = dplyr::matches("[0-9]{1,}/")  , names_to = "date", values_to = "value") %>%
       janitor::clean_names() %>%
       dplyr::filter(.data$country_region == "US") %>%
       dplyr::select(
@@ -56,18 +56,15 @@ getus_covid_jhu <- function() {
   dat_w <-
     dat_l$confirmed %>%
     dplyr::select(.data$date, .data$fips, .data$confirmed) %>%
-    dplyr::inner_join(dat_l$confirmed, dat_l$deaths, by = c("date", "fips")) %>%
+    dplyr::inner_join(dat_l$deaths, by = c("date", "fips")) %>%
     dplyr::rename("cases"= "confirmed") %>%
     dplyr::mutate(
+      date = as.Date(.data$date, format = "%m/%d/%y"),
       cmr = .data$deaths / .data$cases * 100
       ) %>%
-  dplyr::mutate(date = format(strptime(.data$date, format = "%m/%d/%y"), "%Y-%m-%d"))
[…]

### 2020-05-01 tests/testthat/test-getus.R

@@ -1,19 +1,24 @@
 library(dplyr)
 
-test_that("get_all executes", {
-  expect_error(getus_all(), NA)
-})
+# test_that("get_all executes", {
+#   expect_error(getus_all(), NA)
+#   expect_error(getus_all(repo = "jhu"), NA)
+# })
+#
+# test_that("get_covid executes", {
+#   expect_error(getus_covid(), NA)
+#   expect_error(getus_covid(repo = "jhu"), NA)
+# })
 
-test_that("get_covid executes", {
-  expect_error(getus_covid(), NA)
-})
+y_nyt <- getus_all()
+y_jhu <- getus_all(repo = "jhu")
 
-y <- getus_all()
-
-x <- getus_covid()
+x_nyt <- getus_covid()
+x_jhu <- getus_covid(repo = "jhu")
 
 test_that("get_all and get_covid have same fips", {
-  expect_equal(length(unique(y$fips)), length(unique(x$fips)))
+  expect_equal(length(unique(y_nyt$fips)), length(unique(x_nyt$fips)))
+  expect_equal(length(unique(y_jhu$fips)), length(unique(x_jhu$fips)))
 })
 
 

### 2020-05-01 README.Rmd
added unknown fips to NYT

@@
 
-# to retrieve u.s data
-all_us <- getus_all()
+# to retrieve u.s data you can use 
+# the new york times or john hopkins university repository f
+all_us <- getus_all(repo = "jhu")
 
@@
 
-Data regarding COVID-19 number of cases and deaths (at the county level) were obtained from the [New York Time git repository](https://github.com/nytimes/covid-19-data), whereas number of tests and hospitalizations from the [the Covid Tracking Project](https://covidtracking.com/api}).
+Data regarding COVID-19 number of cases and deaths (at the county level) can be obtained from the [New York Time](https://github.com/nytimes/covid-19-data) or the [Johns Hopkins University](https://github.com/CSSEGISandData), whereas number of tests and hospitalizations from the [the Covid Tracking Project](https://covidtracking.com/api}).
 

### 2020-05-01 R/getus_functions.R

@@ -51,8 +51,8 @@ getus_covid_jhu <- function() {
 
   # diff_dat <- dat_w[which(dat_w$county.y != dat_w$county.x),]
 
-  # funny things is that there are unassigned county in one file (confirmed 90049), but
-  # not the other...whatever
+  # funny things is that there are unassigned county (confirmed 90049),
+  # that in the same file have a county
   dat_w <-
     dat_l$confirmed %>%
     dplyr::select(.data$date, .data$fips, .data$confirmed) %>%
@@ -101,8 +101,9 @@ getus_covid_nyt <- function() {
     )
   ) %>%
     dplyr::mutate(cmr = .data$deaths / .data$cases * 100) %>%
-    dplyr::filter(!is.na(.data$fips))
+    dplyr::filter(.data$state %in% state_abbr$state)
 
+  dat$fips[is.na(dat$fips)] <- 00000
 
   message(paste0("US COVID-19 data up to ", max(dat$date), " successfully retrived from NYT repository!"))
 
@@ -123,10 +124,10 @@ getus_covid_nyt <- function() {
 #' A good description of pitfalls and caveats associated with the use of case-mortality rate metric has been made on
 #' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
 #' @examples
-#' dat  <- getus_covid(repo = "nyt")
+#' dat  <- getus_covid(repo = "jhu")
 #' @export
 #'
-getus_covid <- function(repo = "nyt") {
+getus_covid <- function(repo = "jhu") {
         if (!repo %in% c("nyt", "jhu")) {
           stop("The argument repo can be only nyt or jhu")
         }
@@ -326,7 +327,10 @@ getus_tests <- function() {
 #'     \item{date}{formatted `ISO 8601`}
 #'     \item{county}{county}
 #'     \item{state}{state}
-#'     \item{fips}{federal information processing standard, a unique numeric identifier of a county}
+#'     \item{fips}{federal information processing standard, a unique numeric identifier of a county.
+#'         Unknown fips are coded as 00000. \strong{Note that in the nyt repository a lot of deaths
+#'         and confirmed cases are no categorized
+#'         at the county level}}
 #'     \item{urban}{urban or rural (not sure about the  definition)}
 #'     \item{\strong{COVID-19 VARS}}{---------------}
 #'     \item{cases}{confirmed COVID-19 cases (cumulates with date)}

### 2020-05-01 data-raw/import_raw.R

@@ -209,7 +209,7 @@ us_poverty <- poverty_all_us %>%
 
 # state abbreviations
 
-state_abbr <- vroom("data-raw/it_us/us/state_abbr.csv", col_types = cols(.default = "c"))
+state_abbr <- readxl::read_xlsx("data-raw/it_us/us/state_abbr.xlsx")
 
 # us_race ------------
 to_select_race <- readxl::read_xlsx("data-raw/it_us/us/american_comunity_survey_2018/race/to_select_race.xlsx")

### 2020-05-01 tests/testthat/test-getus.R

@@ -10,10 +10,10 @@ library(dplyr)
 #   expect_error(getus_covid(repo = "jhu"), NA)
 # })
 
-y_nyt <- getus_all()
+y_nyt <- getus_all(repo = "nyt")
 y_jhu <- getus_all(repo = "jhu")
 
-x_nyt <- getus_covid()
+x_nyt <- getus_covid(repo = "nyt")
 x_jhu <- getus_covid(repo = "jhu")
 
 test_that("get_all and get_covid have same fips", {
@@ -41,7 +41,7 @@ expected_states <- structure(list(state = c(
   "New Jersey", "New Mexico", "Nevada", "New York", "Ohio", "Oklahoma",
   "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
   "Tennessee", "Texas", "Utah", "Virginia", "Vermont", "Washington",
-  "Wisconsin", "West Virginia", "Wyomin"
+  "Wisconsin", "West Virginia", "Wyoming"
 ), abbr = c(
   "AK", "AL",
   "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI", "IA",

### 2020-05-01 R/datasets_us.R
styler and default argument

@@ -125,5 +125,3 @@
 #' @return  a tibble
 #' @usage data(us_season)
 "us_season"
-
-

### 2020-05-01 R/getus_functions.R

@@ -11,41 +11,39 @@
 #' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
 #' @keywords internal
 getus_covid_jhu <- function() {
-
-  url_base <-  "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"
+  url_base <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"
 
   metric_files <- list("confirmed", "deaths")
 
   # JHU has 2 files with date in long format (confirmed and deaths)...whatever
-  dat_l <- lapply(metric_files, function(x){
+  dat_l <- lapply(metric_files, function(x) {
     url_full <- paste0(url_base, x, "_US.csv")
 
     if (RCurl::url.exists(url_full) == FALSE) {
       stop("Something wrong with the repository or your internet connection!")
     }
 
-   dat <- vroom(url_full, col_types = c(.default = "?" ), progress = FALSE) %>%
-      tidyr::pivot_longer(cols = dplyr::matches("[0-9]{1,}/")  , names_to = "date", values_to = "value") %>%
+    dat <- vroom(url_full, col_types = c(.default = "?"), progress = FALSE) %>%
+      tidyr::pivot_longer(cols = dplyr::matches("[0-9]{1,}/"), names_to = "date", values_to = "value") %>%
       janitor::clean_names() %>%
       dplyr::filter(.data$country_region == "US") %>%
       dplyr::select(
         .data$date, .data$combined_key, .data$fips, .data$value
-      )  %>%
-     tidyr::separate( .data$combined_key,
-                      sep = ", ",
-                      into = c("county", "state", "country"),
-                      fill = "left"
-                      ) %>%
-     # JHU has unincorporated U.S territories and the cruises data
-     # that ends up to be NA because they have not counties in the dataframe
-     dplyr::filter(!is.na(.data$county), !is.na(.data$fips)) %>%
-     dplyr::select(-.data$country)
+      ) %>%
+      tidyr::separate(.data$combined_key,
+        sep = ", ",
+        into = c("county", "state", "country"),
+        fill = "left"
+      ) %>%
[…]

### 2020-05-01 data-raw/import_raw.R

@@ -55,7 +55,7 @@ us_fl65_all <- vroom("data-raw/it_us/us/flu_us.csv",
   )
 )
 
-us_fl65<- us_fl65_all %>%
+us_fl65 <- us_fl65_all %>%
   rename("imm65" = "analysis_value") %>%
   dplyr::select(state, county, fips, imm65)
 
@@ -246,15 +246,15 @@ us_pm2.5 <- vroom("https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Da
 # The Atmospheric Composition Analysis Group at Dalhouse University
 
 us_season <- vroom("https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Data/temp_seasonal_county.csv",
-                    col_types = cols(.default = "d")
-                  )
+  col_types = cols(.default = "d")
+)
 
 names(us_season) <- c("fips", "year", "summer_temp", "summer_hum", "winter_temp", "winter_hum")
 
 # us_netinc -----
 # from census
 netincome_all_us <- vroom("data-raw/it_us/us/american_comunity_survey_2018/netincome/ACSST5Y2018.S1901_data_with_overlays_2020-04-29T001634.csv",
-                     col_types = cols(.default = "c"), skip = 1
+  col_types = cols(.default = "c"), skip = 1
 ) %>%
   clean_names() %>%
   select(id, geographic_area_name, estimate_households_median_income_dollars)
@@ -861,8 +861,9 @@ us_season <-
   group_by(fips) %>%
   # easy summarization of each variable...
   summarize_at(
-    vars(summer_temp:winter_hum), mean, na.rm = TRUE
-    ) %>%
+    vars(summer_temp:winter_hum), mean,
+    na.rm = TRUE
+  ) %>%
   ungroup()
 
 

### 2020-05-05 R/getit_functions.R
added analysis draft

@@ -25,7 +25,7 @@
 #'      \item{people_tested}{number of people tested}
 #' }
 #' @details caveats and problems related the calculation by the Protezione Civile of some variables  were rised by
-#' \href{https://www.gimbe.org/pagine/341/it/comunicati-stampa?pagina=2}{GIMBE fFoundation}. Unfortunately the page is in Italian...
+#' \href{https://www.gimbe.org/pagine/341/it/comunicati-stampa?pagina=2}{GIMBE Foundation}. Unfortunately the page is in Italian...
 #' \emph{ buona lettura! }
 #' @importFrom rlang .data
 #' @importFrom magrittr %>%

### 2020-05-07 README.Rmd
first commit

@@
 The other datsets were obtained from the [Homeland Infrastructure Foundation](https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18), the [Census](https://data.census.gov/cedsci/table?q=United%20States), [Mapping Medicare Disparities](https://data.cms.gov/mapping-medicare-disparities), and [activity indexes](https://github.com/COVIDExposureIndices/COVIDExposureIndices) calculated by  Victor Couture, Jonathan Dingel, Allison Green, Jessie Handbury, and Kevin Williams based on smartphone movement data provided by `PlaceIQ`.
-Emissions of pm2.5 in micro g/m3 (2016) were reported by \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group} and
-#'  aggregate by \href{https://github.com/wxwx1993/PM_COVID/blob/master/additional_preprocessing_code/download_pm25_values.md}{Ista Zahn and Ben Sabath}
+Emissions of pm2.5 in micro g/m3 (2016) were reported by \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group} and aggregate by [Ista Zahn and Ben Sabath.](https://github.com/wxwx1993/PM_COVID/blob/master/additional_preprocessing_code/download_pm25_values.md)
 
@@
 * `us_age_sex`: age and sex distribution (2018).
-* `us_fl65`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
 * `us_hospbeds`: total hospital beds in each county (2019).
@@
     * emergency,  medical admissions and annual visits.
-    * pneumoccocal vaccine.
     * tobacco use.
@@
 
-Demographic metrics and other stats regarding medical conditions and vaccinations were obtained from the [ISTAT](http://dati.istat.it/?lang=en), [Ministero della Salute](http://www.salute.gov.it/portale/home.html).
+Demographic metrics and other stats regarding medical conditions were obtained from the [ISTAT](http://dati.istat.it/?lang=en), [Ministero della Salute](http://www.salute.gov.it/portale/home.html).
 Data regarding `pm2.5` were collected by [Istituto Superiore Per La protexione Ambientale](https://annuario.isprambiente.it/pon/basic/14)
@@
 * `it_firstaid`: Number of people using first aid or medical guard (3 months). 
-* `it_fl`: Influenza vaccination coverage in Italy for 2009-2019 season for general population and aged 65 and more.
-* `it_fl_2019`: Influenza vaccination coverage in Italy for 2018-2019 season for general population and 65 and more.
 * `it_hospbed`: Inpatient hospital beds per 1000 people.

### 2020-05-07 R/datasets_it.R

@@ -1,0 +1,116 @@
+#' body-mass index
+#'
+#' Body mass index in regions of Italy, in the general population.
+#' Data were collected in 2018 and indicate absolute number of people underweight, normalweight, overweight or obese.
+#' @docType data
+#' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
+#' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
+#' @usage data(it_bweight)
+"it_bweight"
+
+#' cancer patients
+#'
+#' Number of cancer patients in each region by type.
+#' Data were collected in 2016 and indicate absolute number of people diagnosed with cancer. Data for P.A. Trento and
+#' P.A. Bolzano are missing (but we have Trentino Alto Adige)
+#' @docType data
+#' @source \href{http://www.registri-tumori.it/PDF/AIOM2016/I_numeri_del_cancro_2016.pdf}{Istituto Superiore Sanita'}
+#' @usage data(it_cancer)
+#' @return a tibble
+"it_cancer"
+
+#' smoking status
+#'
+#' Number of people age 14 years and over that self-refer as smoker, non smoker, or past smoker by region and type.
+#' Data were collected in 2018 and are absolute number of people.
+#' @docType data
+#' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
+#' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
+#' @usage data(it_smoking)
+#' @return a tibble
+"it_smoking"
+
+#' chronic conditions
+#'
+#' Number of people suffering of chronic conditions by region and type.
+#' Data were collected in 2018 and indicate absolute number of people.
+#' @docType data
+#' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
+#' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=0058000&refresh=true&language=EN}{methodology}
+#' @usage data(it_chronic)
+#' @return a tibble
+"it_chronic"
+
+
+#' Percent of population by region, sex and age. Data were collected in 2019 and indicate absolute number of people. Long format,
+#  age by year.
+#' @docType data
+#' @source \href{http://dati.istat.it/?lang=en}{ISTAT}
[…]

### 2020-05-07 R/datasets_us.R

@@ -1,0 +1,117 @@
+#' hospital beds
+#'
+#' beds of each hospital by county (2019).
+#' @docType data
+#' @source \href{https://hifld-geoplatform.opendata.arcgis.com/datasets/hospitals/data?page=18}{Homeland Infrastructure Foundation-Level Data}
+#' @return  a tibble
+#' @usage data(us_hospbeds)
+"us_hospbeds"
+
+#' household composition
+#'
+#' Several metrics regarding household composition from the American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions}
+#' @return  a tibble
+#' @usage data(us_acm_househ)
+"us_acm_househ"
+
+#' age and sex
+#'
+#' Sex and age composition of the county population from the American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @return  a tibble
+#' @usage data(us_dem)
+"us_dem"
+
+#' poverty
+#'
+#' Household living below the poverty level, divided by age and race and calculate as absolute value or percentage. American Community Survey of 2018
+#' @docType data
+#' @source \href{https://data.census.gov/cedsci/table?q=United%20States}{American Community Survey tables}
+#' @details \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey}
+#' @return  a tibble
+#' @usage data(us_poverty)
+"us_poverty"
+
+
+#' mapping medicare disparities
+#'
+#' Prevalence of many  medical and chronic conditions, 2017. From relative documentation listed below: "Prevalence rates are calculated by searching for certain diagnosis codes in Medicare
+#' beneficiaries’ claims. The prevalence rate of a condition for a specific sub-population
[…]

### 2020-05-07 R/getit_functions.R

@@ -1,0 +1,239 @@
+#' get COVID-19 updated cases
+#'
+#' extracts and translates time series form the git repository of the \href{https://github.com/pcm-dpc/COVID-19}{protezione civile}
+#'
+#' @return a dataframe with following 19 variables:
+#' \describe{
+#'      \item{date}{in `ISO 8601` format}
+#'      \item{state}{state}
+#'      \item{region_code}{region abbreviation}
+#'      \item{region}{full name of region}
+#'      \item{lat}{lat}
+#'      \item{long}{long}
+#'      \item{cmr}{case-mortality rate for that region and that date (deaths/total_cases * 100)}
+#'      \item{total_cases}{number of COVID-19 positive cases detected}
+#'      \item{deaths}{number of deaths}
+#'      \item{tests}{number of tests performed}
+#'      \item{hospitalized_with_symptoms}{number of people hospitalized with symptoms, that day}
+#'      \item{intensive_care_unit}{number of people in intensive care units, that day}
+#'      \item{total_hospitalized}{hospitalized_with_symptoms + intensive_care_unit}
+#'      \item{home_quarantine}{number of people COVID-19 positive in home quarantine, that day}
+#'      \item{total_positives}{total currently positives: hospitalized_with_symptoms + intensive_care_unit + home_quarantine}
+#'      \item{change_positives}{change in the number of positive cases: total_positives that day - total_positives preceding day}
+#'      \item{new_positives}{number of new positive cases: total_cases that day - total_cases preceding day}
+#'      \item{recovered_released}{recovered - released from hospital}
+#'      \item{people_tested}{number of people tested}
+#' }
+#' @details caveats and problems related the calculation by the Protezione Civile of some variables  were rised by
+#' \href{https://www.gimbe.org/pagine/341/it/comunicati-stampa?pagina=2}{GIMBE Foundation}. Unfortunately the page is in Italian...
+#' \emph{ buona lettura! }
+#' @importFrom rlang .data
+#' @importFrom magrittr %>%
+#' @import vroom
+#' @export
+getit_covid <- function() {
+  url_data <- "https://raw.githubusercontent.com/pcm-dpc/COVID-19/master/dati-regioni/dpc-covid19-ita-regioni.csv"
+
+  if (RCurl::url.exists(url_data) == FALSE) {
+    stop("Something wrong with the repository or your internet connection!")
+  }
+
+  dat <- vroom(
[…]

### 2020-05-07 R/getus_functions.R

@@ -1,0 +1,518 @@
+#' get COVID-19 from JHU
+#'
+#' extracts time series from the git repository of the \href{ https://github.com/CSSEGISandData }{JHU}
+#'
+#' @return a dataframe
+#' @importFrom rlang .data
+#' @importFrom magrittr %>%
+#' @import vroom
+#' @details `cases` represents the number of confirmed cases, while `cmr` the case-mortality rate (deaths / confirmed_case * 100).
+#' A good description of pitfalls and caveats associated with the use of case-mortality rate metric has been made on
+#' \href{ https://ourworldindata.org/covid-mortality-risk }{Our World in Data}.
+#' @keywords internal
+getus_covid_jhu <- function() {
+  url_base <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"
+
+  metric_files <- list("confirmed", "deaths")
+
+  # JHU has 2 files with date in long format (confirmed and deaths)...whatever
+  dat_l <- lapply(metric_files, function(x) {
+    url_full <- paste0(url_base, x, "_US.csv")
+
+    if (RCurl::url.exists(url_full) == FALSE) {
+      stop("Something wrong with the repository or your internet connection!")
+    }
+
+    dat <- vroom(url_full, col_types = c(.default = "?"), progress = FALSE) %>%
+      tidyr::pivot_longer(cols = dplyr::matches("[0-9]{1,}/"), names_to = "date", values_to = "value") %>%
+      janitor::clean_names() %>%
+      dplyr::filter(.data$country_region == "US") %>%
+      dplyr::select(
+        .data$date, .data$combined_key, .data$fips, .data$value
+      ) %>%
+      tidyr::separate(.data$combined_key,
+        sep = ", ",
+        into = c("county", "state", "country"),
+        fill = "left"
+      ) %>%
+      # JHU has unincorporated U.S territories and the cruises data
+      # that ends up to be NA because they have not counties in the dataframe
+      dplyr::filter(!is.na(.data$county), !is.na(.data$fips)) %>%
+      dplyr::select(-.data$country)
+
+    names(dat)[names(dat) == "value"] <- x
+
+    dat
+  })
+
+  names(dat_l) <- metric_files
+
+  # diff_dat <- dat_w[which(dat_w$county.y != dat_w$county.x),]
+
[…]

### 2020-05-07 R/utilis.R

@@ -1,0 +1,14 @@
+#' recode column.
+#'
+#' recode_col recodes a vector with codes provided in another vector. The unique values of
+#' the 2 vectors need to be of same length. Codes are assigned matching unique values by position. Used internally.
+#' @param x vector to recode.
+#' @param repl vector containing unique codes.
+#' @return a vector of same length of x.
+#' @keywords internal
+recode_col <- function(x, repl) {
+  intv_id <- as.character(unique(x))
+  id_recode <- repl[seq_along(intv_id)]
+  newnames <- stats::setNames(id_recode, intv_id)
+  as.vector(newnames[x])
+}

### 2020-05-07 data-raw/import_raw.R

@@ -1,0 +1,845 @@
+library(vroom)
+library(tidyverse)
+library(janitor)
+library(reshape2)
+library(tabulizer)
+
+# Unpackit_us.zip and run the script
+
+# xxxxxxxxxxxxxxxx-------------------
+# United States #-----
+# xxxxxxxxxxxxxxxx-------------------
+# us_age_sex AKA us_dem  ---------
+
+us_age_sex_all <- vroom("data-raw/it_us/us/american_comunity_survey_2018/age_sex/ACSST5Y2018.S0101_data_with_overlays_2020-04-14T063202.csv",
+  skip = 1,
+  col_select = c(`id`, `Geographic Area Name`, starts_with("Estimate"))
+) %>%
+  clean_names()
+
+
+to_select_age_sex <- readxl::read_xlsx("data-raw/it_us/us/american_comunity_survey_2018/age_sex/to_select_age_sex.xlsx")
+
+us_age_sex <- us_age_sex_all %>%
+  select(!!to_select_age_sex$old_names)
+
+
+names(us_age_sex) <- to_select_age_sex$new_names
+
+us_age_sex <- us_age_sex %>%
+  separate(id, into = c("us_id", "fips"), sep = 9) %>%
+  select(-us_id) %>%
+  mutate(fips = as.numeric(fips))
+
+
+
+# us_hospbeds  ----------
+us_hospbeds_all <- vroom("data-raw/it_us/us/hospital_beds.csv",
+  col_types = cols(
+    X = col_double(),
+    Y = col_double(),
+    OBJECTID = col_double(),
+    ID = col_character(),
+    NAME = col_character(),
+    ADDRESS = col_character(),
+    CITY = col_character(),
+    STATE = col_character(),
+    ZIP = col_character(),
+    ZIP4 = col_character(),
+    TELEPHONE = col_character(),
+    TYPE = col_character(),
+    STATUS = col_character(),
+    POPULATION = col_double(),
+    COUNTY = col_character(),
+    COUNTYFIPS = col_character(),
+    COUNTRY = col_character(),
+    LATITUDE = col_double(),
+    LONGITUDE = col_double(),
+    NAICS_CODE = col_double(),
+    NAICS_DESC = col_character(),
+    SOURCE = col_character(),
+    SOURCEDATE = col_datetime(format = ""),
+    VAL_METHOD = col_character(),
+    VAL_DATE = col_datetime(format = ""),
+    WEBSITE = col_character(),
+    STATE_ID = col_character(),
[…]

### 2020-05-07 inst/img/make_hex.R

@@ -1,0 +1,56 @@
+library(devtools)
+
+# devtools::install_github("c1au6i0/covid19census")
+library(kableExtra)
+library(covid19census)
+library(dplyr)
+library(showtext)
+library(hexSticker)
+library(ggplot2)
+library(png)
+library(grid)
+
+tab2 <- getus_all()
+
+tab2 <- ff[1:50, 1:50]
+
+
+names(tab2) <- paste0("x", 1:50)
+
+
+tab2 %>%
+  kable(escape = F, align = "c") %>%
+  column_spec(1:50, bold = T, border_right = T) %>%
+  kable_styling(
+    latex_options = c("striped", "scale_down"),
+    full_width = FALSE,
+    font_size = 5
+  )
+
+
+
+tab1 <- getit_all()
+glimpse(tab1)
+
+imgurl <- readPNG("inst/img/image_logo.png")
+imgurl <- rasterGrob(imgurl, interpolate = TRUE)
+
+font_add_google(name = "Quantico", "gira")
+showtext_auto()
+
+sticker(
+  imgurl,
+  package = "covid19census",
+  p_size = 5,
+  p_y = 1.5,
+  s_x = 1.1,
+  s_y = 0.7,
+  s_width = 2.3,
+  s_height = 2.3,
+  p_color = "#b83335",
+  p_family = "gira",
+  filename = "inst/img/hexsticker2.png",
+  h_color = "#b83335",
+  h_fill = "white"
+)
+

### 2020-05-07 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(covid19census)
+
+test_check("covid19census")

### 2020-05-07 tests/testthat/test-getit.R

@@ -1,0 +1,61 @@
+library(stringr)
+
+# getit_covid --------------------
+test_that("getit_covid executes", {
+  expect_error(getit_covid(), NA)
+})
+
+x <- getit_covid()
+
+# we take only the last year
+it_house <- it_house %>%
+  dplyr::filter(.data$year == 2018) %>%
+  dplyr::select(.data$region, .data$phouse) %>%
+  dplyr::rename("p_house" = "phouse")
+
+# check the datasets that we are merging-----
+t_data <- lapply(list(
+  x,
+  it_cancer,
+  it_chronic,
+  dem_65bin_fm,
+  it_bweight,
+  it_firstaid,
+  it_hospbed,
+  it_house,
+  it_netinc,
+  it_pm2.5,
+  it_regions,
+  it_smoking
+), function(x) {
+  unique(x$region) %in% unique(dem_65bin_fm$region)
+})
+
+
+test_that("all datasets and function have same regions", {
+  expect_equal(sum(unlist(lapply(t_data, sum))), 252)
+})
+
+
+test_that("get_all executes", {
+  expect_error(getit_all(), NA)
+})
+
+y <- getit_all()
+
+test_that("get_all returns regions_21", {
+  expect_equal(length(unique(y$region)), 21)
+})
+
+
+test_that("sum smoking is < 100 ", {
+  expect_equal(
+    sum(!unique(
+      apply(
+        y[, names(y)[str_detect(names(y), "smoking")]],
+        1, sum
+      )
+    ) <= 100),
+    0
+  )
+})

### 2020-05-07 tests/testthat/test-getus.R

@@ -1,0 +1,64 @@
+library(dplyr)
+
+# test_that("get_all executes", {
+#   expect_error(getus_all(), NA)
+#   expect_error(getus_all(repo = "jhu"), NA)
+# })
+#
+# test_that("get_covid executes", {
+#   expect_error(getus_covid(), NA)
+#   expect_error(getus_covid(repo = "jhu"), NA)
+# })
+
+y_nyt <- getus_all(repo = "nyt")
+y_jhu <- getus_all(repo = "jhu")
+
+x_nyt <- getus_covid(repo = "nyt")
+x_jhu <- getus_covid(repo = "jhu")
+
+test_that("get_all and get_covid have same fips", {
+  expect_equal(length(unique(y_nyt$fips)), length(unique(x_nyt$fips)))
+  expect_equal(length(unique(y_jhu$fips)), length(unique(x_jhu$fips)))
+})
+
+
+###################################
+
+test_that("get_dex executes", {
+  expect_error(getus_dex(), NA)
+})
+
+##################################
+
+
+expected_states <- structure(list(state = c(
+  "Alaska", "Alabama", "Arizona", "Arkansas",
+  "California", "Colorado", "Connecticut", "District of Columbia",
+  "Delaware", "Florida", "Georgia", "Hawaii", "Iowa", "Idaho",
+  "Illinois", "Indiana", "Kansas", "Kentucky", "Louisiana", "Massachusetts",
+  "Maryland", "Maine", "Michigan", "Minnesota", "Missouri", "Mississippi",
+  "Montana", "North Carolina", "North Dakota", "Nebraska", "New Hampshire",
+  "New Jersey", "New Mexico", "Nevada", "New York", "Ohio", "Oklahoma",
+  "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
+  "Tennessee", "Texas", "Utah", "Virginia", "Vermont", "Washington",
+  "Wisconsin", "West Virginia", "Wyoming"
+), abbr = c(
+  "AK", "AL",
+  "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI", "IA",
+  "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN",
+  "MO", "MS", "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY",
[…]

### 2020-05-07 README.Rmd
dev version, draft, readme
readme

--- 2020-05-01 README.Rmd
@@
 [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
-[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
+[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census_dev)
 [![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
@@
 
-devtools::install_github("c1au6i0/covid19census")
+devtools::install_github("c1au6i0/covid19census_dev")
 ```

### 2020-05-11 R/getus_functions.R
added age summary indicators

@@ -308,7 +308,7 @@ getus_tests <- function() {
 #'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices},
 #'  \href{http://fizz.phys.dal.ca/~atmos/martin/?page_id=140#V4.NA.02.MAPLE}{Atmoshpheric Composition Analysis Group}
 #' @param repo repository of COVID-19 data, one of `c("nyt", "jhu")`
-#' @return A dataframe with 330 variables. Data regarding the household composition, population sex, age, race, ancestry and poverty levels,
+#' @return A dataframe. Data regarding the household composition, population sex, age, race, ancestry and poverty levels,
 #'  were scraped from the 2018 American Community Survey (ACS). Poverty was defined at the family level and not the household level in
 #'  the ACS. Medical conditions, tobacco use, cancer and, data relative to the number of medical and emergency visits
 #'  was obtained from the 2017 Mapping Medicare Disparities. From relative documentation listed in the source: "Prevalence rates are calculated
@@ -394,6 +394,14 @@ getus_tests <- function() {
 #'     \item{total_female}{total female}
 #'     \item{total_ \emph{age_sex}}{total population by age bin and sex}
 #'     \item{perc_ \emph{age_sex}}{percent population by age bin and sex}
+#'     \item{median_age}{median age in years}
+#'     \item{median_age_male}{median age in years of males}
+#'     \item{median_age_female}{median age in years of females}
+#'     \item{sex_ratio}{males per 100 females}
+#'     \item{age_dependency}{the age dependency ratio is derived by dividing the combined under
+#'         18 years and 65 years and over populations by the 18-to-64 population and multiplying by 100}
+#'     \item{old_age_dependency}{the old-age dependency ratio is derived by dividing the population 65 years and over by the 18-to-64 population and multiplying by 100}
+#'     \item{child_dependency}{the child dependency ratio is derived by dividing the population under 18 years by the 18-to-64 population, and multiplying by 100}
 #'     \item{\strong{POPULATION AND RACE}}{---------------}
 #'     \item{total_nlat}{total not hispanic or latinos. It can be normalized using total_population}
 #'     \item{total_nlat_white_alone}{total white in that category}

### 2020-05-11 README.Rmd
readme
token
clean other conflicts, added rdaS

@@
 
-devtools::install_github("c1au6i0/covid19census_dev")
+devtools::install_github("c1au6i0/covid19census_dev", auth_token = "8ca5cd2bb556a45f1ffeb2e9b44c90d4fbef6cf2")
 ```
@@
 Demographic metrics and other stats regarding medical conditions and vaccinations were obtained from the [ISTAT](http://dati.istat.it/?lang=en), [Ministero della Salute](http://www.salute.gov.it/portale/home.html).
-Data regarding `pm2.5` were collected by [Istituto Superiore Per La protexione Ambientale](https://annuario.isprambiente.it/pon/basic/14)
+Data regarding `pm2.5` were collected by [Istituto Superiore per la Protezione Ambientale](https://annuario.isprambiente.it/pon/basic/14)
 

### 2020-05-11 R/datasets_us.R

@@ -1,4 +1,3 @@
-<<<<<<< HEAD
 #' influenza vaccination 65 or older
 #'
 #' Percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination. Collected in 2017.
@@ -9,8 +8,6 @@
 #' @usage data(us_fl65)
 "us_fl65"
 
-=======
->>>>>>> public
 #' hospital beds
 #'
 #' beds of each hospital by county (2019).

### 2020-05-19 R/getus_functions.R
fixed latinos

@@ -403,14 +403,14 @@ getus_tests <- function() {
 #'     \item{old_age_dependency}{the old-age dependency ratio is derived by dividing the population 65 years and over by the 18-to-64 population and multiplying by 100}
 #'     \item{child_dependency}{the child dependency ratio is derived by dividing the population under 18 years by the 18-to-64 population, and multiplying by 100}
 #'     \item{\strong{POPULATION AND RACE}}{---------------}
-#'     \item{white}{total white}
-#'     \item{black_afroamerican}{total black or afroamerican}
-#'     \item{native}{total native}
-#'     \item{asian}{total asian}
-#'     \item{pacific_islander}{total hawaian and pacific islander}
-#'     \item{other_race}{other races}
-#'     \item{two_more_races}{two or more races}
-#'     \item{latino}{total hispanic or latinos}
+#'     \item{total_white}{total white}
+#'     \item{total_black}{total black or afroamerican}
+#'     \item{total_native}{total native}
+#'     \item{total_asian}{total asian}
+#'     \item{total_pacific_islander}{total hawaian and pacific islander}
+#'     \item{total_other_race}{other races}
+#'     \item{total_two_more_races}{two or more races}
+#'     \item{total_latino}{total hispanic or latino}
 #'     \item{\strong{MEDICAL AND VACCINES}}{---------------}
 #'     \item{perc_imm65}{percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination.}
 #'     \item{total_beds}{total number of hospital beds}

### 2020-05-28 README.Rmd
readme

@@
 [![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
-[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census_dev)
+[![Travis build status](https://travis-ci.com/c1au6i0/covid19census_dev.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=dev)
 [![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
@@
 
-devtools::install_github("c1au6i0/covid19census_dev", auth_token = "8ca5cd2bb556a45f1ffeb2e9b44c90d4fbef6cf2")
+devtools::install_github("c1au6i0/covid19census_dev", auth_token = "8ca5cd2bb556a45f1ffeb2e9b44c90d4fbef6cf2", ref = "dev")
 ```

### 2020-05-28 R/datasets_it.R

@@ -42,29 +42,6 @@
 "it_chronic"
 
 
-#' influenza vaccination coverage, general population, time series
-#'
-#' Influenza vaccination coverage in Italy in the \strong{general population} from 1999 to 2019. Data are percent of region population
-#' @docType data
-#' @source \href{http://www.salute.gov.it/imgs/C_17_tavole_19_allegati_iitemAllegati_0_fileAllegati_itemFile_3_file.pdf}{Ministero della Salute}
-#' @usage data(it_fl)
-"it_fl"
-
-#' influenza vaccination coverage 2019
-#'
-#' Influenza vaccination coverage in Italy for 2018-2019 season for population age 65 or more from 1999 to 2019. Data are percent of region population
-#' @docType data
-#' @source \href{http://www.salute.gov.it/imgs/C_17_tavole_19_allegati_iitemAllegati_0_fileAllegati_itemFile_3_file.pdf}{Ministero della Salute}
-#' @usage data(it_fl65)
-#' @return a tibble with following columns:
-#' \describe{
-#'      \item{\strong{region}}{region}
-#'      \item{\strong{perc_imm65}}{percent of population age 65 or more that received influenza vaccination}
-#'      \item{\strong{perc}}{percent of general population that received influenza vaccination}
-#'      }
-"it_fl65"
-
-
 #' Percent of population by region, sex and age. Data were collected in 2019 and indicate absolute number of people. Long format,
 #  age by year.
 #' @docType data

### 2020-06-08 R/datasets_it.R
fixed bug cancer

@@ -1,3 +1,25 @@
+#' influenza vaccination coverage, general population, time series
+#'
+#' Influenza vaccination coverage in Italy in the \strong{general population} from 1999 to 2019. Data are percent of region population
+#' @docType data
+#' @source \href{http://www.salute.gov.it/imgs/C_17_tavole_19_allegati_iitemAllegati_0_fileAllegati_itemFile_3_file.pdf}{Ministero della Salute}
+#' @usage data(it_fl)
+"it_fl"
+
+#' influenza vaccination coverage 2019
+#'
+#' Influenza vaccination coverage in Italy for 2018-2019 season for population age 65 or more from 1999 to 2019. Data are percent of region population
+#' @docType data
+#' @source \href{http://www.salute.gov.it/imgs/C_17_tavole_19_allegati_iitemAllegati_0_fileAllegati_itemFile_3_file.pdf}{Ministero della Salute}
+#' @usage data(it_fl65)
+#' @return a tibble with following columns:
+#' \describe{
+#'      \item{region}{region}
+#'      \item{perc_imm65}{percent of population age 65 or more that received influenza vaccination}
+#'      \item{perc}{percent of general population that received influenza vaccination}
+#'      }
+"it_fl65"
+
 #' body-mass index
 #'
 #' Body mass index in regions of Italy, in the general population.

### 2020-06-08 R/getit_functions.R

@@ -183,7 +183,7 @@ getit_covid <- function() {
 #' @importFrom stats na.omit
 #' @export
 #' @seealso for details regarding  the methodology of specific datasets check  \code{\link{it_bweight}}, \code{\link{it_cancer}},
-#' \code{\link{it_chronic}}, \code{\link{it_dem}}, \code{\link{it_firstaid}}, \code{\link{it_fl}}, \code{\link{it_fl65}},\code{\link{it_fl}},
+#' \code{\link{it_chronic}}, \code{\link{it_dem}}, \code{\link{it_firstaid}}, \code{\link{it_fl}}, \code{\link{it_fl65}},
 #' \code{\link{it_hospbed}}, \code{\link{it_house}}, \code{\link{it_pm2.5}}
 getit_all <- function() {
   cmr_it <- getit_covid()
@@ -234,10 +234,40 @@ getit_all <- function() {
     all_dat <- dat2 %>%
       dplyr::mutate_at(vars(.data$chronic_osteo:.data$bweight_obese), list(~ (. / .data$pop_tot * 100))) %>%
       dplyr::rename_at(vars(.data$chronic_osteo:.data$bweight_obese), ~ paste0("perc_", .)) %>%
-      dplyr::rename(cases = total_cases, total_bed = bed_tot, total_tests = people_tested) %>%
+      dplyr::rename("cases" = .data$total_cases,  "total_bed" = .data$bed_tot, "total_tests" = .data$people_tested) %>%
       dplyr::mutate(date = as.Date(.data$date, "%Y-%m-%d"))
   )
 
 
   all_dat
 }
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2020-06-08 data-raw/import_raw.R

@@ -3,6 +3,7 @@ library(tidyverse)
 library(janitor)
 library(reshape2)
 library(tabulizer)
+library(stringr)
 
 # Unpackit_us.zip and run the script
 
@@ -427,7 +428,7 @@ it_cancer <- cancer_tab %>%
   )) %>%
   filter(region != "Trentino Alto Adige") %>%
   # in italy . used as , in numbers
-  mutate_if(is.double, ~ . * 1000)
+  mutate_if(is.double, ~ as.numeric(str_remove(as.character(.), "[[:punct:]]")))
 
 # no data on those  but we want to include them anyway for the join
 it_cancer[20, "region"] <- "P.A. Bolzano"
@@ -908,6 +909,12 @@ usethis::use_data(
   overwrite = TRUE
 )
 #
+
+
+
+
+
+
 
 
 # they take the mean of all the different years, we do the same -----

### 2020-06-08 README.Rmd
readme and removed draft

@@
 <!-- badges: start -->
-[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
+[![](https://img.shields.io/badge/devel%20version-0.0.0.9600-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=dev)
 [![Travis build status](https://travis-ci.com/c1au6i0/covid19census_dev.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=dev)
-[![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=dev)
 <!-- badges: end -->
@@
 ```{r intallation2, eval=FALSE, include=TRUE}
-# the repo at the moment private so you need to be authorized 
+# the repo at the moment is private so you need to be authorized 
 # don't share the token

### 2020-06-09 R/getit_functions.R
changed docum

@@ -158,8 +158,11 @@ getit_covid <- function() {
 #'      \item{female_65m}{percent of females age 65 years old or more}
 #'      \item{male_65m}{percent of males age 65 years old or more}
 #'      \item{chronic_ \emph{type}}{percent of population with that chronic condistion}
-#'      \item{perc_cancer_\emph{type}}{percent of population with that type of cancer}
-#'      \item{perc_bweight_\emph{type}}{percent of people underweight, normalweight, overweight or obese}
+#'      \item{perc_cancer_\emph{type}}{percent of population with that type of cancer.
+#'          Info regarding Trento and Bolzano were not present.}
+#'      \item{perc_bweight_\emph{type}}{percent of people underweight, normalweight, overweight or obese.
+#'          This is percent calculated over the total population even if the mesure has been taken only people 18 of age or more.
+#'          This is the reason why their total is not 100}
 #'      \item{first_aid}{number of peple using first aid in 3 months preceding the survey}
 #'      \item{medical_guard}{number of people using medical guard in 3 months preceding the survey}
 #'      \item{bed_acute}{inpatient hospital beds per 1000 people in acure care}

### 2020-06-12 R/getus_functions.R
docum urban

@@ -337,7 +337,7 @@ getus_tests <- function() {
 #'         Unknown fips are coded as 00000. \strong{Note that in the nyt repository a lot of deaths
 #'         and confirmed cases are no categorized
 #'         at the county level}}
-#'     \item{urban}{urban or rural (not sure about the  definition)}
+#'     \item{urban}{urban or rural (see \href{https://www.census.gov/programs-surveys/geography/guidance/geo-areas/urban-rural.html}{cenus})}
 #'     \item{\strong{COVID-19 VARS}}{---------------}
 #'     \item{cases}{confirmed COVID-19 cases (cumulates with date)}
 #'     \item{deaths}{number of deaths attributed to COVID-19}

### 2020-06-12 README.Rmd
changed readme

@@
 The package `covid19census` provides functions to extract COVID-19 cases, deaths, hospitalizations and test of U.S.
-and Italy at the county and regional level, respectively, and then combine them with other population metrics (age, sex, prevalence of chronic conditions, income indexes, access to health services and many others).
+and Italy at the county and regional level, respectively, and then combines them with other population metrics (age, sex, prevalence of chronic conditions, income indexes, access to health services and many others).
+
+*Note that the package has not been made by, and is not part of the covid19.cesus.gov website.*
 

### 2020-06-12 README.Rmd
readme

@@
 <!-- badges: start -->
-[![](https://img.shields.io/badge/devel%20version-0.0.0.9600-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=dev)
-[![Travis build status](https://travis-ci.com/c1au6i0/covid19census_dev.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=dev)
-[![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=dev)
+[![](https://img.shields.io/badge/devel%20version-0.0.0.9600-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![Travis build status](https://travis-ci.com/c1au6i0/covid19census_dev.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=master)
+[![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 <!-- badges: end -->
@@
 
-*Note that the package has not been made by, and is not part of the covid19.cesus.gov website.*
+*Note that the package has not been made by and is not part of the covid19.census.gov website.*
 

### 2020-06-13 R/getus_functions.R
fixed bug segfaul

@@ -51,7 +51,7 @@ getus_covid_jhu <- function() {
 
     names(dat)[names(dat) == "value"] <- x
 
-    dat
+     dat
   })
 
   names(dat_l) <- metric_files
@@ -144,7 +144,7 @@ getus_covid <- function(repo = "jhu") {
     dat <- getus_covid_jhu()
   }
 
-  dat
+  as.data.frame(dat)
 }
 
 
@@ -260,7 +260,7 @@ getus_tests <- function() {
       onVentilatorCumulative = col_double(),
       recovered = col_double(),
       hash = col_character(),
-      dateChecked = col_datetime(format = ""),
+      dateChecked = col_skip(),
       death = col_double(),
       hospitalized = col_double(),
       total = col_double(),
@@ -291,7 +291,7 @@ getus_tests <- function() {
     "date", "state", "abbr", "positive", "negative", "pending", "hospitalized_currently",
     "hospitalized_cumulative", "in_icu_currently", "in_icu_cumulative",
     "on_ventilator_currently", "on_ventilator_cumulative", "recovered",
-    "hash", "date_checked", "death", "fips", "death_increase", "hospitalized_increase",
+    "hash", "death", "fips", "death_increase", "hospitalized_increase",
     "negative_increase", "positive_increase", "total_test_results_increase"
   )]
 
@@ -299,7 +299,7 @@ getus_tests <- function() {
     "date", "state", "abbr", "positive", "negative", "pending", "hospitalized_curr",
     "hospitalized_cumul", "icu_curr", "icu_cumul",
     "ventilator_curr", "ventilator_cumul", "recovered",
-    "hash", "date_checked", "death", "fips", "death_increase", "hospitalized_increase",
+    "hash", "death", "fips", "death_increase", "hospitalized_increase",
     "negative_increase", "positive_increase", "total_test_increase"
   )
   dat2
@@ -521,5 +521,5 @@ getus_all <- function(repo = "jhu") {
 
   names(dat2) <- stringr::str_replace(names(dat2), "tot_", "total_")
 
-  as.data.frame(dat2)
+  dat2
 }

### 2020-06-13 README.Rmd
readme

@@
 <!-- badges: start -->
-[![](https://img.shields.io/badge/devel%20version-0.0.0.9600-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.0.9700-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Travis build status](https://travis-ci.com/c1au6i0/covid19census_dev.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=master)

### 2020-06-13 README.Rmd
fixed travis badge

@@
 [![](https://img.shields.io/badge/devel%20version-0.0.0.9700-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
-[![Travis build status](https://travis-ci.com/c1au6i0/covid19census_dev.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=master)
+[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
 [![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+
 <!-- badges: end -->

### 2020-06-13 README.Rmd
readme

@@
 [![](https://img.shields.io/badge/devel%20version-0.0.0.9700-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
-[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
+[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=master)](https://travis-ci.com/c1au6i0/covid19census)
 [![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)

### 2020-06-15 R/getus_functions.R
hard fix url RCurl

@@ -240,9 +240,9 @@ getus_dex <- function() {
 getus_tests <- function() {
   url_data <- "https://covidtracking.com/api/v1/states/daily.csv"
 
-  if (RCurl::url.exists(url_data) == FALSE) {
-    stop("Something wrong with the repository or your internet connection!")
-  }
+  # if (RCurl::url.exists(url_data) == FALSE) {
+  #   stop("Something wrong with the repository or your internet connection!")
+  # }
 
 
   dat <- vroom::vroom(url_data,

### 2020-06-15 README.Rmd
readme

--- 2020-06-13 README.Rmd
@@
 <!-- badges: start -->
-[![](https://img.shields.io/badge/devel%20version-0.0.0.9700-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.0.9800-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2020-06-26 README.Rmd

@@
 <!-- badges: start -->
+
 [![](https://img.shields.io/badge/devel%20version-0.0.0.9800-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)

### 2020-06-26 README.Rmd
readme broken link

@@
 [![](https://img.shields.io/badge/devel%20version-0.0.0.9800-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
-[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
+[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=master)
 [![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
@@
 
-Raw data and the code used to import it can be found in the [data-raw](https://github.com/c1au6i0/convid19census/tree/master/data-raw) folder.
+Raw data and the code used to import it can be found in the [data-raw](https://github.com/c1au6i0/covid19census/tree/master/data-raw) folder.
 

### 2020-06-26 README.Rmd
badge fixed

@@
 [![](https://img.shields.io/badge/devel%20version-0.0.0.9800-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
-[![Travis build status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census.svg?token=PM6d4w2yt4HVWw7s2Jgy&branch=master)
+[![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
 [![Codecov test coverage](https://codecov.io/gh/c1au6i0/covid19census/branch/master/graph/badge.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)

### 2020-06-28 README.Rmd

@@
 ```{r intallation2, eval=FALSE, include=TRUE}
-# the repo at the moment is private so you need to be authorized 
-# don't share the token
 library(devtools)
 
-devtools::install_github("c1au6i0/covid19census_dev", auth_token = "8ca5cd2bb556a45f1ffeb2e9b44c90d4fbef6cf2", ref = "dev")
+devtools::install_github("c1au6i0/covid19census")
 ```

### 2020-07-01 README.Rmd
fixed fips numeric

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.0.9800-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.0.9900-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2020-07-01 notebooks/descriptive_tables_exploration.Rmd
fixed fips numeric

---
title: "Intersection COVID-19: Explore data"
date: "`r Sys.time()`"
categories:
  - covid-19
lastmod: "`r Sys.time()`"
keywords: []
description: ''
output:
  html_document:
    df_print: kable
    toc: true
    theme: journal
---

# Italy Descriptive Table

```{r library_data}
path_scripts <- ("scripts/")

source(paste0(path_scripts, "libraries_functions.R"))
```

## flu

```{r}

new_order_it <- c("n", "deaths", "case_ratio"," pop_km2", "lat", "perc_chronic_diab", "perc_chronic_bronc", "perc_chronic_hyp", "perc_chronic_heart" )

data_it <- getit_all()

today_it <- 
  data_it %>% 
    filter(date == max(date))

tab_it_flu <- 
  today_it %>% 
  filter(date == max(date)) %>% 
  select(region, deaths, cases, lat, perc_imm65, pop_km2, perc_chronic_diab,
         perc_chronic_bronc, perc_chronic_hyp, perc_chronic_heart, pop_tot) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  # 53 is median value
  mutate(perc_imm65= if_else(perc_imm65 > 53.00000, "high53", "low53")) %>% 
  mutate( deaths = deaths / pop_tot * 100000,
          case_ratio = cases / pop_tot * 100000
          ) %>% 
  group_by(perc_imm65) %>% 
  mutate(n = n()) %>% 
  summarise_if(is.numeric, funs(mean, sd)) %>% 
  pivot_longer(-perc_imm65) %>% 
  separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
  unite(perc_imm65, perc_imm65, metric) %>% 
  pivot_wider(names_from = perc_imm65, values_from = value) %>% 
  mutate_if(is.numeric, round,1)  %>% 
  mutate(
     "<53% (St.Dev.)"= paste(low53_mean,"(", low53_sd, ")"),
     ">53% (St.Dev.)"= paste(high53_mean,"(",high53_sd, ")")) %>% 
  select(var, "<53% (St.Dev.)", ">53% (St.Dev.)") %>% 
  filter(!var %in% c("pop_tot", "perc_imm65", "cases"))%>%
  arrange(match(var, !!new_order_it)) %>% 
  mutate(var = recode(var,
    "n" = "Number of counties",
    "deaths" = "COVID-19 death rate (per 100,000 people)",
    "case_ratio" = "COVID-19 confirmed-case rate (per 100,000 people)",
    "lat" = "Latitude",
    "chronic_heart" = "% chronic heart disease",
    "chronic_diab" = "% chronic diabetes",
    "chronic_bronc" =  "% chronic bronchitis",
    "chronic_hyp" = "% chronic hypertension",
    "pop_km2" = "Population density (people per squared km)"
    ))

```

## tot

```{r}
tab_it_tot <- 
  today_it %>% 
  select(region, deaths, cases, lat, perc_imm65, pop_km2, perc_chronic_diab,
         perc_chronic_bronc, perc_chronic_hyp, perc_chronic_heart, pop_tot) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  # 53 is median value
  mutate( deaths = deaths/pop_tot * 100000,
          case_ratio = cases / pop_tot * 100000,
          all_d = "all") %>% 
  group_by(all_d) %>% 
  mutate(n = n()) %>% 
  summarise_if(is.numeric, funs(mean, sd)) %>% 
  pivot_longer(-all_d) %>% 
  separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
  unite(all_d, all_d, metric ) %>% 
  pivot_wider(names_from = all_d, values_from = value) %>% 
  mutate_if(is.numeric, round, 1)  %>%

  mutate(
     "All (St.Dev.)" = paste(all_mean,"(", all_sd, ")")) %>% 
  select(var, "All (St.Dev.)") %>% 
  filter(!var %in% c("pop_tot", "imm65", "total_cases"))%>%
  arrange(match(var, !!new_order_it)) %>% 
  mutate(var = recode(var,
    "n" = "Number of counties",
    "deaths" = "COVID-19 death rate (per 100,000 people)",
     "case_ratio" = "COVID-19 confirmed-case rate (per 100,000 people)",
    "lat" = "Latitude",
    "chronic_heart" = "% chronic heart disease",
    "chronic_diab" = "% chronic diabetes",
    "chronic_bronc" =  "% chronic bronchitis",
    "chronic_hyp" = "% chronic hypertension",
    "pop_km2" = "Population density (people per squared km)"
    ))

tab_it_all <- inner_join(tab_it_tot, tab_it_flu, by = "var")

print(
  xtable(
    tab_it_all
  ),
  include.rownames=FALSE,
  comment = FALSE,
  align = rep("c", ncol(tab_it))
)
```

# US Descriptive Table

```{r us_scripts, warning=FALSE}
# path_scripts <- ("scripts/")
# us_scripts <- list(
#   "libraries_functions.R",
#   "us_preprocess.R"
#   # ,
#   # "us_analysis.R"
# )
# 
lapply(us_scripts, function(x) source(paste0(path_scripts, x)))

library(kableExtra)

```

```{r}
new_order <- c(
  #5: COVID-19
  "n",
  "death_rate",
  "case_ratio",
  "XX_days_f0",

  #5: Socioeconomic factors
  "XX_perc_families",
  "XX_perc_family_only_onep",
  "XX_perc_edu_bachelor",
  "XX_perc_withinternet",
  "XX_median_income",

  #19 Health factors
  "XX_ratio_beds",
  "XX_perc_alzheimer_dementia",
  "XX_perc_asthma",
  "XX_perc_atrial_fibrillation",
  "XX_perc_cancer_breast",
  "XX_perc_cancer_colorectal",
  "XX_perc_cancer_lung",
  "XX_perc_ch_obstructive_pulm",
  "XX_perc_chronic_kidney_disease",
  "XX_perc_depression",
  "XX_perc_diabetes",
  "XX_perc_heart_failure",
  "XX_perc_hypertension",
  "XX_perc_ischemic_heart_disease",
  "XX_perc_obesity",
  "XX_perc_osteoporosis",
  "XX_perc_rheumatoid_arthritis",
  "XX_perc_stroke",
  "XX_perc_tobacco_use",

  #5: Environmental factors
  "XX_pm2.5",
  "XX_summer_temp",
  "XX_summer_hum",
  "XX_winter_temp",
  "XX_winter_hum",

  #4 Population demographics
  "XX_median_age",
  "XX_perc_age65_over",
  "XX_sex_ratio",
  "XX_child_dependency",

  #7 Race
  "XX_perc_black",
  "XX_perc_lat",
  "XX_perc_white",
  "XX_perc_asian",
  "XX_perc_island",
  "XX_perc_other",
  "XX_perc_two_more_races"
)
```

## flu
```{r}
today_filt <-   today_us %>% 
  # THIS IMPORTANT
   filter(NC_cases >= 10)

median(today_filt$ZZ_perc_imm65)

tab_us_flu <-
  today_us %>% 
  # THIS IMPORTANT
   filter(NC_cases >= 10) %>% 
   mutate(median = median(ZZ_perc_imm65)) %>% 
   select(matches("ZZ|XX|YY|NP|NC")) %>% 
   select(
     -logitZZ_perc_imm65,
     -XX_total_tests,
     -XX_total_beds,

     ) %>% 
   mutate_at(vars(matches("perc")), ~ . * 100) %>%
   mutate(ZZ_perc_imm65 = if_else(ZZ_perc_imm65 > 45, "high45", "low45")) %>%

   mutate(
     death_rate = YY_deaths / NP_total_pop * 100000,
     case_ratio = NC_cases / NP_total_pop * 100000,
     XX_ratio_beds = XX_ratio_beds * 1000,
     XX_median_income	= XX_median_income / 1000
     ) %>%

   group_by(ZZ_perc_imm65) %>% 
   mutate(n = n()) %>% 
   summarize_if(is.numeric, funs(mean, sd)) %>%

   pivot_longer(-ZZ_perc_imm65) %>% 
   separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
   unite(imm65, ZZ_perc_imm65, metric ) %>%

   pivot_wider(names_from = imm65, values_from = value) %>% 
   mutate_if(is.numeric, round,1)  %>% 
   mutate(
     "\u226445% (St.Dev.)"= paste(low45_mean,"(", low45_sd, ")"),
     ">45% (St.Dev.)"= paste(high45_mean,"(", high45_sd, ")")) %>% 
  select(var, "\u226445% (St.Dev.)", ">45% (St.Dev.)") %>%
  filter(!var %in% c("NP_total_pop", "YY_deaths", "NC_cases")) %>% 
  arrange(match(var, !!new_order)) %>% 
  rename(Variable = var) %>% 
  mutate(Variable = recode(Variable,
    "n" = "Number of counties",

    "death_rate" = "COVID-19 death rate (per 100,000 people)",
    "case_ratio" = "COVID-19 confirmed-case rate (per 100,000 people)",
    "XX_days_f0" =  "Number of days since first case",

    "XX_perc_families" = "% families",
    "XX_perc_family_only_onep" = "% families with only one parent",
    "XX_perc_edu_bachelor" =   "% with bachelor",
    "XX_perc_withinternet" = "% with Internet",
    "XX_median_income" = "% median income ($1,000)",

    "XX_ratio_beds" = "Rate of hospital beds (per 1,000 people)",
    "XX_perc_alzheimer_dementia" = "% with Alzheimer's disease",
    "XX_perc_asthma" = "% with asthma",
    "XX_perc_atrial_fibrillation" = "% with atrial fibrillation",
    "XX_perc_cancer_breast" = "% with breast cancer",
    "XX_perc_cancer_colorectal" = "% with colorectal cancer",
    "XX_perc_cancer_lung" = "% with lung cancer",
    "XX_perc_ch_obstructive_pulm" = "% with obstructive pulmonary disease",
    "XX_perc_chronic_kidney_disease" = "% with chronic kidney disease",
    "XX_perc_depression" = "% with depression",
    "XX_perc_diabetes" = "% with diabetes",
    "XX_perc_heart_failure" = "% with heart failure",
    "XX_perc_hypertension" = "% with hypertension",
    "XX_perc_ischemic_heart_disease" = "% with ischemic heart disease",
    "XX_perc_obesity" = "% with obesity",
    "XX_perc_rheumatoid_arthritis" = "% with rheumatoid arthritis", 
    "XX_perc_stroke" = "% stroke transient ischemic attack",
    "XX_perc_tobacco_use" =  "% using tobacco",

    "XX_pm2.5" = "% average PM2.5(microg/m3)",
    "XX_summer_temp" = "Summer temperature (K)",
    "XX_summer_hum" =  "% summer humidity",
    "XX_winter_temp" = "Winter temperature (K)",
    "XX_winter_hum" = "% winter humidity",
    "XX_median_age" = "% median age",

    "XX_perc_age65_over" = "% 65 years or more of age",
    "XX_sex_ratio" = "Sex ratio",
    "XX_child_dependency" = "Child dependency ratio",

    "XX_perc_black" =   "% black",
    "XX_perc_lat" =  "% latino",
    "XX_perc_white" = "% white",
    "XX_perc_asian" = "% asian",
    "XX_perc_island" = "% island native",
    "XX_perc_other" =   "% other",
    "XX_perc_two_more_races" = "% two more races"

  )
  )

# total test need to be added after
```

## tot

```{r us_descriptive}

tab_us_tot <-
  today_us %>% 
  # THIS IMPORTANT
  filter(NC_cases >= 10) %>% 
   mutate(median = median(ZZ_perc_imm65)) %>% 
   select(matches("ZZ|XX|YY|NP|NC")) %>% 
   select(
     -logitZZ_perc_imm65,
     -XX_total_tests,
     -XX_total_tests,
     -XX_total_beds

     ) %>% 
   mutate_at(vars(matches("perc")), ~ . * 100) %>% 
   mutate(ZZ_perc_imm65 = if_else(ZZ_perc_imm65 > 45, "high45", "low45")) %>%

   mutate(
     death_rate = YY_deaths / NP_total_pop * 100000,
     case_ratio = NC_cases / NP_total_pop * 100000,
     # XX_urban = XX_urban * 100,
     XX_ratio_beds = XX_ratio_beds * 1000,
     XX_median_income	= XX_median_income / 1000,
     all_d = "all_d"
     ) %>%

   group_by(all_d) %>% 
   mutate(n = n()) %>% 
   summarize_if(is.numeric, funs(mean, sd)) %>%

   pivot_longer(-all_d) %>% 
   separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
   unite(all_d, all_d, metric ) %>% 
   pivot_wider(names_from = all_d, values_from = value) %>% 
    mutate_if(is.numeric, round, 1)  %>%
  mutate(
     "All (St.Dev.)" = paste(all_d_mean,"(", all_d_sd, ")")) %>% 
  select(var, "All (St.Dev.)") %>% 
  mutate_if(is.numeric, round, 1)  %>% 
  filter(!var %in% c("NP_total_pop", "YY_deaths", "NC_cases")) %>% 
  arrange(match(var, !!new_order)) %>% 
  rename(Variable = var) %>% 
  mutate(Variable = recode(Variable,
    "n" = "Number of counties",

    "death_rate" = "COVID-19 death rate (per 100,000 people)",
    "case_ratio" = "COVID-19 confirmed-case rate (per 100,000 people)",
    "XX_days_f0" =  "Number of days since first case",

    "XX_perc_families" = "% families",
    "XX_perc_family_only_onep" = "% families with only one parent",
    "XX_perc_edu_bachelor" =   "% with bachelor",
    "XX_perc_withinternet" = "% with Internet",
    "XX_median_income" = "% median income ($1,000)",

    "XX_ratio_beds" = "Rate of hospital beds (per 1,000 people)",
    "XX_perc_alzheimer_dementia" = "% with Alzheimer's disease",
    "XX_perc_asthma" = "% with asthma",
    "XX_perc_atrial_fibrillation" = "% with atrial fibrillation",
    "XX_perc_cancer_breast" = "% with breast cancer",
    "XX_perc_cancer_colorectal" = "% with colorectal cancer",
    "XX_perc_cancer_lung" = "% with lung cancer",
    "XX_perc_ch_obstructive_pulm" = "% with obstructive pulmonary disease",
    "XX_perc_chronic_kidney_disease" = "% with chronic kidney disease",
    "XX_perc_depression" = "% with depression",
    "XX_perc_diabetes" = "% with diabetes",
    "XX_perc_heart_failure" = "% with heart failure",
    "XX_perc_hypertension" = "% with hypertension",
    "XX_perc_ischemic_heart_disease" = "% with ischemic heart disease",
    "XX_perc_obesity" = "% with obesity",
    "XX_perc_rheumatoid_arthritis" = "% with rheumatoid arthritis", 
    "XX_perc_stroke" = "% stroke transient ischemic attack",
    "XX_perc_tobacco_use" =  "% using tobacco",

    "XX_pm2.5" = "% average PM2.5(microg/m3)",
    "XX_summer_temp" = "Summer temperature (K)",
    "XX_summer_hum" =  "% Summer humidity",
    "XX_winter_temp" = "Winter temperature (K)",
    "XX_winter_hum" = "% Winter humidity",
    "XX_median_age" = "% median age",

    "XX_perc_age65_over" = "% 65 years or more of age",
    "XX_sex_ratio" = "Sex ratio",
    "XX_child_dependency" = "Child dependency ratio",

    "XX_perc_black" =   "% black",
    "XX_perc_lat" =  "% latino",
    "XX_perc_white" = "% white",
    "XX_perc_asian" = "% asian",
    "XX_perc_island" = "% island native",
    "XX_perc_other" =   "% other",
    "XX_perc_two_more_races" = "% two more races"

  )
  )

tab_us_all <- inner_join(tab_us_tot, tab_us_flu, by = "Variable")

tab_us_clean <- tab_us_all %>% 
  mutate( Category = 
    c( "",
      rep("COVID-19", 3),
       rep("Socioeconomic factors", 5),
       rep("Health factors", 18),
       rep("Environmental factors", 5),
       rep("Population demographics", 4),
       rep("Race", 7)
    )
) %>% 
  tibble() %>% 
  relocate(Category, .before =  Variable) %>% 
  rename(All = "All (St.Dev.)", `≤Median` ="≤45% (St.Dev.)", `>Median` = ">45% (St.Dev.)"  )

first_row <- sub("\\s\\(\\s0\\s\\)","", tab_us_clean[1, ])
names(first_row) <- names(tab_us_clean)

tab_us_clean <- tab_us_clean %>% 
  filter(Variable != "Number of counties")

tab_us_clean <- bind_rows(first_row, tab_us_clean)

kable(tab_us_clean, "latex", longtable = T, booktabs = T) %>%  
  add_header_above(c(" ", " ", "Vaccination Coverage" = 3), bold = TRUE) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE) %>% 
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle") 
```

# #####################

# Some figures US

```{r us_ggcorr, fig.width=12, fig.height=12}
library(ggcorrplot)
M <-   
  today_us %>% 
  select_if(is.numeric) %>% 
  cor( use ="pairwise.complete.obs")
ggcorrplot(M,
           hc.order = TRUE
           # pch.cex = 30,
           # tl.cex = 30
)
```

# Rural -OUTDATED

```{r us_imm_state_rural, fig.width=4, fig.height=12, message=FALSE, warning=FALSE}
library(ggridges)

ordered_levels <-   
  today_us %>% 
  group_by(state) %>% 
  mutate(mean_z = mean(ZZ_perc_imm65)) %>%
  arrange(mean_z) %>% 
  ungroup() %>%
  select(state) %>% 
  unique() %>% 
  unlist()

today_us %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                         `1` = "urban"
                        ))
         ) %>%

  ggplot(aes(ZZ_perc_imm65, state, fill = urban)) +
      geom_density_ridges2(alpha = 0.5, scale = 1.1) +
      labs(x = "influenza vaccination in 65 and older (percent)",
           y = NULL
           ) +
  theme(legend.title = element_blank(),
        legend.position="top")
```

```{r us_imm_rural}
today_us %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                         `1` = "urban"
                        ))
         ) %>%

  ggplot(aes(ZZ_perc_imm65, fill = urban)) +
      geom_density(alpha = 0.5) +
      labs(x = "influenza vaccination in age 65 and older (percent)",
           y = NULL
           ) +
      scale_x_continuous(labels = scales::percent) +

  theme(legend.title = element_blank(),
        legend.position="top")
```

```{r us_cases_rural}
today_us %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                         `1` = "urban"
                        ))
         ) %>%

  ggplot(aes(log(NC_cases), fill = urban)) +
      geom_density(alpha = 0.5) +
      labs(x = "number of confirmed cases (log2)",
           y = NULL
           ) +
  theme(legend.title = element_blank(),
        legend.position="top")
```

```{r us_deaths, fig.width=4, fig.height=12, message=FALSE, warning=FALSE}
ordered_levels <-   
  today_us %>% 
  group_by(state) %>% 
  mutate(mean_z = mean(YY_deaths)) %>%
  arrange(mean_z) %>% 
  ungroup() %>%
  select(state) %>% 
  unique() %>% 
  unlist()

today_us %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                        `1` = "urban"
                        ))) %>%

  ggplot(aes(log10(YY_deaths), state, col = urban)) +
      geom_jitter(alpha = 0.5) +
      labs(x = "number of deaths (log2)") +
     theme(legend.title = element_blank(),
           legend.position="top")

# today_us %>% 
#   filter(YY_deaths > 15000)
```

# Deaths

```{r us_deaths2}
today_us %>% 
   ungroup() %>% 
   group_by(state) %>% 
   summarize(deaths  = sum(YY_deaths)) %>%

   ggplot(aes(reorder(state, deaths), deaths)) +
     geom_col() +
     scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
     labs(y = NULL)

```

```{r us_perc_state}
today_us %>% 
   group_by(date) %>% 
   mutate(tot = sum(YY_deaths)) %>% 
   group_by(state) %>%
   summarize(perc_death = round((sum(YY_deaths)/first(tot) * 100), 2)) %>% 
   arrange(desc(perc_death))
```

```{r us_perc_case_state}
today_us %>% 
   group_by(date) %>% 
   mutate(tot = sum(NC_cases)) %>% 
   group_by(state) %>%
   summarize(perc_cases = round((sum(NC_cases)/first(tot) * 100),2)) %>% 
   arrange(desc(perc_cases))
```

```{r us_cases_state}
options(scipen = 999)
today_us %>% 
   ungroup() %>% 
   group_by(state) %>% 
   summarize(cases  = sum(NC_cases)) %>%

   ggplot(aes(reorder(state, cases), cases)) +
     geom_col() +
     scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
    labs(y = NULL)

```

```{r u_age_by_state}
today_us %>% 
  filter(date == max(date)) %>% 
  ggplot(aes(state, XX_perc_age65_over)) +
      geom_boxplot() +
      scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45)) 
```

```{r us_urban_pop}
# Is urban defined by pop size?
today_us %>% 
  mutate(XX_urban = factor(recode(XX_urban,
                        `0` = "rural",
                        `1` = "urban"
                        ))) %>% 
  ggplot(aes(XX_urban, NP_total_pop/1000, col = XX_urban)) +
    geom_jitter() +

   labs(y = " Population in thousands",
        x = NULL) +
  theme(legend.title = element_blank())

```

# Lord of the Rings

```{r}

path_f <- "/Users/heverz/Dropbox (MechPred)/COVID19/Interference/figs/IT/no_log"

to_plot <- today_it_all %>% 
  select(where(is.numeric)) %>% 
  names()
# 
lord_rings(dat = dat_it_cl, "YY_deaths", model_dat = TRUE, nation = "it")

lapply(to_plot, lord_rings,
       today_it_all,
       model_dat = FALSE,
       log_v = FALSE,
       nation = "it",
       path_f

       )

today_it_all %>% 
  ggplot(aes(deaths, region)) +
    geom_col()

```

```{r}
path_f = "/Users/heverz/Dropbox (MechPred)/COVID19/Interference/figs/US/model/nolog"

names(today_us) <- stringr::str_replace(names(today_us), "-", "_")

numeric_vars <- today_us %>% 
  select(where(is.numeric)) %>% 
  names()

lapply(numeric_vars, lord_rings,
       today_us,
       TRUE,
       FALSE, #log
       path_f 
)

dput(names(today_us))

```

# NYT vs JHU

```{r}
jhu <- getus_all(repo = "jhu") %>% 
    filter(date == max(date))

nyt <- getus_all(repo = "nyt") %>% 
   filter(date == max(date))

sum(unique(nyt$fips)) - sum(unique(jhu$fips))

j_data <- inner_join(jhu, nyt, by = "fips", suffix = c("_jhu", "_nyt") )

j_data_s <- 
  j_data %>% 
  select( matches("date|county|state|fips|cases|deaths")) %>% 
  mutate(case_diff = cases_jhu - cases_nyt, deaths_diff = deaths_jhu - deaths_nyt)

j_data_s %>% 
  ggplot(aes(fips, deaths_diff)) +
    geom_line()
```

```{r}
j_data_s %>% 
  ggplot(aes(fips, case_diff)) +
    geom_line()
```

```{r fig.height=10, fig.width=5}

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

library(plotly)
x <- j_data_s %>% 
  filter(deaths_diff != 0) %>% 
  ggplot(aes( case_diff, state_jhu, col = state_jhu, fips = fips), county_jhu = county_jhu, county_nyt = county_nyt) +
    geom_jitter() +
    theme(legend.position = "none") +
  labs(x = "JHU deaths - NYT deaths",
       y = NULL)

p <- ggplotly(x, width = 500, height = 800, tooltip = c("deaths_diff",
                                                   "fips",
                                                   "county_jhu",
                                                   "state_jhu"

                                                   )) %>% 
  config(p = .)

p
htmlwidgets::saveWidget(p, "diff_deaths.html")
```

# Unassigned for JHU

```{r fig.width=5, fig.height=6}
data_freeze == "2020-06-10"

dat_all_jhu <- getus_all()

dat_all_jhu %>% 
  filter(date == !!data_freeze) %>% 
  mutate(county_un = if_else(county == "Unassigned", "unassigned", "assigned")) %>% 
  group_by(state, county_un) %>% 
  summarize(n = n(), deaths = sum(deaths, na.rm = TRUE), .groups = "drop") %>% 
  pivot_wider(id_cols = "state", names_from = "county_un", values_from = "deaths") %>% 
  mutate(perc_unas = unassigned/sum(assigned,unassigned, na.rm = TRUE) * 100) %>%

  ggplot(aes(unassigned, state)) +
    geom_col() +
    labs(y = NULL,
             x = "Deaths not assigned to a County (JHU repository)")

```

# Ny State Counties

```{r fig.height=6, fig.width=4}
today_us_filt %>% 
  filter(state == "New York") %>%

  ggplot(aes(YY_deaths, reorder(county, YY_deaths))) +
    geom_col()

```

# Explore quintiles

```{r}
# again we select the names that start we XX
XX <- grep("XX", names(all_dat), value = TRUE)
# we don't include total beds because we have the bed X population already in
XX <- XX[!XX %in% c("XX_total_beds")]

# formula for propensity score
form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

PropScores_dat <- lm(form_ps, data = today_us_filt)

# dat quint.
dat_quint_split <- 
  today_us_filt %>% 
  mutate(PP = as.factor(ntile(fitted.values(PropScores_dat), 5)),
         PP_cont = fitted.values(PropScores_dat)
         )

print(dat_quint_split , width = 100)

```

```{r}

path_f <- "/Users/heverz/Dropbox (MechPred)/COVID19/Interference/figs/US_quint/log"

to_plot <- dat_quint_split %>% 
  select(where(is.numeric)) %>% 
  names()

lapply(
      to_plot,
      fifth_element,
      dat = dat_quint_split,
      log_v = TRUE,
      path_f
       )
```

```{r}
dat_quint_split %>% 
  filter(state == "New York") %>% 
  select(PP) %>% 
  summary()
```

### 2020-07-01 notebooks/old_notebooks/report_italy_wp.Rmd
fixed fips numeric

--- 2020-04-21 data-raw/it/how_get_data.Rmd
@@
 ---
+
+# Objective
+
+Investigate the relation (if any) between influenza vaccination coverage and measures related to COVID-19 epidemic in Italy.
+Demigrafical and health related data of each italian region are also taken into account.
 
@@
 dem_it_p <- vroom(
-  "data/italy/pop_it.csv",
+  "/Users/heverz/Documents/R_projects/covid19_interference/data/italy/pop_it.csv",
   col_types = cols(
@@
 
-chronic_it_p <- vroom("data/italy/chronic_conditions.csv",
+chronic_it_p <- vroom("/Users/heverz/Documents/R_projects/covid19_interference/data/italy/chronic_conditions.csv",
   col_types = cols(
@@
 ```{r smoking_it, warning=FALSE, message=FALSE}
-smoking_it <- vroom("data/italy/smoking.csv",
+smoking_it <- vroom("/Users/heverz/Documents/R_projects/covid19_interference/data/italy/smoking.csv",
   col_types = cols(
@@
 ```{r obesity_it, warning=FALSE, message=FALSE}
-bweight_it <- vroom("data/italy/bweight_it.csv", col_types = cols(
+bweight_it <- vroom("/Users/heverz/Documents/R_projects/covid19_interference/data/italy/bweight_it.csv", col_types = cols(
   ITTER107 = col_character(),
@@
 
+With this we check that the regions are the same across datasets. They are not always very consistent with names.
+```{r check_join_all}
+# check that the names are right otherwise when we join we loose data
+# lapply(list(
+#   cmr_it,
+#   dem_it_wider,
+#   regions_area,
+#   chronic_it,
+#   cancer_it,
+#   smoking_it,
+#   bweight_it
+# ), function(x) {
+#   unique(x$region) %in% unique(dem_it_wider$region)
+# })
+```
+
 ```{r join_all}
@@
 Now we have our datasets ready for being analyzed...what we need to do is just filter for the `COVID19_var` and date of interest and then run the stats.
+
+## Now We Rock`n`roll baby
+
+Let' start looking at COVID-19 measures.
+
+```{r}
+unique(all_dat$COVID19_var)
+```
+
+### total_cases
+
+```{r total_cases, fig.height= 6, fig.width=6}
+all_dat %>% 
+    filter(COVID19_var == "total_cases") %>% 
+    ggplot(aes(date, value, group = region, color = region)) + 
+      geom_line() +
+      # geom_smooth(method = "lm", formula = y~x, se = FALSE) +
+      scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
+      theme_bw() +
+      theme(legend.position = "none") +
+      facet_wrap(vars(region)) +
+        labs(
+        x = "date",
+        y = "confirmed cases") 
+```
+```{r total_cases_log, fig.height= 6, fig.width=6}
+all_dat %>% 
+    filter(COVID19_var == "total_cases") %>% 
+    ggplot(aes(date, log2(value), group = region, color = region)) + 
+      geom_line() +
+      # geom_smooth(method = "lm", formula = y~x, se = FALSE) +
+      scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
+      theme_bw() +
+      theme(legend.position = "none") +
+      facet_wrap(vars(region)) +
+        labs(
+        x = "date",
+        y = "confirmed cases (log2)") 
+```
+Clearly at the extreme of date we lose linearity. But that might make sense
+
+Let's do some smooth analysis.
+
+```{r total_cases_smooth, fig.height= 6, fig.width=6, warning=FALSE}
+p <- ggplotly(all_dat %>% 
+    filter(COVID19_var == "total_cases") %>% 
+    ggplot(aes(date, log2(value), group = region, color = region)) + 
+      # geom_line() +
+      geom_smooth(method = "lm", formula = y~x, se = FALSE, lwd = 0.5) +
+      scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
+      theme_bw() +
+      theme(legend.position = "none") +
+      # facet_wrap(vars(region)) +
+        labs(
+        x = "date",
+        y = "confirmed cases"),
+    tooltip = c("date", "log2(value)", "group")
+
+    ) %>% 
+   config(p = ., displayModeBar = FALSE)
+
+p
+```
+
+Well beside `Piemonte` and `Molise`, rate is very similar.
+
+othe vars:
+
+cmr: altready linear. saturation hospitals
+chr: funny at the beginning they were detecting only peple that were hospitalizing
+cir:
+```{r mortality_log, fig.height= 6, fig.width=6, message=FALSE, warning=FALSE}
+all_dat %>% 
+    filter(COVID19_var == "chr") %>% 
+    ggplot(aes(date, value, group = region, color = region)) + 
+      geom_line() +
+      # geom_smooth(method = "lm", formula = y~x, se = FALSE) +
+      scale_x_date(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
+      theme_bw() +
+      theme(legend.position = "none") +
+      facet_wrap(vars(region)) +
+        labs(
+        x = "date")
+```
+
+<!-- This chunk below save datasets and data in my folder that then sync with dropbox. -->
+```{r savedata, include=FALSE, eval=FALSE}
+# files   ############
+# report_workingprogress: notebook with all the progresses
+# renv.lock: file to restore libraries
+saveRDS(all_dat, "data/italy/all_dat.RDS")
+
+drop_upload("data/italy/all_dat.RDS", path = "covid19/interference/data/")
+drop_upload("tidy_data_script.R", path = "covid19/interference/code/")
+drop_upload("data/italy/all_dat.RDS", path = "covid19/interference/data/italy/")
+drop_upload("report_italy_wp.html", path = "covid19/interference/text/")
+drop_upload("renv.lock", path = "covid19/interference/renv/")
+drop_upload("report_italy_wp.Rmd", path = "covid19/interference/code/")
+drop_upload("data/italy/NPI_it.xlsx", path = "covid19/interference/data/italy/raw/")
+```

### 2020-07-01 notebooks/old_notebooks/us_explore2.Rmd
fixed fips numeric

---
title: "Intersection COVID-19: Explore data"
date: "`r Sys.time()`"
categories:
  - covid-19
tags: 
  - covid-19
  - Italy
lastmod: "`r Sys.time()`"all
keywords: []
description: ''
output:
  html_document:
    df_print: kable
    toc: true
    theme: journal
---

# Objective

```{r}
library(facetscales)
library(covid19census)
library(tidyverse)
library(ggcorrplot)
library(GGally)
library(gganimate)
library(ggcorrplot)
library(broom)
library(purrr)
library(knitr)
library(kableExtra)
library(htmlwidgets)
library(plotly)
library(stringr)
library(ggridges)
library(xtable)
theme_set(theme_bw())
```

# Italy Table

## flu block
```{r}
data_it <- getit_all()

today_it <- 
  data_it %>% 
    filter(date == max(date))

new_order_it <- c("n", "deaths", "pop_km2", "lat", "chronic_diab", "chronic_bronc", "chronic_hyp")

tab_it_flu <- 
  today_it %>% 
  select(region, deaths, lat, imm65, pop_km2, chronic_diab, chronic_bronc, chronic_hyp, pop_tot) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  # 53 is median value
  mutate(imm65= if_else(imm65 > 50, "high53", "low53")) %>% 
  mutate( deaths = deaths/pop_tot * 100000) %>% 
  group_by(imm65) %>% 
  mutate(n = n()) %>% 
  summarise_if(is.numeric, funs(mean, sd)) %>% 
  pivot_longer(-imm65) %>% 
  separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
  unite(imm65, imm65, metric ) %>% 
  pivot_wider(names_from = imm65, values_from = value) %>% 
  mutate_if(is.numeric, round,1)  %>% 
  mutate(
     "<53% (St.Dev.)"= paste(low53_mean,"(", low53_sd, ")"),
     ">53% (St.Dev.)"= paste(high53_mean,"(",high53_sd, ")")) %>% 
  select(var, "<53% (St.Dev.)", ">53% (St.Dev.)") %>% 
  filter(var != "pop_tot") %>%
  arrange(match(var, !!new_order_it)) %>% 
  mutate(var = recode(var,
    "n" = "Number of counties",
    "deaths" = "COVID-19 death rate (per 100,000 people)",
    "lat" = "Latitude",
    "chronic_diab" = "% chronic diabetes",
    "chronic_bronc" =  "% chronic bronchitis",
    "chronic_hyp" = "% chronic hypertension",
    "pop_km2" = "Population density (people per squared km)"
    ))

```

### tot

```{r}
new_order_it <- c("n", "deaths", "pop_km2", "lat", "chronic_diab", "chronic_bronc", "chronic_hyp")

tab_it_tot <- 
  today_it %>% 
  select(region, deaths, lat, imm65, pop_km2, chronic_diab, chronic_bronc, chronic_hyp, pop_tot) %>% 
  mutate_if(is.numeric, round, 1) %>% 
  # 53 is median value
  mutate( deaths = deaths/pop_tot * 100000,
          all_d = "all") %>% 
  group_by(all_d) %>% 
  mutate(n = n()) %>% 
  summarise_if(is.numeric, funs(mean, sd)) %>% 
  pivot_longer(-all_d) %>% 
  separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
  unite(all_d, all_d, metric ) %>% 
  pivot_wider(names_from = all_d, values_from = value) %>% 
  mutate_if(is.numeric, round, 1)  %>%

  mutate(
     "All (St.Dev.)" = paste(all_mean,"(", all_sd, ")")) %>% 
  select(var, "All (St.Dev.)") %>% 
  filter(!var %in% c("pop_tot", "imm65"))%>%
  arrange(match(var, !!new_order_it)) %>% 
  mutate(var = recode(var,
    "n" = "Number of counties",
    "deaths" = "COVID-19 death rate (per 100,000 people)",
    "lat" = "Latitude",
    "chronic_diab" = "% chronic diabetes",
    "chronic_bronc" =  "% chronic bronchitis",
    "chronic_hyp" = "% chronic hypertension",
    "pop_km2" = "Population density (people per squared km)"
    ))

tab_it_all <- inner_join(tab_it_tot, tab_it_flu, by = "var")

print(
  xtable(
    tab_it_all
  ),
  include.rownames=FALSE,
  comment = FALSE,
  align = rep("c", ncol(tab_it))
)

```

# US Table

```{r warning=FALSE}

logit = function(p){ log(p) - log(1-p) }

# Thi are the variable that are initially selected from the dataframe
# var new to select -------
to_select<- c("date", "county", "state", "fips", "cases", "deaths", "perc_families", 
"perc_family_only_onep", "perc_edu_bachelor", 
"perc_withinternet", "total_pop", "perc_imm65", "total_imm65", 
"total_beds", "ratio_beds", "urban",  
"perc_alzheimer_dementia", "perc_asthma", "perc_atrial_fibrillation", 
"perc_cancer_breast", "perc_cancer_colorectal", "perc_cancer_lung", 
"perc_ch_obstructive_pulm", "perc_chronic_kidney_disease", "perc_depression", 
"perc_diabetes", "perc_heart_failure", "perc_hypertension", "perc_ischemic_heart_disease", "perc_obesity", "perc_osteoporosis", "perc_rheumatoid_arthritis", 
 "perc_stroke", "perc_tobacco_use", 
"median_income", "pm2.5", "summer_temp", "summer_hum", "winter_temp", 
"winter_hum", "age65_over", "median_age", "sex_ratio", "child_dependency", 
"perc_black", "perc_lat", "perc_white", "perc_asian", "perc_island", 
"perc_other", "perc_two_more_races", "total_tests", "days_f0"
)

# df1   -----------------------------
df1 <- getus_all()

# df2 ------------------------------
df2 <-
  df1 %>%
  # calculate age65_over and perc_black, white and latinos, total tests
  mutate(age65_over = rowSums(dplyr::select(., matches("perc_[6][5-9]|[7|8][0-9]")))) %>%
  mutate(urban = if_else(urban == "Urban", 1, 0)) %>%
  mutate(perc_black = total_black / total_pop * 100) %>%
  mutate(perc_white = total_white / total_pop * 100) %>%
  mutate(perc_lat = total_latino / total_pop * 100) %>%
  mutate(total_tests = positive + negative) %>%

  # calculate day since first case
  mutate(f_date = case_when(cases >= 1 ~ date)) %>%
  group_by(fips) %>%
  mutate(f_date = min(f_date, na.rm = TRUE), days_f0 = as.numeric(date - f_date)) %>%
  ungroup() %>%
  mutate(days_f0 = if_else(is.finite(days_f0), days_f0, NA_real_)) %>%

  # newly calculated-----------------------------------  
  # total beds in ratio
  mutate(ratio_beds = total_beds/total_pop) %>% 
  # asian are here
  mutate(perc_asian = total_asian / total_pop * 100) %>%
  mutate(perc_island = total_pacific_islander / total_pop * 100) %>%
  mutate(perc_native = total_native / total_pop * 100) %>%
  mutate(perc_other = total_other_race / total_pop * 100) %>%
  mutate(perc_two_more_races = total_two_more_races / total_pop * 100) %>%

  # family with one parent together
  mutate(perc_family_only_onep = perc_families_only_female + perc_families_only_male) %>% 
  mutate(total_imm65 = age65_over/100 *perc_imm65) %>%

  # select variables ----------
  dplyr::select(!!to_select) %>%

  # perc divided by 100
  mutate_at(vars(starts_with("perc")), function(x) x / 100) %>%
  mutate(age65_over = age65_over / 100) %>%

  # we remove NAs
  na.omit() %>%

  # at least 1 death
  filter(deaths >= 1)

# all_dat -------------
all_dat <- df2 %>%
  # make variables numeric
  mutate_at(vars(
    -date, -county, -state, -fips), 
     as.numeric
  ) %>%
  # rename
  # add XX_ to all the variables excepts those
  rename_at(
    vars(
      -date, -county, -state, -fips, -cases, -deaths,
      -perc_imm65, -cases, -total_pop, -total_imm65
    ),
    ~ paste0("XX_", .)
  ) %>%
  rename_at(vars(perc_imm65, total_imm65), ~ paste0("ZZ_", .)) %>%
  rename_at(vars(deaths), ~ paste0("YY_", .)) %>%
  rename_at(vars(cases), ~ paste0("NC_", .)) %>%
  rename_at(vars(total_pop), ~ paste0("NP_", .)) %>%
  # calculate logitZZ
  mutate(logitZZ_perc_imm65 = logit(ZZ_perc_imm65)) 
  # mutate(logitZZ_total_imm65 = log(ZZ_total_imm65) - log(1 - ZZ_total_imm65))

# today_dat ------------------------------
today_dat <- all_dat %>%
  filter(date == max(date))
```

```{r}
new_order <- c(
  "n",
  "death_rate",
  "XX_perc_families",
  "XX_perc_family_only_onep",
  "XX_perc_edu_bachelor",
  "XX_perc_withinternet",
  "XX_ratio_beds",
  "XX_urban",
  "XX_perc_alzheimer_dementia",
  "XX_perc_asthma",
  "XX_perc_atrial_fibrillation",
  "XX_perc_cancer_breast",
  "XX_perc_cancer_colorectal",
  "XX_perc_cancer_lung",
  "XX_perc_ch_obstructive_pulm",
  "XX_perc_chronic_kidney_disease",
  "XX_perc_depression",
  "XX_perc_diabetes",
  "XX_perc_heart_failure",
  "XX_perc_hypertension",
  "XX_perc_ischemic_heart_disease",
  "XX_perc_obesity",
  "XX_perc_osteoporosis",
  "XX_perc_rheumatoid_arthritis",
  "XX_perc_stroke",
  "XX_perc_tobacco_use",
  "XX_median_income",
  "XX_pm2.5",
  "XX_summer_temp",
  "XX_summer_hum",
  "XX_winter_temp",
  "XX_winter_hum",
  "XX_median_age",
  "XX_age65_over",
  "XX_sex_ratio",
  "XX_child_dependency",
  "XX_perc_black",
  "XX_perc_lat",
  "XX_perc_white",
  "XX_perc_asian",
  "XX_perc_island",
  "XX_perc_other",
  "XX_perc_two_more_races",
  "XX_days_f0"
)
```

## flu blocks

```{r}

# RATIO BEDS X THOUSANDS
pm2.5_name <- expression("% average PM2.5(" ~ mu ~ g / m^3 ~ ")")

tab_us_flu <- today_dat %>% 
   mutate(median = median(ZZ_perc_imm65)) %>% 
   select(matches("ZZ|XX|YY|NP")) %>% 
   select(
     -ZZ_total_imm65,
     -logitZZ_perc_imm65,
     -XX_total_tests,
     -XX_total_tests,
     -XX_total_beds

     ) %>% 
   mutate_at(vars(matches("perc")), ~ . * 100) %>% 
   mutate(ZZ_perc_imm65 = if_else(ZZ_perc_imm65 > 46, "high46", "low46")) %>%

   mutate(
     death_rate = YY_deaths / NP_total_pop * 100000,
     XX_urban = XX_urban * 100,
     XX_ratio_beds = XX_ratio_beds * 1000,
     XX_median_income	= XX_median_income / 1000,
     XX_sex_ratio = XX_sex_ratio * 100,
     XX_child_dependency = XX_sex_ratio * 100,
     ) %>%

   group_by(ZZ_perc_imm65) %>% 
   mutate(n = n()) %>% 
   summarize_if(is.numeric, funs(mean, sd)) %>%

   pivot_longer(-ZZ_perc_imm65) %>% 
   separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
   unite(imm65, ZZ_perc_imm65, metric ) %>%

   pivot_wider(names_from = imm65, values_from = value) %>% 
   mutate_if(is.numeric, round,1)  %>% 
   mutate(
     "<46% (St.Dev.)"= paste(low46_mean,"(", low46_sd, ")"),
     ">46% (St.Dev.)"= paste(high46_mean,"(",high46_sd, ")")) %>% 
  select(var, "<46% (St.Dev.)", ">46% (St.Dev.)") %>%
  filter(!var %in% c("NP_total_pop", "YY_deaths")) %>% 
  arrange(match(var, !!new_order)) %>% 
  mutate(var = recode(var,
    "n" = "Number of counties",
    "death_rate" = "COVID-19 death rate (per 100,000 people)",
    "XX_perc_families" = "% families",
    "XX_perc_family_only_onep" = "% families with only one parent",
    "XX_perc_edu_bachelor" =   "% with bachelor",
    "XX_perc_withinternet" = "% with Internet",
    "XX_ratio_beds" = "Rate of hospital beds (per 100,000 people)",
    "XX_urban" = "% urban counties",
    "XX_perc_alzheimer_dementia" = "% with Alzheimer's disease",
    "XX_perc_asthma" = "% with asthma",
    "XX_perc_atrial_fibrillation" = "% with atrial fibrillation",
    "XX_perc_cancer_breast" = "% with breast cancer",
    "XX_perc_cancer_colorectal" = "% with colorectal cancer",
    "XX_perc_cancer_lung" = "% with lung cancer",
    "XX_perc_ch_obstructive_pulm" = "% with obstructive pulmonary disease",
    "XX_perc_chronic_kidney_disease" = "% with chronic kidney disease",
    "XX_perc_depression" = "% with depression",
    "XX_perc_diabetes" = "% with diabetes",
    "XX_perc_heart_failure" = "% with heart failure",
    "XX_perc_hypertension" = "% with hypertension",
    "XX_perc_ischemic_heart_disease" = "% with ischemic heart disease",
    "XX_perc_obesity" = "% with obesity",
    "XX_perc_osteoporosis" = "% with osteoporosis",
    "XX_perc_rheumatoid_arthritis" = "% with rheumatoid arthritis", 
    "XX_perc_stroke" = "% stroke transient ischemic attack",
    "XX_perc_tobacco_use" =  "% using tobacco",
    "XX_median_income" = "% median income ($1,000)",
    "XX_pm2.5" = "% average PM2.5(microg/m3)",
    "XX_summer_temp" = "Summer temperature (K)",
    "XX_summer_hum" =  "Summer humidity (mixing ratio)", 
    "XX_winter_temp" = "Winter temperature (K)",
    "XX_winter_hum" = "Winter humidity (mixing ratio)",
    "XX_median_age" = "% median age",
    "XX_age65_over" = "Number 65 years of age",
    "XX_sex_ratio" = "Sex ratio",
    "XX_child_dependency" = "Child dependency ratio",
    "XX_perc_black" =   "% black",
    "XX_perc_lat" =  "% latino",
    "XX_perc_white" = "% white",
    "XX_perc_asian" = "% asian",
    "XX_perc_island" = "% island native",
    "XX_perc_other" =   "% other",
    "XX_perc_two_more_races" = "% two more races",
    "XX_days_f0" =  "Number of days since first case"

  )
  )

# total test need to be added after
```

## tot

```{r US_ALL}
tab_us_tot <-
  today_dat %>% 
   mutate(median = median(ZZ_perc_imm65)) %>% 
   select(matches("ZZ|XX|YY|NP")) %>% 
   select(
     -ZZ_total_imm65,
     -logitZZ_perc_imm65,
     -XX_total_tests,
     -XX_total_tests,
     -XX_total_beds

     ) %>% 
   mutate_at(vars(matches("perc")), ~ . * 100) %>% 
   mutate(ZZ_perc_imm65 = if_else(ZZ_perc_imm65 > 46, "high46", "low46")) %>%

   mutate(
     death_rate = YY_deaths / NP_total_pop * 100000,
     XX_urban = XX_urban * 100,
     XX_ratio_beds = XX_ratio_beds * 1000,
     XX_median_income	= XX_median_income / 1000,
     XX_sex_ratio = XX_sex_ratio * 100,
     XX_child_dependency = XX_sex_ratio * 100,
     ) %>%

   group_by(all_d) %>% 
   mutate(n = n()) %>% 
   summarize_if(is.numeric, funs(mean, sd)) %>%

   pivot_longer(-all_d) %>% 
   separate(name, c("var", "metric"), "_(?=(mean)|(sd))") %>% 
   unite(all_d, all_d, metric ) %>% 
   pivot_wider(names_from = all_d, values_from = value) %>% 
    mutate_if(is.numeric, round, 1)  %>%
  mutate(
     "All (St.Dev.)" = paste(all_d_mean,"(", all_d_sd, ")")) %>% 
  select(var, "All (St.Dev.)") %>% 
  mutate_if(is.numeric, round, 1)  %>% 
  filter(!var %in% c("NP_total_pop", "YY_deaths")) %>% 
  arrange(match(var, !!new_order)) %>% 
 mutate(var = recode(var,
    "n" = "Number of counties",
    "death_rate" = "COVID-19 death rate (per 100,000 people)",
    "XX_perc_families" = "% families",
    "XX_perc_family_only_onep" = "% families with only one parent",
    "XX_perc_edu_bachelor" =   "% with bachelor",
    "XX_perc_withinternet" = "% with Internet",
    "XX_ratio_beds" = "Rate of hospital beds (per 100,000 people)",
    "XX_urban" = "% urban counties",
    "XX_perc_alzheimer_dementia" = "% with Alzheimer's disease",
    "XX_perc_asthma" = "% with asthma",
    "XX_perc_atrial_fibrillation" = "% with atrial fibrillation",
    "XX_perc_cancer_breast" = "% with breast cancer",
    "XX_perc_cancer_colorectal" = "% with colorectal cancer",
    "XX_perc_cancer_lung" = "% with lung cancer",
    "XX_perc_ch_obstructive_pulm" = "% with obstructive pulmonary disease",
    "XX_perc_chronic_kidney_disease" = "% with chronic kidney disease",
    "XX_perc_depression" = "% with depression",
    "XX_perc_diabetes" = "% with diabetes",
    "XX_perc_heart_failure" = "% with heart failure",
    "XX_perc_hypertension" = "% with hypertension",
    "XX_perc_ischemic_heart_disease" = "% with ischemic heart disease",
    "XX_perc_obesity" = "% with obesity",
    "XX_perc_osteoporosis" = "% with osteoporosis",
    "XX_perc_rheumatoid_arthritis" = "% with rheumatoid arthritis", 
    "XX_perc_stroke" = "% stroke transient ischemic attack",
    "XX_perc_tobacco_use" =  "% using tobacco",
    "XX_median_income" = "% median income ($1,000)",
    "XX_pm2.5" = "% average PM2.5(microg/m3)",
    "XX_summer_temp" = "Summer temperature (K)",
    "XX_summer_hum" =  "Summer humidity (mixing ratio)", 
    "XX_winter_temp" = "Winter temperature (K)",
    "XX_winter_hum" = "Winter humidity (mixing ratio)",
    "XX_median_age" = "% median age",
    "XX_age65_over" = "Number 65 years of age",
    "XX_sex_ratio" = "Sex ratio",
    "XX_child_dependency" = "Child dependency ratio",
    "XX_perc_black" =   "% black",
    "XX_perc_lat" =  "% latino",
    "XX_perc_white" = "% white",
    "XX_perc_asian" = "% asian",
    "XX_perc_island" = "% island native",
    "XX_perc_other" =   "% other",
    "XX_perc_two_more_races" = "% two more races",
    "XX_days_f0" =  "Number of days since first case"
  )
  )

tab_us_all <- inner_join(tab_us_tot, tab_us_flu, by = "var")

print(
  xtable(
    tab_us_all
  ),
  include.rownames=FALSE,
  comment = FALSE,
  align = rep("c", ncol(tab_it))
)

```

# Some figures US

```{r fig.width=12, fig.height=12}
library(ggcorrplot)
M <-   
  today_dat %>% 
  select_if(is.numeric) %>% 
  cor( use ="pairwise.complete.obs")
ggcorrplot(M,
           hc.order = TRUE
           # pch.cex = 30,
           # tl.cex = 30
)
```

```{r fig.width=4, fig.height=12, message=FALSE, warning=FALSE}
ordered_levels <-   
  today_dat %>% 
  group_by(state) %>% 
  mutate(mean_z = mean(ZZ_perc_imm65)) %>%
  arrange(mean_z) %>% 
  ungroup() %>%
  select(state) %>% 
  unique() %>% 
  unlist()

today_dat %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                         `1` = "urban"
                        ))
         ) %>%

  ggplot(aes(ZZ_perc_imm65, state, fill = urban)) +
      geom_density_ridges2(alpha = 0.5, scale = 1.1) +
      labs(x = "influenza vaccination in 65 and older (percent)",
           y = NULL
           ) +
  theme(legend.title = element_blank(),
        legend.position="top")
```
```{r}
today_dat %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                         `1` = "urban"
                        ))
         ) %>%

  ggplot(aes(ZZ_perc_imm65, fill = urban)) +
      geom_density(alpha = 0.5) +
      labs(x = "influenza vaccination in 65 and older (percent)",
           y = NULL
           ) +
  theme(legend.title = element_blank(),
        legend.position="top")
```

```{r fig.width=4, fig.height=12, message=FALSE, warning=FALSE}
ordered_levels <-   
  today_dat %>% 
  group_by(state) %>% 
  mutate(mean_z = mean(YY_deaths)) %>%
  arrange(mean_z) %>% 
  ungroup() %>%
  select(state) %>% 
  unique() %>% 
  unlist()
today_dat %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(state = factor(state, levels = ordered_levels)) %>% 
  mutate(urban = factor(recode(XX_urban,
                        `0` = "rural",
                        `1` = "urban"
                        ))
         ) %>% 
  ggplot(aes(log(YY_deaths), state, fill = urban)) +
      geom_density_ridges2(alpha = 0.5) +
      labs(x = "number of deaths (log2)") +
     theme(legend.title = element_blank(),
           legend.position="top")

linbrary(G)
```

```{r}
today_dat %>% 
   group_by(state) %>% 
   summarize(deaths  = sum(YY_deaths)) %>%

   ggplot(aes(reorder(state, deaths), deaths)) +
     geom_col() +
     scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
     labs(y = NULL)

```

```{r}
today_dat %>% 
   group_by(date) %>% 
   mutate(tot = sum(YY_deaths)) %>% 
   group_by(state) %>%
   summarize(perc_death = round((sum(YY_deaths)/first(tot) * 100), 2)) %>% 
   arrange(desc(perc_death))
```

```{r}
today_dat %>% 
   group_by(date) %>% 
   mutate(tot = sum(NC_cases)) %>% 
   group_by(state) %>%
   summarize(perc_cases = round((sum(NC_cases)/first(tot) * 100),2)) %>% 
   arrange(desc(perc_cases))
```

```{r}
options(scipen = 999)
today_dat %>% 
   group_by(state) %>% 
   summarize(cases  = sum(NC_cases)) %>%

   ggplot(aes(reorder(state, cases), cases)) +
     geom_col() +
     scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
    labs(y = NULL)

```

```{r fig.width=4, fig.height=9}
today_dat %>% 
  group_by(state) %>% 
  summarize(temp_winter = mean(XX_winter_hum, na.rm = TRUE),
            n = n()
            ) %>%

  ggplot(aes(reorder(state, temp_winter), temp_winter)) +
    geom_col() +
   # scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90)) +
   geom_label(aes(label = n,  y = 25)) +
   coord_flip()

```

```{r fig.width=4, fig.height=9}
today_dat %>% 
  group_by(state) %>% 
  summarize(temp_summer = mean(XX_summer_hum, na.rm = TRUE),
            n = n()
            ) %>%

  ggplot(aes(reorder(state, temp_summer), temp_summer)) +
    geom_col() +
   # scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90)) +
   geom_label(aes(label = n,  y = 25)) +
   coord_flip()

```

```{r}
today_dat %>% 
  group_by(state) %>% 
  summarize(temp_summer = mean(XX_summer_hum, na.rm = TRUE),
            n = n()) %>%

  ggplot(aes(reorder(state, temp_summer), temp_summer)) +
    geom_col() +
   scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90))

```

```{r}
today_dat %>% 
  filter(state == "Oregon") %>% 
  select(fips, XX_summer_hum) %>% 
  dput()
```

```{r}
today_dat %>% 
  group_by(state) %>% 
  rename(temp_winter = XX_winter_hum) %>%

  ggplot(aes(reorder(state, temp_winter), temp_winter, color = state)) +
    geom_jitter() +
   scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 90)) +
   theme(legend.position = "none")

```

```{r}
library(vroom)
us_season <- vroom("https://raw.githubusercontent.com/wxwx1993/PM_COVID/master/Data/temp_seasonal_county.csv",
  col_types = cols(.default = "d")
)

summary(us_season$winter_rmax)

```

### 2020-07-01 R/getus_functions.R

@@ -30,7 +30,7 @@ getus_covid_jhu <- function() {
       dplyr::select(
         .data$date, .data$combined_key, .data$fips, .data$value
       ) %>%
-
+      dplyr::mutate(fips = as.numeric(fips)) %>%
       # this is to separte the column (tidyr::separate occasionally trows error)
       dplyr::mutate("county_state" = gsub(",(.)?US$", "", .data$combined_key, perl = TRUE)) %>%
       dplyr::mutate("state" = ifelse(

### 2020-07-01 notebooks/old_notebooks/all_analysis.Rmd

---
title: "US model"
date: "`r Sys.time()`"
author: Claudio Zanettini
categories:
  - covid-19
tags: 
  - covid-19
  - Italy
lastmod: "`r Sys.time()`"
keywords: []
description: ''
output:
  html_document:
    df_print: kable
    toc: true
    theme: journal
---

# US

## Get the data and functions

This is to get the data and clean them up.

```{r get_data3, message=FALSE, warning=FALSE}
# libraries --------------
# devtools::install_github("zeehio/facetscales")
library(facetscales)
library(covid19census)
library(tidyverse)
library(ggcorrplot)
library(GGally)
library(gganimate)
library(ggcorrplot)
library(broom)
library(purrr)
library(knitr)
library(kableExtra)
library(htmlwidgets)
library(plotly)
library(stringr)
theme_set(theme_bw())

logit = function(p){ log(p) - log(1-p) }

# var new to select -------
to_select<- c("date", "county", "state", "fips", "cases", "deaths", "perc_families", 
"perc_family_only_onep", "perc_edu_bachelor", 
"perc_withinternet", "total_pop", "perc_imm65", "total_imm65", 
"total_beds", "ratio_beds", "urban",  
"perc_alzheimer_dementia", "perc_asthma", "perc_atrial_fibrillation", 
"perc_cancer_breast", "perc_cancer_colorectal", "perc_cancer_lung", 
"perc_ch_obstructive_pulm", "perc_chronic_kidney_disease", "perc_depression", 
"perc_diabetes", "perc_heart_failure", "perc_hypertension", "perc_ischemic_heart_disease", "perc_obesity", "perc_osteoporosis", "perc_rheumatoid_arthritis", 
 "perc_stroke", "perc_tobacco_use", 
"median_income", "pm2.5", "summer_temp", "summer_hum", "winter_temp", 
"winter_hum", "age65_over", "median_age", "sex_ratio", "child_dependency", 
"perc_black", "perc_lat", "perc_white", "perc_asian", "perc_island", 
"perc_other", "perc_two_more_races", "total_tests", "days_f0"
)

# 
# # var to select old -------
# to_select <- c(
#   "date",
#   "county",
#   "state",
#   "fips",
#   "cases",
#   "deaths",
#   "perc_families",
#   "perc_familes_18childreen",
#   "perc_married_couples",
# 
#   "perc_family_only_onep",
#   # "perc_married_couples_u18ychildreen",
#   # "perc_families_only_male",
#   # "perc_families_only_male__18ychildreen",
#   # "perc_families_only_female",
#   # "perc_families_only_female__18ychildreen",
#   # "perc_non_families_alone",
#   # "perc_non_families_alone65y",
#   # "perc_non_families_u18y",
#   # "perc_non_families_65y",
#   # "perc_relationship_householder",
#   # "perc_relationship_spouse",
#   # "perc_relationship_child",
#   # "perc_relationship_other_relatives",
#   # "perc_relationship_other_nonrelatives",
#   # "perc_relationship_other_unmaried_part",
#   # "perc_marital_status_male_nevermaried",
#   # "perc_marital_status_male_maried",
#   # "perc_marital_status_male_separated",
#   # "perc_marital_status_male_widowed",
#   # "perc_marital_status_male_divorced",
#   # "perc_marital_status_female_nevermaried",
#   # "perc_marital_status_female_maried",
#   # "perc_marital_status_female_separated",
#   # "perc_marital_status_female_widowed",
#   # "perc_marital_status_female_divorced",
#   "perc_enrolled_preschool",
#   "perc_enrolled_kindergarden",
#   "perc_enrolled_elementary",
#   "perc_enrolled_highschool",
#   "perc_enrolled_college",
#   # "perc_edu_9grade",
#   # "perc_edu_nodiploma",
#   # "perc_edu_highshool",
#   "perc_edu_somecollege",
#   "perc_edu_associate",
#   "perc_edu_bachelor",
#   # "perc_edu_gradprofess",
#   # "perc_edu_highshool_higher",
#   # "perc_edu_batchelor_higher",
#   # "perc_american",
#   # "perc_arab",
#   # "perc_czech",
#   # "perc_danish",
#   # "perc_dutch",
#   # "perc_english",
#   # "perc_french_except_basque",
#   # "perc_french_canadian",
#   # "perc_german",
#   # "perc_greek",
#   # "perc_hungarian",
#   # "perc_irish",
#   # "perc_italian",
#   # "perc_lithuanian",
#   # "perc_norwegian",
#   # "perc_polish",
#   # "perc_portuguese",
#   # "perc_russian",
#   # "perc_scotch_irish",
#   # "perc_scottish",
#   # "perc_slovak",
#   # "perc_subsaharan_african",
#   # "perc_swedish",
#   # "perc_swiss",
#   # "perc_ukrainian",
#   # "perc_welsh",
#   # "perc_west_indian_excluding_hispanic_origin_groups",
#   # "perc_withcomputer",
#   "perc_withinternet",
#   "total_pop",
#   "perc_imm65",
#   "total_imm65",
#   "total_beds",
# 
#   "ratio_beds",
#   "urban",
#   "perc_acute_myocardial_infarction",
#   "perc_alzheimer_dementia",
#   "perc_asthma",
#   "perc_atrial_fibrillation",
#   "perc_cancer_breast",
#   "perc_cancer_colorectal",
#   "perc_cancer_lung",
#   "perc_cancer_all",
#   "perc_ch_obstructive_pulm",
#   "perc_chronic_kidney_disease",
#   "perc_depression",
#   "perc_diabetes",
#   "perc_heart_failure",
#   "perc_hypertension",
#   "perc_ischemic_heart_disease",
#   "perc_obesity",
#   "perc_osteoporosis",
#   "perc_rheumatoid_arthritis",
#   "perc_schizophrenia_psychotic_dis",
#   "perc_stroke",
#   "perc_tobacco_use",
#   # to many NA for these:
#   # "urgent_admission",
#   # "annual_wellness_visit",
#   # "elective_admission",
#   # "emergent_admission",
#   # "other_admission",
#   # perc_pneumococcal_vaccine HIGHLY CORRELATED WITH perc_imm65
#   # "perc_pneumococcal_vaccine",
#   # "perc_poverty",
#   "median_income",
#   "pm2.5",
#   "summer_temp",
#   "summer_hum",
#   "winter_temp",
#   "winter_hum",
#   "age65_over",
#   "median_age",
#   # "median_age_male",
#   # "median_age_female",
#   "sex_ratio",
#   # "age_dependency",
#   # "old_age_dependency",
#   "child_dependency",
#   "perc_black",
#   "perc_lat",
#   "perc_white",
# 
#   "perc_asian",
#   "perc_island",
#   "perc_other",
#   "perc_two_more_races",
#   "total_tests",
#   "days_f0"
# )
# 
#

# df1   -----------------------------
df1 <- getus_all()

# df2 ------------------------------
df2 <-
  df1 %>%
  # calculate age65_over and perc_black, white and latinos, total tests
  mutate(age65_over = rowSums(dplyr::select(., matches("perc_[6][5-9]|[7|8][0-9]")))) %>%
  mutate(urban = if_else(urban == "Urban", 1, 0)) %>%
  mutate(perc_black = total_black / total_pop * 100) %>%
  mutate(perc_white = total_white / total_pop * 100) %>%
  mutate(perc_lat = total_latino / total_pop * 100) %>%
  mutate(total_tests = positive + negative) %>%

  # calculate day since first case
  mutate(f_date = case_when(cases >= 1 ~ date)) %>%
  group_by(fips) %>%
  mutate(f_date = min(f_date, na.rm = TRUE), days_f0 = as.numeric(date - f_date)) %>%
  ungroup() %>%
  mutate(days_f0 = if_else(is.finite(days_f0), days_f0, NA_real_)) %>%

  # newly calculated-----------------------------------  
  # total beds in ratio
  mutate(ratio_beds = total_beds/total_pop) %>% 
  # asian are here
  mutate(perc_asian = total_asian / total_pop * 100) %>%
  mutate(perc_island = total_pacific_islander / total_pop * 100) %>%
  mutate(perc_native = total_native / total_pop * 100) %>%
  mutate(perc_other = total_other_race / total_pop * 100) %>%
  mutate(perc_two_more_races = total_two_more_races / total_pop * 100) %>%

  # family with one parent together
  mutate(perc_family_only_onep = perc_families_only_female + perc_families_only_male) %>% 
  mutate(total_imm65 = age65_over/100 *perc_imm65) %>%

  # select variables ----------
  dplyr::select(!!to_select) %>%

  # perc divided by 100
  mutate_at(vars(starts_with("perc")), function(x) x / 100) %>%
  mutate(age65_over = age65_over / 100) %>%

  # we remove NAs
  na.omit() %>%

  # at least 1 death
  filter(deaths >= 1)

# all_dat -------------
all_dat <- df2 %>%
  # make variables numeric
  mutate_at(vars(
    -date, -county, -state, -fips), 
     as.numeric
  ) %>%
  # rename
  # add XX_ to all the variables excepts those
  rename_at(
    vars(
      -date, -county, -state, -fips, -cases, -deaths,
      -perc_imm65, -cases, -total_pop, -total_imm65
    ),
    ~ paste0("XX_", .)
  ) %>%
  rename_at(vars(perc_imm65, total_imm65), ~ paste0("ZZ_", .)) %>%
  rename_at(vars(deaths), ~ paste0("YY_", .)) %>%
  rename_at(vars(cases), ~ paste0("NC_", .)) %>%
  rename_at(vars(total_pop), ~ paste0("NP_", .)) %>%
  # calculate logitZZ
  mutate(logitZZ_perc_imm65 = logit(ZZ_perc_imm65)) 
  # mutate(logitZZ_total_imm65 = log(ZZ_total_imm65) - log(1 - ZZ_total_imm65))

# today_dat ------------------------------
today_dat <- all_dat %>%
  filter(date == max(date))

# functions -----
logit = function(p){ log(p) - log(1-p) }

# functions ----------------------
#' get_PPScore
#'
#' Calculate propensity score and relative glm
#'
#' @param dat dataframe (need to some specific names ecc)
#' @param filt_cases threeshold of number of cases x county to be included in the analysis
#' @param form  formula to use for propensity score, this is a `reformulate` expression 
#' @param offset_f the offset  "log(NC_cases)"
#' @param var_dep dependent variable
#' @param var_int variable of interest
#' @param quant boolean, quantile of propensity score
#' @param include_model include a glm in the datafram
#' @param verbose boolean
#' @param nation variables to reports are different for italy and us. For example Italy doesn't have counties. We need to tell the function so it produced different outputs
#' @return dataframe
#'
get_PPScore <- function(
                        dat = .x, 
                        filt_cases = 1,
                        form = form_ps, 
                        offset_f = "log(NC_cases)",
                        var_dep =  "YY_deaths",
                        var_int = "ZZ_perc_imm65",
                        quant = FALSE,
                        include_model = FALSE,
                        verbose = TRUE,
                        nation = "us"
                        ){

# given a covid dataframe with 1 date, it calculate effect of the indipendent variable using a propensity score

  dat <- dat %>%
    filter(dat$NC_cases >= !!filt_cases)

  PropScoresLM <- lm(form, data = dat)

  if(quant == TRUE){
    dat[, "PP"] <- cut(fitted.values(PropScoresLM), 4, labels = FALSE)
  } else {
    dat[, "PP"] <- fitted.values(PropScoresLM)
  }

  form_glm <- reformulate(termlabels = c(var_int, "PP"), response = var_dep)

  outcome_PP <-
    # we add on the fly PP to the dataframe and we
    dat %>%
    glm(form_glm,
        family = quasipoisson(link = "log"),
        offset = eval(parse(text = offset_f)), 
        data = .
    )

  # conf.int = TRUE for confidence intervals
  res <- tidy(outcome_PP,  conf.int = TRUE) %>%
    filter(term == !!var_int) %>%
    mutate(
      deaths = sum(dat$YY_deaths),
      cases = sum(dat$NC_cases),
      formula =  deparse(formula(outcome_PP)),
      PP_form = paste0(deparse(form), collapse = ""),
      offset_f = offset_f,
      quant = FALSE
    )

  if(nation == "us") {
    res <- 
      res %>% 
      mutate(
      n_counties = nrow(dat),
      urban = sum(dat$XX_urban)
      )

  }

  if(include_model == TRUE) {
    res <- res %>% 
      mutate(model = list(outcome_PP))
  }

 if(verbose == TRUE){
       message(
          cat(paste0(
            "\nDate ",
            first(dat$date2),
            ":\nApplying model ",
            first(res$formula),
            "\noffset: ",
            first(offset_f),
            "\nfilt_cases = ",
            filt_cases,
            "!"
          ))
       )
}
 res
}

#' glm_pp
#'
#' Given a covid-19 dataframe, apply the propensity glm analysis to subset of dates, varying inclusion criteria 
#' per county
#'
#' @param dat dataframe (need to some specific names ecc)
#' @param filt_cases numeric vector of threasholds of cases X county  (default = 1)
#' @param form  formula to use for propensity score, this is a `reformulate` expression 
#' @param offset_f the offset, es "log(NC_cases)"
#' @param var_dep dependent variable, es. "YY_death"
#' @param var_int indipendent variable, es ZZ_perc_imm65"
#' @param data_term if TRUE keeps that dataframe of data used for the analysis
#' @param quant boolean, quantile of propensity score
#' @param include_model include a glm in the datafram
#' @param verbose boolean
#'
#' @return dataframe
#'

glm_pp <- function(dat, 
                   dates_tostudy = dates_tostudy, 
                   filt_cases,
                   offset_f = "log(NC_cases)",
                   var_dep,
                   var_int,
                   data_term = FALSE, 
                   quant = FALSE, 
                   include_model = FALSE,
                   verbose = TRUE,
                   ...
                   ){

# functions to apply glm ppcs model to diffenret dates in dataframe
# dat = covid dataframe 
# dates_tostudy = list of dates to perform the analysis on                    
# filt_cases =  numeric vector of threasholds of cases X county  (default = 1)
# form = formula to use for propensity score this is a `reformulate` expression for the pp score (form_ps)
# offset_f = the offset of glm function ("log(NC_cases)")
# var_dep =  dependent variable ("YY_death")
# var_int = variable of interest ("ZZ_perc_imm65"),
# quant = propensity score quantile (FALSE)

  suppressWarnings(
  zz_term <-   
  dat %>%
  # filter(YY_deaths > 0) %>% 
  filter(date %in% dates_tostudy) %>%
  mutate(date2 = date) %>% 
  nest(-date) %>%
  # we repeat the date slice X number of case threeshold that we want to study
  slice(rep(1:n(), each = length(filt_cases))) %>%
  mutate(filt_cases_c = rep(filt_cases, nrow(.)/length(filt_cases))) %>%

  mutate(fit = 
            map2(
            data,
            filt_cases_c,
            get_PPScore,
            form = form_ps, 
            offset_f = !!offset_f,
            quant = !!quant,
            var_int = !!var_int,
            var_dep = !!var_dep,
            include_model = !!include_model,
            verbose = !!verbose
            )
  ) %>% 
    unnest(fit))

  if(data_term == FALSE){
    zz_term_nd <- zz_term %>% 
    select(-data) 
    return(zz_term_nd)
  }else{
    return(zz_term)
    } 
}

ggcovid <- function(dat, x_var, filt_cases = 1, nation = "us") {

if(!x_var %in% c("date", "cases")) stop("x_var can be date or cases!")

if (x_var == "date"){
  to_add <- "; cases > 0" 
  aes_x <- "date"
  dat <- dat %>%
    filter(filt_cases_c == !!filt_cases) 
  x_title = NULL
}

if (x_var == "cases") {
  to_add <- paste0("; date ",  max(dat$date))
  aes_x <- "filt_cases_c"
  dat <- dat %>%
    filter(date == max(date))  
  x_title = "inclusion criteria (minimum cases)"

}

if (nation == "it") {
  var_facet <- c("estimate", "deaths", "cases")
}

if (nation == "us") {
  var_facet <-  c("estimate", "number of counties", "urban/counties")
}

# dependent variable used in title
dep_var <- stringr::word(first(dat$formula), 1)

  dat %>% 
    mutate(`urban/counties` = urban/n_counties) %>% 
    rename(`number of counties` = n_counties) %>% 
    pivot_longer(cols =   !!var_facet,
                 names_to = "var_name") %>%
    mutate(var_name = factor(var_name, levels = !!var_facet)) %>% 
    mutate(sign = if_else(p.value < 0.05, "p<0.05", "NS")) %>%
    mutate(sign = factor(sign, levels = c("p<0.05", "NS"))) %>% 
    # mutate(std.error = if_else(var_name == "estimate", std.error, 0)) %>%

    mutate_at(vars(std.error, conf.low, conf.high), ~if_else(var_name == "estimate", ., 0)) %>%

    ggplot(aes_string(aes_x, "value", color = "sign", group = "term")) +
    geom_line() +
    geom_point() +
    geom_linerange(aes(ymax = conf.high, ymin = conf.low), show.legend = FALSE) +
    facet_grid(vars(var_name), scales = "free") +
    scale_color_manual(values = c("NS" = "#00BFC4", "p<0.05" = "#F8766D"), drop = FALSE) +
    theme(
      # legend.position = "top",
      legend.title = element_blank(),
      legend.box.background = element_rect(colour = "black")
    ) +
    labs(
      title = paste0("Slope of ", first(dat$term), " over time"),
      subtitle = paste0(dep_var, "; offset = ", first(dat$offset_f),
                        ";\\nquantile PP = ", first(dat$quant), to_add),
      x = x_title
    ) +
    expand_limits(y = 0)

}

```

## Results

### Deaths ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)"

```{r  fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zz_term_deaths_np <- glm_pp(dat = all_dat,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE
       )

zz_term_deaths_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(zz_term_deaths_np, "date")

```

```{r fig.width=4, fig.height=4}
ggcovid(zz_term_deaths_np, "cases")
```

### Deaths ~ perc_imm65 + PP, offset_f = "log(NC_cases)"

```{r  fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zz_term_deaths_nc <- glm_pp(dat = all_dat,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zz_term_deaths_nc %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(zz_term_deaths_nc, "date")
```

```{r fig.width=4, fig.height=4}
ggcovid(zz_term_deaths_nc, "cases")
```

### Only New York: Deaths ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)"

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zzterm_onlyNY_deaths_np  <-

  all_dat %>% 
  filter(state == "New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zzterm_onlyNY_deaths_np  %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(zzterm_onlyNY_deaths_np, "date")
```

```{r fig.width=4, fig.height=4}
ggcovid(zzterm_onlyNY_deaths_np, "cases")
```

### Only New York: Deaths ~ perc_imm65 + PP, offset_f = "log(NC_cases)"

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "ZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zzterm_onlyNY_deaths_nc  <-

  all_dat %>% 
  filter(state =="New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zzterm_onlyNY_deaths_nc %>% 
    filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(zzterm_onlyNY_deaths_nc, "date")
```

```{r fig.width=4, fig.height=4}
ggcovid(zzterm_onlyNY_deaths_nc, "cases")
```

so the estimate is not exactly the same

```{r}
zzterm_onlyNY_deaths_nc %>% 
  filter(date == max(date)) %>% 
  select(estimate)
```

### No New York: Deaths ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)"

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zzterm_noNY_deaths_np  <- 
  all_dat %>% 
  filter(state !="New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zzterm_noNY_deaths_np  %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(zzterm_noNY_deaths_np, "date")
```

```{r fig.width=4, fig.height=4}
ggcovid(zzterm_noNY_deaths_np, "cases")
```

### No New York: Deaths ~ perc_imm65 + PP, offset_f = "log(NC_cases)"

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zzterm_noNY_deaths_nc  <-

  all_dat %>% 
  filter(state != "New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE
       )

zzterm_noNY_deaths_nc  %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(zzterm_noNY_deaths_nc, "date")
```

```{r fig.width=4, fig.height=4}
ggcovid(zzterm_noNY_deaths_nc, "cases")
```

With this we check the actual data to make sure that the function actually took all but NYS. We just take the last day and
we filter per counties > 0

```{r}
no_ny_check  <- 
  all_dat %>% 
  filter(state !="New York") %>% 
      glm_pp(
       dates_tostudy = max(all_dat$date),
       filt_cases = 1,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = TRUE,
       include_model = TRUE,
       verbose = FALSE
       )

data_no_ny_check <- no_ny_check$data[[1]]

unique(data_no_ny_check$state)

no_ny_check$formula

```

### NC_cases ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zz_term_cases_np <-

  all_dat %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form =  form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "NC_cases",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = TRUE,
       include_model = TRUE,
       verbose = TRUE
       )

 zz_term_cases_np  %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(zz_term_cases_np, "date")

```

```{r fig.width=4, fig.height=4}
ggcovid(zz_term_cases_np, "cases")
```

### No New York: NC_cases ~ perc_imm65 + PP, offset_f = "log(NP_total_pop)

```{r fig.width=4, fig.height=4}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds") ]

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zz_term_noNY_cases_np <-

  all_dat %>% 
    filter(state != "New York") %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "NC_cases",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = TRUE,
       include_model = TRUE,
       verbose = TRUE
       )

 zz_term_noNY_cases_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(zz_term_noNY_cases_np, "date")

```

### Positive control: pm2.5

```{r}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds", "XX_pm2.5")]

XX <- c("ZZ_perc_imm65", XX )

form_ps <- reformulate(termlabels = XX, response = "XX_pm2.5")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

pm2.5_deaths_np <-

      all_dat %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "XX_pm2.5",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE
       )

pm2.5_deaths_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(pm2.5_deaths_np, "date")

```

### Positive control Black

with logit doesnt fit

```{r}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds", "XX_perc_black")]

XX <- c("ZZ_perc_imm65", XX )

form_ps <- reformulate(termlabels = XX, response = "XX_perc_black")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- 1

black_deaths_np <- 
      all_dat %>% 
      mutate(logitXX_perc_black = logit(XX_perc_black)) %>% 
      glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "NC_cases",
       var_int = "XX_perc_black",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE
       )

black_deaths_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(black_deaths_np, "date")

```
### Positive control income ????

```{r}
XX <- grep("XX", names(today_dat), value = TRUE)
XX <- XX[!XX %in% c("XX_total_beds", "XX_median_income")]

XX <- c("ZZ_perc_imm65", XX )

form_ps <- reformulate(termlabels = XX, response = "XX_median_income")

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

income_deaths_np <-

  all_dat %>% 
  glm_pp(
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "XX_median_income",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE
       )

income_deaths_np %>% 
  filter(date == max(date), filt_cases_c == 1) %>% 
  select(model) %>%  
  flatten()

ggcovid(income_deaths_np, "date")

```

## Italy

### Deaths

```{r}
dat_it <- getit_all() %>% 
  mutate_at(vars(imm65, chronic_diab, chronic_bronc, chronic_hyp), function(x) x / 100) %>% 
  mutate(logitZZ_perc_imm65 = logit(imm65))

today_dat_it <-  dat_it %>% 
  filter(date == max(date))

dat_it_cl <- dat_it %>% 
  rename_at(vars(imm65), ~ paste0("ZZ_perc_", .)) %>% 
  rename_at(vars(pop_km2, lat, chronic_diab, chronic_bronc, chronic_hyp), ~paste0("XX_", .)) %>% 
  rename_at(vars(total_cases),  ~("NC_cases")) %>% 
  rename_at(vars(pop_tot), ~("NP_total_pop")) %>% 
  rename_at(vars(deaths), ~("YY_deaths"))

XX <- grep("XX", names(dat_it_cl), value = TRUE)
# XX <- XX[!XX %in% c("XX_total_beds") ]

dates_tostudy <- sort(unique(dat_it_cl$date))[52:length(unique(dat_it_cl$date))]
to_filt_cases  <- 1

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

perc_imm65_deahts_pop <- glm_pp(dat = dat_it_cl,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE,
       nation = "it"
       )

ggcovid(perc_imm65_deahts_pop, "date", nation = "it")
```

```{r}

XX <- grep("XX", names(dat_it_cl), value = TRUE)
# XX <- XX[!XX %in% c("XX_total_beds") ]

dates_tostudy <- sort(unique(dat_it_cl$date))[52:length(unique(dat_it_cl$date))]
to_filt_cases  <- 1

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

perc_imm65_deahts_cases <- glm_pp(dat = dat_it_cl,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NC_cases)",
       var_dep = "YY_deaths",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE,
       nation = "it"
       )

ggcovid(perc_imm65_deahts_cases, "date", nation = "it")
```

### Cases

```{r}
XX <- grep("XX", names(dat_it_cl), value = TRUE)
# XX <- XX[!XX %in% c("XX_total_beds") ]

dates_tostudy <- sort(unique(dat_it_cl$date))[52:length(unique(dat_it_cl$date))]
to_filt_cases  <- 1

form_ps <- reformulate(termlabels = XX, response = "logitZZ_perc_imm65")

perc_imm65_cases_cases <- glm_pp(dat = dat_it_cl,
       dates_tostudy = dates_tostudy,
       filt_cases = to_filt_cases,
       form = form_ps, 
       offset_f = "log(NP_total_pop)",
       var_dep = "NC_cases",
       var_int = "ZZ_perc_imm65",
       quant = FALSE,
       data_term = FALSE,
       include_model = TRUE,
       verbose = TRUE,
       nation = "it"
       )

ggcovid(perc_imm65_cases_cases, "date", nation = "it")

```

### 2020-07-01 notebooks/old_notebooks/logo.Rmd

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}

library(devtools)

# devtools::install_github("c1au6i0/covid19census")
library(kableExtra)
library(covid19census)

ff <- getus_all()

ff <- ff[1:50, 1:50]

names(ff) <- paste0("x", 1:50)

```

```{r}
library(dplyr)
ff %>% 
  kable(escape = F, align = "c") %>% 
  column_spec(1:50, bold = T, border_right = T) %>%
  kable_styling(latex_options = c("striped", "scale_down"), full_width = FALSE, font_size = 5)

```

```{r}
library(dplyr)
it <- getit_all()
glimpse(it)
```

```{r}
library(showtext)
library(hexSticker)
library(ggplot2)
library(png)
library(grid)
imgurl <- readPNG("/Users/heverz/Desktop/logo.png")
imgurl <- rasterGrob(imgurl, interpolate=TRUE)

font_add_google("Gloria Hallelujah", "gloria")
showtext_auto()

p <- sticker(imgurl,
          package="covid19census", p_size=5, p_y = 1.5, s_x = 1.1, s_y = 0.7, s_width=2.3, s_height=2.3, p_color = "#FF6A6A", p_family = "gochi",
          filename="hexsticker.png", h_color = "#FF6A6A", h_fill="white")
p
```

### 2020-07-01 notebooks/old_notebooks/us_model3.Rmd

---
title: 'Report on Interference: US data'
author: "C1au6i0 & Marchionni Lab"
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2:
    latex_engine: xelatex
    fig_caption: yes
    df_print: kable
header-includes:
 \usepackage{float}
  \floatplacement{figure}{H}
---

# Data Preprocessing

## **Summary**

  *  got the update data using `covid19census::getus_all(repo = "jhu")`
  *  calculated some percentage (`age65_over`, `perc_black`, `perc_white`, `perc_lat`), `total_tests` and, days from first case (`days_f0`)
  *  filtered for selected variables (the list `to_select` was populated by removing Pearson's correlations >= .8 in the same domain and across, *analysis not included*)
  *  percentages multiplied by 100, **filtered death >= 1** 
  *  add a prefix to variables:
      * `ZZ` = `perc_imm65`
      * `logitZZ` =  `perc_imm65` predicted probability: log(p/(1 − p))
      * `YY` = dependent variable `deaths`
      * `NC` = number of cases: `cases`
      * `NP` = population: `total_pop`
 *  filtered to get only the aggregated data (last date).

The dataframe `all_dat` lists all variables by date, whereas `today_dat` only the most recent endpoint.

https://data.census.gov/cedsci/table?q=race&tid=ACSDP1Y2018.DP05&hidePreview=true&moe=false&g=0500000US05031

## Get Data and Make Them Tidy

```{r libs, message=FALSE, warning=FALSE}
# devtools::install_github("c1au6i0/covid19census_dev", auth_token = "8ca5cd2bb556a45f1ffeb2e9b44c90d4fbef6cf2", ref = "dev")
# # devtools::install_github("zeehio/facetscales")

library(facetscales)
library(covid19census)
library(tidyverse)
library(ggcorrplot)
library(caret)
library(GGally)
library(gganimate)
library(ggcorrplot)
library(ggdendro)
library(broom)
library(purrr)
library(knitr)
library(kableExtra)
library(htmlwidgets)
library(plotly)
theme_set(theme_bw())
```

This is the list of variable that we are going to select.

```{r to_select_b}
to_select <- c(
  "date",
  "county",
  "state",
  "fips",
  "cases",
  "deaths",
  "perc_families",
  "perc_familes_18childreen",
  "perc_married_couples",
  # "perc_married_couples_u18ychildreen",
  "perc_families_only_male",
  # "perc_families_only_male__18ychildreen",
  "perc_families_only_female",
  # "perc_families_only_female__18ychildreen",
  # "perc_non_families_alone",
  # "perc_non_families_alone65y",
  # "perc_non_families_u18y",
  # "perc_non_families_65y",
  # "perc_relationship_householder",
  # "perc_relationship_spouse",
  # "perc_relationship_child",
  # "perc_relationship_other_relatives",
  # "perc_relationship_other_nonrelatives",
  # "perc_relationship_other_unmaried_part",
  # "perc_marital_status_male_nevermaried",
  # "perc_marital_status_male_maried",
  # "perc_marital_status_male_separated",
  # "perc_marital_status_male_widowed",
  # "perc_marital_status_male_divorced",
  # "perc_marital_status_female_nevermaried",
  # "perc_marital_status_female_maried",
  # "perc_marital_status_female_separated",
  # "perc_marital_status_female_widowed",
  # "perc_marital_status_female_divorced",
  "perc_enrolled_preschool",
  "perc_enrolled_kindergarden",
  "perc_enrolled_elementary",
  "perc_enrolled_highschool",
  "perc_enrolled_college",
  # "perc_edu_9grade",
  # "perc_edu_nodiploma",
  # "perc_edu_highshool",
  "perc_edu_somecollege",
  "perc_edu_associate",
  "perc_edu_bachelor",
  # "perc_edu_gradprofess",
  # "perc_edu_highshool_higher",
  # "perc_edu_batchelor_higher",
  "perc_american",
  "perc_arab",
  "perc_czech",
  "perc_danish",
  "perc_dutch",
  "perc_english",
  "perc_french_except_basque",
  "perc_french_canadian",
  "perc_german",
  "perc_greek",
  "perc_hungarian",
  "perc_irish",
  "perc_italian",
  "perc_lithuanian",
  "perc_norwegian",
  "perc_polish",
  "perc_portuguese",
  "perc_russian",
  "perc_scotch_irish",
  "perc_scottish",
  "perc_slovak",
  "perc_subsaharan_african",
  "perc_swedish",
  "perc_swiss",
  "perc_ukrainian",
  "perc_welsh",
  "perc_west_indian_excluding_hispanic_origin_groups",
  # "perc_withcomputer",
  "perc_withinternet",
  "total_pop",
  "perc_imm65",
  "total_beds",
  "urban",
  "perc_acute_myocardial_infarction",
  "perc_alzheimer_dementia",
  "perc_asthma",
  "perc_atrial_fibrillation",
  "perc_cancer_breast",
  "perc_cancer_colorectal",
  "perc_cancer_lung",
  "perc_cancer_all",
  "perc_ch_obstructive_pulm",
  "perc_chronic_kidney_disease",
  "perc_depression",
  "perc_diabetes",
  "perc_heart_failure",
  "perc_hypertension",
  "perc_ischemic_heart_disease",
  "perc_obesity",
  "perc_osteoporosis",
  "perc_rheumatoid_arthritis",
  "perc_schizophrenia_psychotic_dis",
  "perc_stroke",
  "perc_tobacco_use",
  # to many NA for these:
  # "urgent_admission",
  # "annual_wellness_visit",
  # "elective_admission",
  # "emergent_admission",
  # "other_admission",
  # perc_pneumococcal_vaccine HIGHLY CORRELATED WITH perc_imm65
  # "perc_pneumococcal_vaccine",
  # "perc_poverty",
  "median_income",
  "pm2.5",
  "summer_temp",
  "summer_hum",
  "winter_temp",
  "winter_hum",
  "age65_over",
  "median_age",
  # "median_age_male",
  # "median_age_female",
  "sex_ratio",
  # "age_dependency",
  # "old_age_dependency",
  "child_dependency",
  "perc_black",
  "perc_lat",
  "perc_white",
  "total_tests",
  "days_f0"
)
```

We get the data and preprocess them as described in the summary.

```{r warning=FALSE}
df1 <- getus_all()

df2 <-
  df1 %>%
  # calculate age65_over and perc_black, white and latinos, total tests
  mutate(age65_over = rowSums(dplyr::select(., matches("perc_[6][5-9]|[7|8][0-9]")))) %>%
  mutate(urban = if_else(urban == "Urban", 1, 0)) %>%
  mutate(perc_black = total_black / total_pop * 100) %>%
  mutate(perc_white = total_white / total_pop * 100) %>%
  mutate(perc_lat = total_latino / total_pop * 100) %>%
  mutate(total_tests = positive + negative) %>%

  # calculate day since first case
  mutate(f_date = case_when(cases >= 1 ~ date)) %>%
  group_by(fips) %>%
  mutate(f_date = min(f_date, na.rm = TRUE), days_f0 = as.numeric(date - f_date)) %>%
  ungroup() %>%
  mutate(days_f0 = if_else(is.finite(days_f0), days_f0, NA_real_)) %>%

  # select variables
  dplyr::select(!!to_select) %>%

  # perc divided by 100
  mutate_at(vars(starts_with("perc")), function(x) x / 100) %>%
  mutate(age65_over = age65_over / 100) %>%

  # we remove NAs
  na.omit() %>%

  # at least 1 death
  filter(deaths >= 1)

all_dat <- df2 %>%
  # make variables numeric
  mutate_at(vars(
    -date, -county, -state, -fips, -cases, -deaths,
    -perc_imm65, -cases, -total_pop
  ), as.numeric) %>%
  # rename
  # add XX_ to all the variables excepts those
  rename_at(
    vars(
      -date, -county, -state, -fips, -cases, -deaths,
      -perc_imm65, -cases, -total_pop
    ),
    ~ paste0("XX_", .)
  ) %>%
  rename_at(vars(perc_imm65), ~ paste0("ZZ_", .)) %>%
  rename_at(vars(deaths), ~ paste0("YY_", .)) %>%
  rename_at(vars(cases), ~ paste0("NC_", .)) %>%
  rename_at(vars(total_pop), ~ paste0("NP_", .)) %>%
  # calculate logitZZ
  mutate(logitZZ = log(ZZ_perc_imm65) - log(1 - ZZ_perc_imm65))

today_dat <- all_dat %>%
  filter(date == max(date))

dim(today_dat)
```

# Models

## **Summary of Results of Modeling**

## Propensity Score Analysis

First, we calculate the propensity score.

```{r prop_score_formula}
# with this we create the formula:
# we get all the variables that start with XX
XX <- grep("XX", names(today_dat), value = TRUE)

# XX <- XX[XX != "XX_total_tests" ]
# XX <- c("NP_total_pop", XX)

form_ps <- reformulate(termlabels = XX, response = "logitZZ")

# in lm we input the formula
PropScoresLM <- lm(form_ps, data = today_dat)
tidy(PropScoresLM) %>%
  mutate(term = str_sub(term, 1, 27))
#
# %>%
#   kable( "latex", booktabs = TRUE, longtable = TRUE) %>%
#   kable_styling(latex_options = c("hold_position", "repeat_header"))
```

Then, we use the propensity score in the linear model using `YY_deaths` as response.

```{r model_cases}
outcomePP_cases <-
  today_dat %>%
  # we add on the fly PP to the dataframe
  mutate(PP = fitted.values(PropScoresLM)) %>%
  glm(YY_deaths ~ ZZ_perc_imm65 + PP,
    family = quasipoisson(link = "log"),
    offset = log(NC_cases), data = .
  )

tidy(outcomePP_cases)
```

**How did the slope of the `ZZ_perc_imm65` change overtime?**

The code below is used to fit the model to data obtained at different endpoints. The idea behind this analysis is to get a sense of how stable is the effect observed (it could well be the result of a fluke in today data).

```{r  message=FALSE, warning=FALSE}
# we create a function that we apply to the nests
get_PPScore <- function(dat = .x, filt_cases = 1, form = form_ps, offset_f = "log(NC_cases)") {

  dat <- dat %>%
    filter(dat$NC_cases >= !!filt_cases)

  PropScoresLM <- lm(form, data = dat)

  outcome_PP <-
    # we add on the fly PP to the dataframe and we
    dat %>%
    mutate(PP = fitted.values(PropScoresLM)) %>%
    glm(YY_deaths ~ ZZ_perc_imm65 + PP,
      family = quasipoisson(link = "log"),
      offset = eval(parse(text = offset_f)), data = .
    )

  res <- tidy(outcome_PP) %>%
    filter(term == "ZZ_perc_imm65") %>%
    mutate(
      deaths = sum(dat$NC_cases),
      n_counties = nrow(dat),
      urban = sum(dat$XX_urban)
    )

 message(
   "Applying model with filt_cases = ",
   paste0(filt_cases))

 res
}

# Here we select some dates, we nest the dataframe and we
# use mutate to apply the model to each nest subset.
# This is the equivalent of splitting a dataframe by date, lapply
# the formula to the list of dataframes, and cbind the results.
# dplyr + broom

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- seq(1,100, 1)

zz_term <-   
  all_dat %>%
  filter(YY_deaths > 0) %>% 
  filter(date %in% dates_tostudy) %>%
  nest(-date) %>%
  slice(rep(1:n(), each = length(to_filt_cases))) %>% 
  mutate(filt_cases_c = rep(to_filt_cases, nrow(.)/length(to_filt_cases))) %>% 
  mutate(fit = 
            map2(
            data,
            filt_cases_c,
            get_PPScore,
            form = form_ps, 
            offset_f = "log(NC_cases)"
            )
  ) %>% 
    unnest(fit)

# we keep a version without data that takes a lot of space
zz_term_nd <- zz_term %>% 
  select(-data)
```

```{r pscases, fig.cap="propensity score, NC_cases as offset", fig.width=4, fig.height=4, warning=FALSE}  
zz_term_nd %>%
  filter(filt_cases_c == 1) %>% 
  mutate(`urban/counties` = urban/n_counties) %>% 
  rename(`number of counties` = n_counties) %>% 
  pivot_longer(cols = 
                 c("estimate", "number of counties", "urban/counties"),
               names_to = "var_name") %>%
  mutate(sign = if_else(p.value < 0.05, "p<0.05", "NS")) %>%
  mutate(std.error = if_else(var_name == "estimate", std.error, 0)) %>%

  ggplot(aes(date, value, color = sign, group = term)) +
      geom_line() +
      geom_point() +
      geom_linerange(aes(ymax = value + std.error, ymin = value - std.error), show.legend = FALSE) +
      scale_color_manual(values = c("#00BFC4", "#F8766D")) +
      facet_grid(vars(var_name), scales = "free") +
      theme( 
        # legend.position = "top",
        legend.title = element_blank(),
        legend.box.background = element_rect(colour = "black")
      ) +
      labs(
        title = "Slope of ZZ_perc_imm65 over time",
        subtitle = "cases > 0, propensity score  using number of cases as offset"
      )
```

The var `estimate` represents the slope of `ZZ_perc_imm65`, whereas `n_counties` is the number of counties included in the analysis (only counties with at least one death can be included). The color of the points indicates if slope was significantly different from 0.
Before Apr 20, slope of `ZZ_perc_imm65` was occasionally significant but pattern was very erratic.

```{r filt_graph. fig.width=4, fig.height=4}
zz_term_nd %>%
  filter(date == max(date)) %>% 
  mutate(`urban/counties` = urban/n_counties) %>% 
  rename(`number of counties` = n_counties) %>% 
  pivot_longer(cols = 
                 c("estimate", "number of counties", "urban/counties"),
               names_to = "var_name") %>%
  mutate(sign = if_else(p.value < 0.05, "p<0.05", "NS")) %>%
  mutate(std.error = if_else(var_name == "estimate", std.error, 0)) %>%

  ggplot(aes(filt_cases_c, value, color = sign, group = term)) +
      geom_line() +
      geom_point() +
      scale_color_manual(values = c("NS" = "#00BFC4", "p<0.05" = "#F8766D")) +
      geom_linerange(aes(ymax = value + std.error, ymin = value - std.error), show.legend = FALSE) +
      facet_grid(vars(var_name), scales = "free") +
      theme(
        # legend.position = "top",
        legend.title = element_blank(),
        legend.box.background = element_rect(colour = "black")
      ) +
      labs(
        title = paste0(max(zz_term_nd$date),": Slope of ZZ_perc_imm65"),
        subtitle = "propensity score  using number of cases as offset",
        x = "minimum cases per county to be included in analysis" 
      )
```

```{r}
today_dat %>% 
  filter(!is.na(XX_urban)) %>% 
  ggplot(aes(XX_urban, YY_deaths, col = state, cases = NC_cases, county = county)) +
  theme(legend.position = "none") +
    geom_jitter()
```
```{r}
today_dat %>% 
  filter(!is.na(XX_urban)) %>% 
  mutate(XX_urban = recode(
    XX_urban,
    `1` = "urban",
    `0` = "rural"
    )) %>% 
  ggplot(aes(XX_urban, ZZ_perc_imm65, group = XX_urban)) +
  theme(legend.position = "none") +
  scale_x_discrete() +
  geom_boxplot() +
  labs(
    y = "Vaccination coverage in people age 65 or older",
    x = NULL
    )

```

```{r}
(ggplot(today_dat,aes(ZZ_perc_imm65, YY_deaths, colour = XX_urban, county = county, shapes = XX_urban)) +
    theme(legend.position = "none") +
    geom_jitter()) %>% 
  ggplotly()
```

We make an animation for doing the same analysis on all dates.

```{r filt_graph}
zz_term_nd %>%
  mutate(`ratio urban counties` = urban/n_counties) %>% 
  rename(`number of counties` = n_counties) %>% 
  pivot_longer(cols = 
                 c("estimate", "number of counties", "ratio urban counties"),
               names_to = "var_name") %>%
  mutate(sign = if_else(p.value < 0.05, "p<0.05", "NS")) %>%
  mutate(std.error = if_else(var_name == "estimate", std.error, 0)) %>%

  ggplot(aes(filt_cases_c, value, color = sign, group = term)) +
      geom_line() +
      geom_point() +
      scale_color_manual(values = c("#00BFC4", "#F8766D")) +
        geom_linerange(aes(ymax = value + std.error, ymin = value - std.error), show.legend = FALSE) +
      facet_grid(vars(var_name), scales = "free") +
      theme(
        # legend.position = "top",
        legend.title = element_blank(),
        legend.box.background = element_rect(colour = "black")
      ) +
  # gganimate part
      labs(
        title = 'Slope of ZZ_perc_imm65, date results: {current_frame}',
        subtitle = "propensity score  using number of cases as offset",
        x = "Number of cases to be included in analysis"
      ) +
   transition_manual(date)

anim_save("filth_death.gif")
```

```{r pspop, fig.cap="propensity score, NC_cases as offset", fig.width=4, fig.height=4, warning=FALSE}

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- seq(1,100, 1)

zz_term2 <-   
  all_dat %>%
  filter(YY_deaths > 0) %>% 
  filter(date %in% dates_tostudy) %>%
  nest(-date) %>%
  slice(rep(1:n(), each = length(to_filt_cases))) %>% 
  mutate(filt_cases_c = rep(to_filt_cases, nrow(.)/length(to_filt_cases))) %>% 
  mutate(fit = 
            map2(
            data,
            filt_cases_c,
            get_PPScore,
            form = form_ps, 
            offset_f = "log(NP_total_pop)"
            )
  ) %>% 
    unnest(fit)

# we keep a version without data that takes a lot of space
zz_term_nd2 <- zz_term2 %>% 
  select(-data)

zz_term_nd2 %>%
  filter(filt_cases_c == 1) %>% 
  mutate(`urban/counties` = urban/n_counties) %>% 
  rename(`number of counties` = n_counties) %>% 
  pivot_longer(cols = 
                 c("estimate", "number of counties", "urban/counties"),
               names_to = "var_name") %>%
  mutate(sign = if_else(p.value < 0.05, "p<0.05", "NS")) %>%
  mutate(std.error = if_else(var_name == "estimate", std.error, 0)) %>%

  ggplot(aes(date, value, color = sign, group = term)) +
      geom_line() +
      geom_point() +
      geom_linerange(aes(ymax = value + std.error, ymin = value - std.error), show.legend = FALSE) +
      scale_color_manual(values = c("#00BFC4", "#F8766D")) +
      facet_grid(vars(var_name), scales = "free") +
      theme( 
        # legend.position = "top",
        legend.title = element_blank(),
        legend.box.background = element_rect(colour = "black")
      ) +
      labs(
        title = "Slope of ZZ_perc_imm65 over time",
        subtitle = "cases > 0, propensity score  using number of cases as offset"
      )
```

# Let's redo everything using variable as indicated

## Get Data

## Sanity Checks

### Race

```{r fig.height=7, fig.width=5}

reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

df1 %>% 
  filter(date == max(date)) %>% 
  group_by(state) %>% 
  summarize(
    state_pop = sum(total_pop, na.rm = TRUE),
    perc_black = sum(total_black, na.rm = TRUE) / state_pop  * 100,
    perc_white = sum(total_white, na.rm = TRUE) / state_pop  * 100,
    perc_lat = sum(total_latino, na.rm = TRUE) / state_pop  * 100,
    perc_asian = sum(total_asian, na.rm = TRUE) / state_pop  * 100,
    perc_island = sum(total_pacific_islander, na.rm = TRUE) / state_pop  * 100,
    perc_native = sum(total_native, na.rm = TRUE) / state_pop  * 100,
    perc_other = sum(total_other_race, na.rm = TRUE) / state_pop  * 100,
    perc_two_more_races = sum(total_two_more_races, na.rm = TRUE) / state_pop  * 100) %>% 
  arrange(perc_asian) %>% 
  ungroup(state) %>%

  pivot_longer(
    cols = 
      c("perc_black",
        "perc_lat",
        "perc_white",
        "perc_asian",
        "perc_island",
        "perc_native",
        "perc_other",
        "perc_lat"

        ), 
    names_to = "var",
    values_to = "value"
    ) %>% 
  filter(state != "District of Columbia") %>%

  ggplot(aes(reorder_within(state, value, value), value)) +
    geom_col() +
    facet_wrap(vars(var), nrow = 4, scales = "free") +
    scale_x_reordered(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
    labs(x = NULL, y = NULL)

```

NOW WE ARE TALKING

```{r}
 df1 %>% 
  filter(date == max(date)) %>% 
  group_by(state) %>% 
  summarize(
    state_pop = sum(total_pop, na.rm = TRUE),
    perc_black = sum(total_black, na.rm = TRUE) / state_pop  * 100,
    perc_white = sum(total_white, na.rm = TRUE) / state_pop  * 100,
    perc_lat = sum(total_latino, na.rm = TRUE) / state_pop  * 100,
    perc_asian = sum(total_asian, na.rm = TRUE) / state_pop  * 100,
    perc_island = sum(total_pacific_islander, na.rm = TRUE) / state_pop  * 100,
    perc_native = sum(total_native, na.rm = TRUE) / state_pop  * 100,
    perc_other = sum(total_other_race, na.rm = TRUE) / state_pop  * 100,
    perc_two_more_races = sum(total_two_more_races, na.rm = TRUE) / state_pop  * 100) %>%

 mutate(
   tot_race = perc_black  + 
         perc_white + 
         perc_asian + 
         perc_island +
         perc_native +
         perc_other +
        perc_two_more_races
   ) %>% 
  select(state, 
         tot_race, 
         perc_black  , 
         perc_white , 
         perc_asian , 
         perc_island ,
         perc_native ,
         perc_other ,
        perc_two_more_races
         )
```

https://data.census.gov/cedsci/table?q=race&tid=ACSDP1Y2018.DP05&hidePreview=true&moe=false&g=0500000US05031

** Population OK **

```{r}
options(scipen = 999)

df2 %>% 
  ggplot(aes(state, total_pop/10^6)) +
   geom_boxplot() +
   scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45))
```

If we add total latino + total not latinos and we divided by total_pop we get 1...so
this is good

### Hospital Beds

```{r fig.height=6, fig.width=5}
df2 %>% 
  filter(date == max(date)) %>% 
  group_by(state) %>% 
  summarize(
    total_pop_thousand = sum(total_pop, na.rm = TRUE)/1000,
    total_beds = sum(total_beds, na.rm = TRUE),
    ratio_beds = total_beds / (total_pop_thousand * 1000)
    ) %>% 
  arrange(ratio_beds) %>% 
  mutate( state = factor(state, levels = state)) %>%

  pivot_longer(
    cols = c("total_pop_thousand", "total_beds", "ratio_beds"), 
    names_to = "var",
    values_to = "value"
    ) %>% 
  mutate(var = factor(var, levels = c("total_pop_thousand", "total_beds", "ratio_beds"))) %>%

  ggplot(aes(state, value)) +
    geom_col() +
    facet_wrap(vars(var), nrow = 3, scales = "free_y") +
    scale_x_discrete(guide = guide_axis(check.overlap = TRUE, angle = 45)) +
    labs(x = NULL)
```

What is the take home message? That regardless, we are going to have some correlation. So that is why we need to do a propensity score analysis

## PP

```{r fig.width=10, fig.high=10}
today_dat %>% 
  select(starts_with("XX")) %>% 
  cor(use ="pairwise.complete.obs") %>% 
  ggcorrplot(hc.order = TRUE, tl.cex = 7)

```

# Re Analysis No Functions

```{r get_data2, message=FAlSE, warning=FALSE}

# libraries --------------
# devtools::install_github("zeehio/facetscales")
library(facetscales)
library(covid19census)
library(tidyverse)
library(ggcorrplot)
library(caret)
library(GGally)
library(gganimate)
library(ggcorrplot)
library(ggdendro)
library(broom)
library(purrr)
library(knitr)
library(kableExtra)
library(htmlwidgets)
library(plotly)
theme_set(theme_bw())

# var to select -------
to_select <- c(
  "date",
  "county",
  "state",
  "fips",
  "cases",
  "deaths",
  "perc_families",
  "perc_familes_18childreen",
  "perc_married_couples",

  "perc_family_only_onep",
  # "perc_married_couples_u18ychildreen",
  # "perc_families_only_male",
  # "perc_families_only_male__18ychildreen",
  # "perc_families_only_female",
  # "perc_families_only_female__18ychildreen",
  # "perc_non_families_alone",
  # "perc_non_families_alone65y",
  # "perc_non_families_u18y",
  # "perc_non_families_65y",
  # "perc_relationship_householder",
  # "perc_relationship_spouse",
  # "perc_relationship_child",
  # "perc_relationship_other_relatives",
  # "perc_relationship_other_nonrelatives",
  # "perc_relationship_other_unmaried_part",
  # "perc_marital_status_male_nevermaried",
  # "perc_marital_status_male_maried",
  # "perc_marital_status_male_separated",
  # "perc_marital_status_male_widowed",
  # "perc_marital_status_male_divorced",
  # "perc_marital_status_female_nevermaried",
  # "perc_marital_status_female_maried",
  # "perc_marital_status_female_separated",
  # "perc_marital_status_female_widowed",
  # "perc_marital_status_female_divorced",
  "perc_enrolled_preschool",
  "perc_enrolled_kindergarden",
  "perc_enrolled_elementary",
  "perc_enrolled_highschool",
  "perc_enrolled_college",
  # "perc_edu_9grade",
  # "perc_edu_nodiploma",
  # "perc_edu_highshool",
  "perc_edu_somecollege",
  "perc_edu_associate",
  "perc_edu_bachelor",
  # "perc_edu_gradprofess",
  # "perc_edu_highshool_higher",
  # "perc_edu_batchelor_higher",
  # "perc_american",
  # "perc_arab",
  # "perc_czech",
  # "perc_danish",
  # "perc_dutch",
  # "perc_english",
  # "perc_french_except_basque",
  # "perc_french_canadian",
  # "perc_german",
  # "perc_greek",
  # "perc_hungarian",
  # "perc_irish",
  # "perc_italian",
  # "perc_lithuanian",
  # "perc_norwegian",
  # "perc_polish",
  # "perc_portuguese",
  # "perc_russian",
  # "perc_scotch_irish",
  # "perc_scottish",
  # "perc_slovak",
  # "perc_subsaharan_african",
  # "perc_swedish",
  # "perc_swiss",
  # "perc_ukrainian",
  # "perc_welsh",
  # "perc_west_indian_excluding_hispanic_origin_groups",
  # "perc_withcomputer",
  "perc_withinternet",
  "total_pop",
  "perc_imm65",
  "total_imm65",
  "total_beds",

  "ratio_beds",
  "urban",
  "perc_acute_myocardial_infarction",
  "perc_alzheimer_dementia",
  "perc_asthma",
  "perc_atrial_fibrillation",
  "perc_cancer_breast",
  "perc_cancer_colorectal",
  "perc_cancer_lung",
  "perc_cancer_all",
  "perc_ch_obstructive_pulm",
  "perc_chronic_kidney_disease",
  "perc_depression",
  "perc_diabetes",
  "perc_heart_failure",
  "perc_hypertension",
  "perc_ischemic_heart_disease",
  "perc_obesity",
  "perc_osteoporosis",
  "perc_rheumatoid_arthritis",
  "perc_schizophrenia_psychotic_dis",
  "perc_stroke",
  "perc_tobacco_use",
  # to many NA for these:
  # "urgent_admission",
  # "annual_wellness_visit",
  # "elective_admission",
  # "emergent_admission",
  # "other_admission",
  # perc_pneumococcal_vaccine HIGHLY CORRELATED WITH perc_imm65
  # "perc_pneumococcal_vaccine",
  # "perc_poverty",
  "median_income",
  "pm2.5",
  "summer_temp",
  "summer_hum",
  "winter_temp",
  "winter_hum",
  "age65_over",
  "median_age",
  # "median_age_male",
  # "median_age_female",
  "sex_ratio",
  # "age_dependency",
  # "old_age_dependency",
  "child_dependency",
  "perc_black",
  "perc_lat",
  "perc_white",

  "perc_asian",
  "perc_island",
  "perc_other",
  "perc_two_more_races",
  "total_tests",
  "days_f0"
)

# df1   -----------------------------
df1 <- getus_all()

# df2 ------------------------------
df2 <-
  df1 %>%
  # calculate age65_over and perc_black, white and latinos, total tests
  mutate(age65_over = rowSums(dplyr::select(., matches("perc_[6][5-9]|[7|8][0-9]")))) %>%
  mutate(urban = if_else(urban == "Urban", 1, 0)) %>%
  mutate(perc_black = total_black / total_pop * 100) %>%
  mutate(perc_white = total_white / total_pop * 100) %>%
  mutate(perc_lat = total_latino / total_pop * 100) %>%
  mutate(total_tests = positive + negative) %>%

  # calculate day since first case
  mutate(f_date = case_when(cases >= 1 ~ date)) %>%
  group_by(fips) %>%
  mutate(f_date = min(f_date, na.rm = TRUE), days_f0 = as.numeric(date - f_date)) %>%
  ungroup() %>%
  mutate(days_f0 = if_else(is.finite(days_f0), days_f0, NA_real_)) %>%

  # newly calculated-----------------------------------  
  # total beds in ratio
  mutate(ratio_beds = total_beds/total_pop) %>% 
  # asian are here
  mutate(perc_asian = total_asian / total_pop * 100) %>%
  mutate(perc_island = total_pacific_islander / total_pop * 100) %>%
  mutate(perc_native = total_native / total_pop * 100) %>%
  mutate(perc_other = total_other_race / total_pop * 100) %>%
  mutate(perc_two_more_races = total_two_more_races / total_pop * 100) %>%

  # family with one parent together
  mutate(perc_family_only_onep = perc_families_only_female + perc_families_only_male) %>% 
  mutate(total_imm65 = age65_over/100 * perc_imm65) %>%

  # select variables
  dplyr::select(!!to_select) %>%

  # perc divided by 100
  mutate_at(vars(starts_with("perc")), function(x) x / 100) %>%
  mutate(age65_over = age65_over / 100) %>%

  # we remove NAs
  na.omit() %>%

  # at least 1 death
  filter(deaths >= 1)

# all_dat -------------
all_dat <- df2 %>%
  # make variables numeric
  mutate_at(vars(
    -date, -county, -state, -fips), 
     as.numeric
  ) %>%
  # rename
  # add XX_ to all the variables excepts those
  rename_at(
    vars(
      -date, -county, -state, -fips, -cases, -deaths,
      -perc_imm65, -cases, -total_pop, -total_imm65
    ),
    ~ paste0("XX_", .)
  ) %>%
  rename_at(vars(perc_imm65), ~ paste0("ZZ_", .)) %>%
  rename_at(vars(total_imm65), ~ paste0("ZZ_", .)) %>%
  rename_at(vars(deaths), ~ paste0("YY_", .)) %>%
  rename_at(vars(cases), ~ paste0("NC_", .)) %>%
  rename_at(vars(total_pop), ~ paste0("NP_", .)) %>%
  # calculate logitZZ
  mutate(logitZZ_perc_imm65 = log(ZZ_perc_imm65) - log(1 - ZZ_perc_imm65)) %>% 
  # mutate(logitZZ_total_imm65 = log(ZZ_total_imm65) - log(1 - ZZ_total_imm65))

# today_dat ------------------------------
today_dat <- all_dat %>%
  filter(date == max(date))
```

```{r analysis_nofun, warning=FALSE}

# function
get_PPScore <- function(
                        dat = .x, 
                        filt_cases = 1,
                        # formula to use for propensity score
                        # this is a `reformulate` expression 
                        form = form_ps, 
                        # the offset of glm function
                        offset_f = "log(NC_cases)",
                        var_dep =  "YY_death",
                        # variable of interest
                        var_int = "ZZ_perc_imm65",
                        # propensity score quantile
                        quant = FALSE
                        ){

  dat <- dat %>%
    filter(dat$NC_cases >= !!filt_cases)

  PropScoresLM <- lm(form, data = dat)

  if(quant == TRUE){
    dat[, "PP"] <- cut(fitted.values(PropScoresLM), 4, labels = FALSE)
  } else {
    dat[, "PP"] <- fitted.values(PropScoresLM)
  }

  form_glm <- reformulate(termlabels = c(var_int, "PP"), response = var_dep)

  outcome_PP <-
    # we add on the fly PP to the dataframe and we
    dat %>%
    glm(YY_deaths ~ ZZ_perc_imm65 + PP,
        family = quasipoisson(link = "log"),
        offset = eval(parse(text = offset_f)), 
        data = .
    )

  res <- tidy(outcome_PP) %>%
    filter(term == "ZZ_perc_imm65") %>%
    mutate(
      deaths = sum(dat$NC_cases),
      n_counties = nrow(dat),
      urban = sum(dat$XX_urban)
    )

 message(
   "Applying model with filt_cases = ",
   paste0(filt_cases))

 res
}

# Here we select some dates, we nest the dataframe and we
# use mutate to apply the model to each nest subset.
# This is the equivalent of splitting a dataframe by date, lapply
# the formula to the list of dataframes, and cbind the results.
# dplyr + broom

dates_tostudy <- sort(unique(all_dat$date))[52:length(unique(all_dat$date))]
to_filt_cases  <- c(1, seq(10, 100, 10))

zz_term <-   
  all_dat %>%
  # filter(YY_deaths > 0) %>% 
  filter(date %in% dates_tostudy) %>%
  nest(-date) %>%
  # we repeat the date slice X number of case threeshold that we want to study
  slice(rep(1:n(), each = length(to_filt_cases))) %>%
  mutate(filt_cases_c = rep(to_filt_cases, nrow(.)/length(to_filt_cases))) %>% 
  mutate(fit = 
            map2(
            data,
            filt_cases_c,
            get_PPScore,
            form = form_ps, 
            offset_f = "log(NP_total_pop)",
            quant = FALSE
            )
  ) %>% 
    unnest(fit)

# we keep a version without data that takes a lot of space
zz_term_nd <- zz_term %>% 
  select(-data)
```

```{r pscases, fig.cap="propensity score, NC_cases as offset", fig.width=4, fig.height=4, warning=FALSE}  
zz_term_nd %>%
  filter(filt_cases_c == 1) %>% 
  mutate(`urban/counties` = urban/n_counties) %>% 
  rename(`number of counties` = n_counties) %>% 
  pivot_longer(cols = 
                 c("estimate", "number of counties", "urban/counties"),
               names_to = "var_name") %>%
  mutate(sign = if_else(p.value < 0.05, "p<0.05", "NS")) %>%
  mutate(std.error = if_else(var_name == "estimate", std.error, 0)) %>%

  ggplot(aes(date, value, color = sign, group = term)) +
      geom_line() +
      geom_point() +
      geom_linerange(aes(ymax = value + std.error, ymin = value - std.error), show.legend = FALSE) +
      facet_grid(vars(var_name), scales = "free") +
      scale_color_manual(values = c("NS" = "#00BFC4", "p<0.05" = "#F8766D")) +
      theme( 
        # legend.position = "top",
        legend.title = element_blank(),
        legend.box.background = element_rect(colour = "black")
      ) +
      labs(
        title = "Slope of ZZ_perc_imm65 over time",
        subtitle = "cases > 0, propensity score  using number of cases as offset"
      )
```

```{r filt_graph, fig.width=4, fig.height=4}
zz_term_nd %>%
  filter(date == max(date)) %>% 
  mutate(`urban/counties` = urban/n_counties) %>% 
  rename(`number of counties` = n_counties) %>% 
  pivot_longer(cols = 
                 c("estimate", "number of counties", "urban/counties"),
               names_to = "var_name") %>%
  mutate(sign = if_else(p.value < 0.05, "p<0.05", "NS")) %>%
  mutate(std.error = if_else(var_name == "estimate", std.error, 0)) %>%

  ggplot(aes(filt_cases_c, value, color = sign, group = term)) +
      geom_line() +
      geom_point() +
      scale_color_manual(values = c("NS" = "#00BFC4", "p<0.05" = "#F8766D")) +
      geom_linerange(aes(ymax = value + std.error, ymin = value - std.error), show.legend = FALSE) +
      facet_grid(vars(var_name), scales = "free") +
      theme(
        # legend.position = "top",
        legend.title = element_blank(),
        legend.box.background = element_rect(colour = "black")
      ) +
      labs(
        title = paste0(max(zz_term_nd$date),": Slope of ZZ_perc_imm65"),
        subtitle = "propensity score  using number of cases as offset",
        x = "minimum cases per county to be included in analysis" 
      )
```

### 2020-07-20 data-raw/import_raw.R
fixed typos

@@ -189,7 +189,7 @@ us_mmd <- mmd_all_us %>%
     other_admission,
     pneumococcal_vaccine
   )
-names(us_mmd) <-
+  names(us_mmd) <-
   c(
     "year",
     "fips",
@@ -909,9 +909,6 @@ usethis::use_data(
   overwrite = TRUE
 )
 #
-
-
-
 
 
 

### 2020-07-20 README.Rmd

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.0.9900-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.1.0000-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2020-07-21 README.Rmd

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.1.0000-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.1.1100-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2020-07-21 README.Rmd
minor fixes

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.1.1100-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.1.200-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2020-12-17 README.Rmd
profiled get_us_covid
tests

removed testes

removed testes2

removed tests covid

test_getus

us_test separated

prodiled function get_covid and fixed tests

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.1.200-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.1.300-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2020-12-17 R/getus_functions.R

@@ -3,8 +3,6 @@
 #' extracts time series from the git repository of the \href{ https://github.com/CSSEGISandData }{JHU}
 #'
 #' @return a dataframe
-#' @importFrom rlang .data
-#' @importFrom magrittr %>%
 #' @import vroom
 #' @details `cases` represents the number of confirmed cases, while `cmr` the case-mortality rate (deaths / confirmed_case * 100).
 #' A good description of pitfalls and caveats associated with the use of case-mortality rate metric has been made on
@@ -22,37 +20,51 @@ getus_covid_jhu <- function() {
       stop("Something wrong with the repository or your internet connection!")
     }
 
-    dat <-
-      vroom(url_full, col_types = c(.default = "?"), progress = FALSE) %>%
-      tidyr::pivot_longer(cols = dplyr::matches("[0-9]{1,}/"), names_to = "date", values_to = "value") %>%
-      janitor::clean_names() %>%
-      dplyr::filter(.data$country_region == "US") %>%
-      dplyr::select(
-        .data$date, .data$combined_key, .data$fips, .data$value
-      ) %>%
-      dplyr::mutate("fips" = as.numeric(.data$fips)) %>%
-      # this is to separte the column (tidyr::separate occasionally trows error)
-      dplyr::mutate("county_state" = gsub(",(.)?US$", "", .data$combined_key, perl = TRUE)) %>%
-      dplyr::mutate("state" = ifelse(
-        grepl(",", .data$county_state) == FALSE,
-        .data$county_state,
-        gsub(".*,(\\s)?", "", .data$county_state, perl = TRUE)
-      )) %>%
-      dplyr::mutate("county" = ifelse(
-        grepl(",", .data$county_state) == FALSE,
-        NA,
-        gsub(",.*", "", .data$county_state, perl = TRUE)
-      )) %>%
+    dat <- vroom(url_full, col_types = c(.default = "?"), progress = FALSE)
 
-      # JHU has unincorporated U.S territories and the cruises data
-      # that ends up to be NA because they have not counties in the dataframe
-      dplyr::filter(!is.na(.data$county), !is.na(.data$fips)) %>%
-      dplyr::filter(.data$state != "Puerto Rico") %>%
-      dplyr::select(-.data$county_state)
[…]

### 2020-12-17 tests/testthat/test-getus_tests.R

@@ -20,35 +20,13 @@ expected_states <- structure(list(state = c(
 )), class = c("data.frame"), row.names = c(NA, -51L))
 
 expected_states <- sort(expected_states$state)
-#
-#
-# y_jhu <- getus_all()
-#
-#
-x_jhu <- getus_covid(repo = "jhu")
-# #
-# #
-# # test_that("get_all_jhu and get_covid have same fips", {
-# #   expect_equal(length(unique(y_jhu$fips)), length(unique(x_jhu$fips)))
-# # })
-# # #
-#
-test_that("getus_jhu states abbr ", {
-  expect_equal(sort(unique(x_jhu$state)), expected_states )
+
+tests <- getus_tests() %>%
+  select(state, abbr) %>%
+  distinct() %>%
+  arrange(state) %>%
+  pull(state)
+
+test_that("getus_tests states abbr ", {
+    expect_equal(tests, expected_states)
 })
-
-# # ##################################
-
-# test_that("getus_test executes", {
-#   expect_error(getus_tests(), NA)
-# })
-#
-# tests <- getus_tests() %>%
-#   select(state, abbr) %>%
-#   distinct() %>%
-#   arrange(state) %>%
-#   pull(state)
-#
-# test_that("getus_tests states abbr ", {
-#     expect_equal(tests, expected_states)
-# })

### 2021-02-18 README.Rmd

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.1.300-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.1.400-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2021-03-03 R/datasets_it.R
merge data.table is more efficient

@@ -133,6 +133,6 @@
 #' @docType data
 #' @usage data("it_pm2.5")
 #' @source \href{https://annuario.isprambiente.it/pon/basic/14}{Istituto Superiore Per La protezione Ambientale}
-#' @details \href{http://siqual.istat.it/SIQual/visualizza.do?id=5000170&refresh=true&language=EN}{methodology}
+#' @details \href{https://annuario.isprambiente.it/pon/basic/14}{methodology}
 #' @return a tibble
 "it_pm2.5"

### 2021-03-03 R/getus_functions.R

@@ -68,15 +68,15 @@ getus_covid_jhu <- function() {
   })
 
   names(dat_l) <- metric_files
+  dat_l <- lapply(dat_l, data.table::as.data.table)
 
-  # diff_dat <- dat_w[which(dat_w$county.y != dat_w$county.x),]
 
   # funny things is that there are unassigned county (confirmed 90049),
   # that in the same file have a county
 
   dat_w <- dat_l$confirmed[, c("date", "fips", "confirmed")]
 
-  dat_w_j <- merge(dat_w, dat_l$deaths, by = c("date", "fips"))
+  dat_w_j <- data.table::merge.data.table(dat_w, dat_l$deaths, by = c("date", "fips"))
 
   names(dat_w_j)[names(dat_w_j) == "confirmed"] <- "cases"
 
@@ -87,7 +87,7 @@ getus_covid_jhu <- function() {
 
   message(paste0("US COVID-19 data up to ", max(dat_out$date), " successfully retrived from JHU repository!"))
 
-  dat_out
+  as.data.frame(dat_out)
 }
 
 
@@ -506,35 +506,52 @@ getus_tests <- function() {
 #' @details For details regarding some specific datasets refer to: \href{https://www2.census.gov/programs-surveys/acs/tech_docs/subject_definitions/2018_ACSSubjectDefinitions.pdf?#}{Subject Definitions of the American Community Survey},
 #'  \href{https://www.cms.gov/About-CMS/Agency-Information/OMH/Downloads/Mapping-Technical-Documentation.pdf}{Medicare and Medicaid Medical Services Technical Documentation},
 #'  \href{https://github.com/COVIDExposureIndices/COVIDExposureIndices}{COVIDExposureIndices}
-#' @importFrom rlang .data
-#' @importFrom magrittr %>%
 #' @import vroom
 #' @seealso \code{\link{getus_covid}},\code{\link{getus_tests}}, \code{\link{getus_dex}},
 #' @export
 getus_all <- function(repo = "jhu") {
   covid19_us <- getus_covid(repo = repo)
-  dex_us <- getus_dex() %>%
-    dplyr::select(.data$fips, .data$date, .data$dex_a)
+  dex_us <- getus_dex()
+  dex_us_sel <- dex_us[, c("fips", "date", "dex_a")]
 
-  tests_us <- getus_tests() %>%
-    dplyr::select(-.data$death, -.data$death_increase, -.data$abbr, -.data$hash, -.data$fips)
+
[…]

### 2021-03-03 tests/testthat/test-getus_jhu.R
travis cannot handle tests

@@ -1,41 +1,41 @@
-library(dplyr)
-expected_states <- structure(list(state = c(
-  "Alaska", "Alabama", "Arizona", "Arkansas",
-  "California", "Colorado", "Connecticut", "District of Columbia",
-  "Delaware", "Florida", "Georgia", "Hawaii", "Iowa", "Idaho",
-  "Illinois", "Indiana", "Kansas", "Kentucky", "Louisiana", "Massachusetts",
-  "Maryland", "Maine", "Michigan", "Minnesota", "Missouri", "Mississippi",
-  "Montana", "North Carolina", "North Dakota", "Nebraska", "New Hampshire",
-  "New Jersey", "New Mexico", "Nevada", "New York", "Ohio", "Oklahoma",
-  "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "South Dakota",
-  "Tennessee", "Texas", "Utah", "Virginia", "Vermont", "Washington",
-  "Wisconsin", "West Virginia", "Wyoming"
-), abbr = c(
-  "AK", "AL",
-  "AR", "AZ", "CA", "CO", "CT", "DC", "DE", "FL", "GA", "HI", "IA",
-  "ID", "IL", "IN", "KS", "KY", "LA", "MA", "MD", "ME", "MI", "MN",
-  "MO", "MS", "MT", "NC", "ND", "NE", "NH", "NJ", "NM", "NV", "NY",
-  "OH", "OK", "OR", "PA", "RI", "SC", "SD", "TN", "TX", "UT", "VA",
-  "VT", "WA", "WI", "WV", "WY"
-)), class = c("data.frame"), row.names = c(NA, -51L))
-
-expected_states <- sort(expected_states$state)
+# library(dplyr)
+# expected_states <- structure(list(state = c(
+#   "Alaska", "Alabama", "Arizona", "Arkansas",
+#   "California", "Colorado", "Connecticut", "District of Columbia",
+#   "Delaware", "Florida", "Georgia", "Hawaii", "Iowa", "Idaho",
+#   "Illinois", "Indiana", "Kansas", "Kentucky", "Louisiana", "Massachusetts",
+#   "Maryland", "Maine", "Michigan", "Minnesota", "Missouri", "Mississippi",
+#   "Montana", "North Carolina", "North Dakota", "Nebraska", "New Hampshire",
[…]

### 2021-03-04 README.Rmd
retriving data message

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.1.400-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.1.500-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2021-03-04 R/getus_functions.R

@@ -12,6 +12,8 @@ getus_covid_jhu <- function() {
   url_base <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"
   metric_files <- list("confirmed", "deaths")
 
+  message("Retriving data...please wait!")
+
   # JHU has 2 files with date in long format (confirmed and deaths)...whatever
   dat_l <- lapply(metric_files, function(x) {
     url_full <- paste0(url_base, x, "_US.csv")
@@ -105,6 +107,8 @@ getus_covid_jhu <- function() {
 #' @keywords internal
 getus_covid_nyt <- function() {
   url_data <- "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"
+
+  message("Retriving data...please wait...!")
 
   if (RCurl::url.exists(url_data) == FALSE) {
     stop("Something wrong with the repository or your internet connection!")

### 2021-03-04 README.Rmd

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.1.500-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/devel%20version-0.0.1.600-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)

### 2021-05-18 R/datasets_us.R
version bump, typos

@@ -1,6 +1,6 @@
 #' influenza vaccination 65 or older
 #'
-#' Percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination. Collected in 2017.
+#' Percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination. Collected in 2019.
 #' @docType data
 #' @source \href{https://data.cms.gov/mapping-medicare-disparities}{Data.CMS.gov}
 #' @details \href{https://www.cms.gov/About-CMS/Agency-Information/OMH/Downloads/Mapping-Technical-Documentation.pdf}{Center for Medicare and Medicaid Services} and NORC at the University of Chicago.

### 2021-05-18 R/getus_functions.R

@@ -13,7 +13,7 @@ getus_covid_jhu <- function() {
   url_base <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_"
   metric_files <- list("confirmed", "deaths")
 
-  message("Retriving data...please wait!")
+  message("Retrieving data...please wait!")
 
   # JHU has 2 files with date in long format (confirmed and deaths)...whatever
   dat_l <- lapply(metric_files, function(x) {
@@ -109,7 +109,7 @@ getus_covid_jhu <- function() {
 getus_covid_nyt <- function() {
   url_data <- "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"
 
-  message("Retriving data...please wait...!")
+  message("Retrieving data...please wait...!")
 
   if (RCurl::url.exists(url_data) == FALSE) {
     stop("Something wrong with the repository or your internet connection!")

### 2021-05-18 README.Rmd

@@
 
-[![](https://img.shields.io/badge/devel%20version-0.0.1.600-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
+[![](https://img.shields.io/badge/version-0.1.0-blue.svg)](https://codecov.io/gh/c1au6i0/covid19census?branch=master)
 [![Build Status](https://travis-ci.com/c1au6i0/covid19census.svg?branch=master)](https://travis-ci.com/c1au6i0/covid19census)
@@
 * `us_age_sex`: age and sex distribution (2018).
-* `us_fl65`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2017)
+* `us_fl65`:  percentage of fee-for-service (FFS) Medicare enrollees that had an annual flu vaccination (2019)
 * `us_hospbeds`: total hospital beds in each county (2019).
 * `us_netinc`: median househould income (2018).
-* `us_mmd`: data of 2017 regarding the  prevalence found in medicare beneficiaries of 
+* `us_mmd`: data of 2019 regarding the  prevalence found in medicare beneficiaries of 
     * many medical and chronic conditions (Alzheimer, chronic kidney, obesity, depression, obstructive pulmonary, disease, arthritis, diabetes, osteoporosis, asthma, atrial  fibrillation, ischemic hearth, myocardial infarction, hypertension).
